### 2015-08-06 paper/frontiers.tex
abstract

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.                                                 %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.1 Generated 2015/22/05 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%

\documentclass{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
%\documentclass{frontiersHLTH} % for Health articles
%\documentclass{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles

%\setcitestyle{square}
\usepackage{url,hyperref,lineno,microtype}
\usepackage[onehalfspacing]{setspace}
\linenumbers

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Stewart {et~al.}} %use et al only if is more than 1 author
\def\Authors{Terrence C. Stewart\,$^{1,*}$, Ashley Kleinhans\,$^{2}$ and Chris Eliasmith\,$^1$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Centre for Theoretical Neuroscience, University of Waterloo,
    Waterloo, ON, Canada
    \\
$^{2}$Mobile Intelligent Autonomous Systems group, 
      Council for Scientific and Industrial Research,
      Pretoria, South Africa}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Terrence C. Stewart}
\def\corrAddress{Centre for Theoretical Neuroscience, University of Waterloo,
    Waterloo, ON, Canada}
\def\corrEmail{tcstewar@uwaterloo.ca}

\begin{document}
\onecolumn
\firstpage{1}

\title[Embodied Neuromorphic Benchmarks]{Embodied Neuromorphic Benchmarks}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% The sections below are for reference only.
%%%
%%% For Original Research Articles, Clinical Trial Articles, and Technology Reports the section headings should be those appropriate for your field and the research itself. It is recommended to organize your manuscript in the
%%% following sections or their equivalents for your field:
%%% Abstract, Introduction, Material and Methods, Results, and Discussion.
%%% Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
%%%
%%%For information about Clinical Trial Registration, please go to http://www.frontiersin.org/about/AuthorGuidelines#ClinicalTrialRegistration
%%%
%%% For Clinical Case Studies the following sections are mandatory: Abstract, Introduction, Background, Discussion, and Concluding Remarks.
%%%
%%% For all other article types there are no mandatory sections.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

    Evaluating the effectiveness and performance of neuromorphic hardware is
    difficult.  It is even more difficult when the task of interest is an
    embodied task; that is a task where the output from the neuromorphic
    hardware affects its future input.  However, these embodied situations
    are one of the primary potential uses of neuromorphic hardware.  To address
    this, we present a methodology for embodied benchmarking and a particular
    example of a benchmark for adaptive control of an arbitrary system.  These
    benchmarks are flexible, in that they allow researchers to explicitly
    modify the benchmark to identify particular task domains where particular
    hardware excels.  Furthermore, the benchmarks make use of a hybrid of
    real physical embodiment and a type of "minimal" simulation that has been
    shown to lead to robust real-world performance, allowing the same benchmark
    to be used by many researchers.

\tiny
 \keyFont{ \section{Keywords:} neuromorphic hardware, benchmarking, 
        minimal simulation, adaptive control, neural networks} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\end{abstract}

\section{Introduction}

% For Original Research Articles, Clinical Trial Articles, and Technology Reports the introduction should be succinct, with no subheadings.
%
% For Clinical Case Studies the Introduction should include symptoms at presentation, physical exams and lab results.
%
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text. Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text. Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text.Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text.

%\begin{methods}
\section{Material \& Methods}

Text Text Text Text Text Text  Text Text Text Text Text Text \cite{conference} Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.  \cite{patent} might want to know about  text text text text Text Text Text Text  Text Text Text Text Text Text  Text Text. \citep{article} might want to know about  text text text text
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.  \cite{book} might want to know about  text text text text \cite{chapter}

\begin{table}[!t]
\textbf{\refstepcounter{table}\label{Tab:01} Table \arabic{table}.}{ Maximum size of the Manuscript }

\processtable{ }
{\begin{tabular}{lllll}\toprule
 & Abstract max. legth (incl. spaces) & Figures or tables & Manuscript max. length \\\midrule
Clinical Case Study & & & &\\
Clinical Trial & & & &\\
Hypothesis and Theory & & & &\\
Methods & 2000 characters  & 15 & 12000 words \\
Original Research & & & &\\
Review & & & &\\
Technology Report & & & &\\\midrule
Focused Review & 2000 characters & 5 & 5000 words \\\midrule
CPC &  1250 characters& 6 & 2500 words  \\\midrule
Perspective & 1250 characters & 2 & 3000 words  \\
Mini Review & & & &\\\midrule
Data Report & None & 2 & 3000 words\\\midrule
Classification & 1250 characters & 10 & 2000 words \\\midrule
Editorial & None & None & 1000 words  \\\midrule
Frontiers Commentary  & & &\\
General Commentary & None & 1 & 1000 words\\
Book review & & & \\\midrule
Opinion   & & &\\
Specialty Grand Challenge & None & 1 & 2000 words\\
Field Grand Challenge & & & &\\\botrule
\end{tabular}}{}
\end{table}

Please note that large tables covering several pages cannot be included in the final PDF for formatting reasons. These tables will be published as supplementary material on the online article abstract page at the time of acceptance. The author will notified during the typesetting of the final article if this is the case. A link in the final PDF will direct to the online material.

\subsection{Original Research Articles, Clinical Trial Articles, and Technology Reports}

For Original Research Articles, Clinical Trial Articles, and Technology Reports the section headings should be those appropriate for your field and the research itself. It is recommended to organize your manuscript in the following sections or their equivalents for your field:

\begin{itemize}
%for bulleted list, use itemize
\item Introduction: Succinct, with no subheadings.
\item Materials and Methods: This section may be divided by subheadings. This section should contain sufficient detail so that when read in conjunction with cited references, all procedures can be repeated.
\item Results: This section may be divided by subheadings. Footnotes should not be used and have to be transferred into the main text.
\item Discussion: This section may be divided by subheadings. Discussions should cover the key findings of the study: discuss any prior art related to the subject so to place the novelty of the discovery in the appropriate context; discuss the potential short-comings and limitations on their interpretations; discuss their integration into the current understanding of the problem and how this advances the current views; speculate on the future direction of the research and freely postulate theories that could be tested in the future.
\end{itemize}

Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.

\subsection{Clinical Case Studies}

For Clinical Case Studies the following sections are mandatory:

\begin{itemize}
%for bulleted list, use itemize
\item Introduction: Include symptoms at presentation, physical exams and lab results.
\item Background: This section may be divided by subheadings. Include history and review of similar cases.
\item Results: This section may be divided by subheadings. Include diagnosis and treatment.
\item Concluding Remarks
\end{itemize}

%\end{methods}

\section{Results}

\subsection{Figures}
Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript. Figures will then be automatically embedded at the bottom of the submitted manuscript. Kindly ensure that each table and figure is mentioned in the text and in numerical order. Permission must be obtained for use of copyrighted material from other sources (including the web). Please note that it is compulsory to follow figure instructions. Figures which are not according to the guidelines will cause substantial delay during the production process.

\begin{table}[!t]
\textbf{\refstepcounter{table}\label{Tab:02} Table \arabic{table}.}{ Resolution Requirements for the figures}

\processtable{}
{\begin{tabular}{lllll}\toprule
Image Type & Description & Format & Color Mode & Resolution\\\midrule
Line Art & An image composed of lines and text,  & TIFF, JPEG & RGB, Bitmap & 900 - 1200 dpi\\
           & which does not contain tonal or shaded areas.& & &\\
           Halftone & A continuous tone photograph, which contains no text. & TIFF, EPS, JPEG & RGB, Grayscale & 300 dpi\\
Combination & Image contains halftone + text or line art elements. & TIFF, JPEG & RGB,Grayscale & 600 - 900 dpi\\\botrule
\end{tabular}}{}
\end{table}

\textbf{Table \ref{Tab:02}} shows the resolution requirements for the figures. The figures must be legible:
\begin{enumerate}
\item The smallest visible text is no less than 8 points in height, when viewed at actual size.
\item Solid lines are not broken up.
\item Image areas are not pixelated or stair stepped.
\item Text is legible and of high quality.
\item Any lines in the graphic are no smaller than 2 points width.
\item The actual size of the figure must be of at least 8.5 cm.
\end{enumerate}

\subsection{Nomenclature}
\begin{itemize}
\item The use of abbreviations should be kept to a minimum. Non-standard abbreviations should be avoided unless they appear at least four times, and defined upon first use in the main text. Consider also giving a list of non-standard abbreviations at the end, immediately before the Acknowledgments.
\item Gene symbols should be italicized; protein products are not italicized.
\item Chemical compounds and biomolecules should be referred to using systematic nomenclature, preferably using the recommendations by IUPAC.
\item We encourage the use of Standard International Units in all manuscripts.
\item To take part in the Resource Identification Initiative, please cite antibodies, genetically modified organisms, software tools, data, databases and services using the corresponding catalog number and RRID in your current manuscript. For more information about the project and for steps on how to search for an RRID, please click \href{http://www.frontiersin.org/files/pdf/letter_to_author.pdf}{here}.
\end{itemize}

\begin{equation}
\sum x+ y =Z\label{eq:01}
\end{equation}

\section{Discussion}

Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text.
Additional Requirements:
\subsection{Corrections}

If you need to communicate important changes to a published article please submit a General Commentary. Submit the article with the title “Corrigendum: Original Title of Article”.

\subsection{Commentaries on Articles}

At the beginning of your Commentary, please provide the citation of the article commented on. Rebuttals may be submitted in response to Commentaries; our limit in place is one commentary and one response. Rebuttals should also be submitted as General Commentary articles.

\subsection{Human Search and Animal Research}

All experiments on live vertebrates or higher invertebrates must be performed in accordance with relevant institutional and national guidelines and regulations. In the manuscript, authors must identify the committee approving the experiments and must confirm that all experiments conform to the relevant regulatory standards. For manuscripts reporting experiments on human subjects, authors must identify the committee approving the experiments and must also include a statement confirming that informed consent was obtained from all subjects. In Original Research Articles and Clinical Trial Articles these statements should appear in the Materials and Methods section.

\subsection{Clinical Trial Registration}

Clinical trials should be registered in a public trials registry in order to become the object of a publication at Frontiers. Trials must be registered at or before the start of patient enrollment. A clinical trial is defined as "any research study that prospectively assigns human participants or groups of humans to one or more health-related interventions to evaluate the effects on health outcomes."(\href{www.who.int/ictrp/en}{www.who.int/ictrp/en}). A list of acceptable registries can be found at \href{www.who.int/ictrp/en}{www.who.int/ictrp/en} and \href{www.icmje.org}{www.icmje.org}.

\subsection{Inclusion of Proteomics Data}

Authors should provide relevant information relating to how the peptide/protein matches were undertaken, including methods used to process and analyze data, false discovery rates (FDR) for large-scale studies and threshold or cut-off rates for peptide and protein matches. Further information could include software used, mass spectrometer type, sequence database and version, number of sequences in database, processing methods, mass tolerances used for matching, variable/fixed modifications, allowable missed cleavages, etc.

Authors should provide as supplementary material information used to identify proteins and/or peptides. This should include information such as accession numbers, observed mass (m/z), charge, delta mass, matched mass, peptide/protein scores, peptide modification, miscleavages, peptide sequence, match rank, matched species (for cross species matching), number of peptide matches, ambiguous protein/peptide matches should be indicated, etc.
For quantitative proteomics analyses authors should provide information to justify the statistical significance including biological replicates, statistical methods, estimates of uncertainty and the methods used for calculating error.

For peptide matches with biologically relevant post-translational modifications (PTM) and for any protein match that has occurred using a single mass spectrum, authors should include this information as raw data, annotated spectra or submit data to an online repository (recommended option).
Authors are encouraged to submit raw or matched data and 2-DE images to public proteomics repositories. Submission codes and/or links to data should be provided within the manuscript.

\subsection{Data Sharing}

Frontiers supports the policy of data sharing, and authors are advised to make freely available any materials and information described in their article, and any data relevant to the article (while not compromising confidentiality in the context of human-subject research) that may be reasonably requested by others for the purpose of academic and non-commercial research. In regards to deposition of data and data sharing through databases, Frontiers urges authors to comply with the current best practices within their discipline.

\section*{Disclosure/Conflict-of-Interest Statement}
%Frontiers follows the recommendations by the International Committee of Medical Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require that all financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed. If no such relationship exists, authors will be asked to declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest. When disclosing the potential conflict of interest, the authors need to address the following points:
%•	Did you or your institution at any time receive payment or services from a third party for any aspect of the submitted work?
%•	Please declare financial relationships with entities that could be perceived to influence, or that give the appearance of potentially influencing, what you wrote in the submitted work.
%•	Please declare patents and copyrights, whether pending, issued, licensed and/or receiving royalties relevant to the work.
%•	Please state other relationships or activities that readers could perceive to have influenced, or that give the appearance of potentially influencing, what you wrote in the submitted work.

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}
%When determining authorship the following criteria should be observed:
%•	Substantial contributions to the conception or design of the work; or the acquisition, analysis, or interpretation of data for the work; AND
%•	Drafting the work or revising it critically for important intellectual content; AND
%•	Final approval of the version to be published ; AND
%•	Agreement to be accountable for all aspects of the work in ensuring that questions related to the accuracy or integrity of any part of the work are appropriately investigated and resolved.
%Contributors who meet fewer than all 4 of the above criteria for authorship should not be listed as authors, but they should be acknowledged. (http://www.icmje.org/roles_a.html)

The statement about the authors and contributors can be up to several sentences long, describing the tasks of individual authors referred to by their initials and should be included at the end of the manuscript before the References section.

\section*{Acknowledgments}
Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text. Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.

\textit{Funding\textcolon} Text Text Text Text Text Text  Text Text.

\section*{Supplemental Data}
Supplementary Material should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure. LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder

Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.

\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
\bibliography{test}

%%% Upload the *bib file along with the *tex file and PDF on submission if the bibliography is not in the main *tex file

\section*{Figures}

%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.jpg and logo2.eps) are in the Frontiers LaTeX folder
%%% If using *.tif files convert them to .jpg or .png

\begin{figure}[h!]
\begin{center}
\includegraphics[width=10cm]{logo1}% This is a *.jpg file
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
\end{figure}

%\begin{figure}
%\begin{center}
%\includegraphics[width=10cm]{logo2}% This is an *.eps file
%\end{center}
%\textbf{\refstepcounter{figure}\label{fig:02} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
%\end{figure}

%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.

%%% Frontiers will add the figures at the end of the provisional pdf automatically %%%

%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.

\end{document}

### 2015-08-06 paper/paper.tex
intro

--- 2015-08-06 paper/frontiers.tex
@@
     difficult.  It is even more difficult when the task of interest is an
-    embodied task; that is a task where the output from the neuromorphic
+    embodied task; that is, a task where the output from the neuromorphic
     hardware affects its future input.  However, these embodied situations
@@
 \section{Introduction}
+
+Neuromorphic hardware holds great promise for a wide variety of applications.
+The combination of massively parallel computation and low power consumption
+means that there is the potential to have complex algorithms running in
+embedded processing situations, without being a significant drain on battery
+life.  The question, of course, is to identify what sort of always-on or
+interactive functionality is feasible with these devices.
+
+To evaluate applications of this hardware, we need benchmark tasks.  These
+tasks must allow us to compare across different instances of neuromorphic
+hardware (and potentially across different algorithms implemented in said
+hardware).  This allows us to quantitatively compare systems, letting 
+researchers both measure the progress in the field and also directly
+compare competing approaches.
+
+In this paper, we focus on the development of \emph{embodied} benchmarks.
+These are tasks where the output of the neuromorphic hardware \emph{influences
+its own future input}.  This is in contrast to standard categorization or pattern
+identification tasks, where the input is some fixed sequence, and the main
+question is whether the hardware produces the correct output for each input
+(or input pattern).
+
+We believe embodied benchmarks
 

### 2015-08-08 paper/paper.tex
intro

@@
 
-We believe embodied benchmarks
+We believe embodied benchmarks should be of particular interest to
+neuromorphic research, given that many of the applications of neuromorphic
+hardware are in exactly this domain of embedded and interactive control
+of robotic or other physical systems.  However, embodiment raises a number
+of issues that complicate the development of such benchmarks.  Rather than
+simply providing a data file of inputs and desired outputs, the benchmark
+must either specify a full physical system for that embodiment, or it must 
+provide software for a simulation of that system.  As we discuss below, either
+approach is problematic, and addressing this problem is the primary goal of
+this paper.
+
+\section{Embodied Benchmarks}
+
+\subsection{Simulation versus Physical Instantiation}
+
+\subsection{Minimal Simulation}
+
+\subsection{Cheap Robotics}
 

### 2015-08-08 paper/paper.tex
sim versus phys

@@
 provide software for a simulation of that system.  As we discuss below, either
-approach is problematic, and addressing this problem is the primary goal of
+approach is problematic, and addressing this difficulty is the primary goal of
 this paper.
@@
 
+An embodied benchmark task is one where the system we are studying has a 
+two-way interaction with some sort of environment.  That is, the outputs from
+the neuromorphic hardware are sent to the environment where they cause some
+sort of effect, the results of which change the subsequent input.  For example,
+the outputs might control the movement of a robot, which in turn affects the
+sensory data received by the robot.
+
 \subsection{Simulation versus Physical Instantiation}
+
+To define such a benchmark, we need to be explicit about the embodiment.  If
+it is a situation where a robot is to be controlled, we need to be explicit
+about all the details of that robot.  What motors are there?  How are they
+configured?  How strong are they?  What sensors are there?  Where are they
+placed?  How accurate are they?  However, even if these questions are
+answered, there is a fundamental problem in that \emph{other researchers
+need access to that exact robot}.  If a benchmark is to be widely used, other
+researchers developing their own hardware should be able to do their own 
+testing.
+
+Furthermore, using a physical robot also imposes significant practical
+difficulties when performing extensive benchmark testing.  When testing, we
+often want to run the same task over and over again, both for robustness and
+to see the effects of varying parameters.  With a physical robot, this means
+manually setting up the task, letting the test run, gathering the resulting
+data, and then resetting the robot back to the initial state.  This means that
+issues like battery life become problematic, and not just because there is a
+limited amount of time available for testing.  As the battery level changes,
+the performance of the robot itself can also change.  Futhermore, for any
+rigorous testing of the benchmark, we will want to examine situations where
+the system fails.  This means that some of the testing will involve parameter
+settings that lead to poor behaviour, which could involve physical damage to
+the robot itself.
+
+However, \emph{not} using a real physical embodiment for testing is also
+problematic.  First and foremost, without an actual real-world task, why
+should anyone have any confidence that the performance on the benchmark is
+reflective of the actual usefulness of the neuromorphic hardware?  It is
+very widely known that \emph{simulations of robots} (or other physical systems) are
+often much easier to control and better-behaved than the real thing.  The
+field of robotics is filled with algorithms that work well ``in theory'', but
+fail when run on actual hardware.  We do not want a benchmark that falls into
+this trap of failing to generalize to real situations.
+
+Furthermore, neuromorphic hardware has another constraint that severely limits
+the possibilities of simulation.  The normal approach when a simulation is
+too simple to reflect reality is to add details to the simulation itself.
+Incredibly finely detailed simulations can be created, filling in all of the
+details needed.  However, those resulting simulations \emph{cannot be run
+in real-time}.  This is a fundamental problem, in that most neuromorphic
+hardware is about real-time interactions, and there is no way to slow down
+the hardware to match the simulated environment.  This means that even if
+we spent the considerable amount of research effort needed to define a
+simulated environment for an embodied benchmark, that simulation could not
+be run fast enough to interact with most of the hardware that we would want
+to test.
 

### 2015-08-08 paper/paper.tex
bibliography

@@
 \subsection{Minimal Simulation}
+
+The above considerations seem to indicate that even though using real-world
+phyiscal hardware for benchmarking is problematic, it is still better than
+using simplistic simulations which may not generalize to real tasks.  However,
+we believe there is an alternate approach known as \emph{Minimal Simulation}
+\cite{Jakobi97evolutionaryrobotics}.
 
@@
 %\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
-\bibliography{test}
+\bibliography{paper}
 

### 2015-08-09 paper/paper.tex
minimal sim

@@
 
+First, we note that the problem faced here is remarkably similar to a problem
+faced by the evolutionary robotics community.  In evolutionary robotics, the
+goal is to use genetic algorithms to \emph{evolve} systems that can control
+robots to perform various tasks.  These tasks can be as simple as navigation
+and obstacle avoidance, but have also included complex eight-legged walking,
+object identification, and visual tracking [refs???].
+
+However, performing this evolution on real physical robots is problematic
+for the same reasons that benchmarks on physical robots are problematic.  The
+robots must be reset to the same state each time; they often involve
+behaviour that can physically damage the robots; and they take a very long
+time to run.  For this reason, attempts were made to evolve using simulated
+robots.  However, the general finding was that algorithms that worked on the
+simulated robots would not work when run on the real physical robots.  If
+the simulations were improved, adding complex physical detail, then it was
+possible to generalize to real behaviour; unfortunately, such complex
+simulations would run slower than real-time [refs].
+
+To address this problem, Nicholas Jacobi proposed the creation of
+``minimal'' simulations.  These are simulations where there is variability
+\emph{within the simulation itself}.  In other words, we make \emph{poor}
+simulations, but ensure that the way in which it is poor is itself
+variable.  We then make sure that the controllers work across that whole
+range of variability.  ``Instead of trying to eliminate the differences between
+simulation and reality, they are acknowledged, and mechanisms are put in place
+to prevent evolving controllers from relying on them.'' [ref: jacobi thesis].
+
+With this approach, it was possible to build minimal simulations that would
+run faster than real-time and yet also be complex enough that if a system
+could successfully control the simulation, it was also likely to successfully
+control a real robot.  To achieve this, the simulations are made to be
+unreliable in almost every respect.  For example, for a simulation of a simple
+motor it would still be the case that if power is applied it would generally
+try to spin, but the exact amount of torque, the amount of sensory noise,
+the amount of time needed, the amount of static and dynamic friction, and so
+on would all be randomly chosen.  A successful controller would have to deal
+with this wide range of variability, and if it could handle that variability
+then we would have reason to believe it could also handle the real system.
+
+It is also worth noting that a minimal simulation \emph{only has to be a
+good simulation for successful behaviour}.  That is, ``if we are evolving 
+corridor following behaviour, the dynamics of the simulation might
+differ wildly from those of reality if the controller hits a wall or
+goes round in circles, but this does
+not matter, since the controllers we are interested in transferring
+across the reality gap will neither
+hit walls nor go round in circles.''  If the controller is poor, we do not 
+need the simulation to be at all accurate in exactly \emph{how} that poor
+behaviour is manifest.  We do not need an exact detailed physics model of
+the collision between a robot and a wall, or a detailed model of what happens
+to a robot arm when it starts oscillating wildly due to a poor control signal.
+All we need is for the simulation to be just good enough to indicate that
+things have gone wrong, and thus give a low score to that controller.
+
+Given the success of this approach for evolutionary robotics, we propose
+that it can be directly used for embodied neuromorphic benchmarks.  To do
+this, we create software simulations for each benchmark task.  These simulations
+must be fast enough to run in realtime (so that they can be controlled by
+real neuromorphic hardware), and they must be extremely variable.  Each time
+the simulation is run, different parameters will be chosed for this variability
+(so one run might have a large degree of sensor noise while the next run has
+none at all; one run might have more delay in the motor response and another
+might have less power available).  Being successful at the benchmark means
+being successful across all this variability.
+
+The result should be a benchmark that can be run by any researcher.  The fact
+that it is a simulation means that source code can be shared, and that no
+specialized hardware is needed.  Furthermore, the variability in the simulation
+itself can be controlled, and this can help give a rich characterization of
+the benchmarked hardware.  For example, some hardware might only work with
+small amounts of sensor noise, or other hardware might only work when there
+is significant delay in the motor response.  This flexibility in parameters
+in the benchmark allows researchers to explicitly characterize that particular
+siuations where their hardware excells.
+
 \subsection{Cheap Robotics}
+
+All of that said, we also need real physical embodiment as part of any
+benchmark of this kind.  While we claim that reliable minimal simulations
+are relatively easy to create, they are not a widely-used technique, and the
+real proof of behaviour is always in the real world.  For this reason, we
+also argue that every benchmark based on a minimal simulation should also
+have an easy-to-construct physical analog.  However, this physical version
+is not to be used as the primary benchmark.  Rather, this is meant as a
+double-check that the system does actually behave as expected.  The worry
+is always that if researchers focus on the simulated system when benchmarking
+their hardware, they may inadvertantly end up with hardware that only performs
+well in simulation.
+
+For this physical aspect of the benchmark, we recommend cheap, widely-available
+components.  This allows a greater chance for other researchers to have
+access to the same (or similar) hardware.  For the particular example 
+benchmark described in the next section, we use the Lego Mindstorms EV3 kit,
+a simple robotics platform available at most toy stores.
+
+It is important to note that there is actually a theoretical advantage to
+using cheap robotics hardware for benchmarking, in addition to the practical
+advantages.  In particular, we \emph{don't want benchmarks that rely on
+particular high-speed, high-accuracy devices}.  The purpose of the benchmark
+is not to indicate how well this neuromorphic hardware works to control this
+one particular robot in this task.  Rather, the purpose of a benchmark is to
+characterise how well this neuromorphic hardware works on this task 
+\emph{in general}.  The variability in the minimal simulation means that it
+should be able to function across a wide variety of physical embodiments, and
+so if we are to choose one particular physical embodiment to test in the
+real world, then we should choose one that is not extremely high-precision.
+For this reason, we believe using cheap Lego robot is actually more useful
+for benchmarking than an expensive high-precision robot.\footnote{Of course,
+for more complex benchmark tasks we may need sensory and motor capabiities
+that are beyond that of a simple Lego robot.}
+
+\section{Embodied Benchmark 1: Adaptive Motor Control}
 

### 2015-08-10 paper/paper.tex
adaptive control

@@
 
-\section{Embodied Benchmark 1: Adaptive Motor Control}
+\section{A Benchmark: Adaptive Motor Control}
+
+To demonstrate this approach to creating embodied neuromorphic benchmarks,
+we now consider a basic control task.  Suppose we have a system with a number
+of joints $q$ and we want to send an output $u$ to those motors such that
+the joints move to a particular desired position $q_d$.  Our only output is
+the signal $u$ (one for each motor) and our only input is the current position
+of each motor $q$.
+
+The simplest controller for such a situation is a P (proportional) controller,
+where $u=K_p(q_d - q)$.  This is often supplemented with a derivative term,
+which helps to slow the system down as it approaches the desired position,
+thus avoiding overshooting and oscillation ($u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$),
+leading to the standard PD controller.  Both $K_p$ and $K_d$ are constants that can be
+hand-tuned to particular situations.
+
+However, this controller has difficulty in the presence of significant external
+forces.  For example, consider a single motor controlling the angle of a single
+arm.  If the arm is held out to the side, gravity acting on the mass of the
+arm itself will pull the arm downward.  Thus to hold the arm still at a
+particular $q_d$ will require the controller to apply a force to counteract
+gravity.  Since the PD controller always produces an output $u=0$ when $q=q_d$,
+it cannot compensate for this. [TODO: add diagram]
+
+The standard solution to this problem is to add an integral term ($K_i \int{q_d-q dt}$)
+to the controller, making it a PID controller.  The idea here is that as the difference between where it is and
+where we want it to be accumulates over time, the $K-i$ term gradually increases
+how much extra force is being applied until it is large enough to counteract
+the external force of gravity (or whatever other external forces are present).
+However, this approach has great difficulty when $q_d$ changes, since the
+external force due to gravity changes depending on the position of the arm $q$.
+The controller ends up having to use the accumulated integral to ``relearn'' the
+correct amount of extra force needed every time $q_d$ changes.
+
+In some robotics applications, the solution to this problem is to mathematically
+analyze the geometry and mass of the system and compute exactly how much extra
+force is needed.  In this particular case, the answer is straight-forward,
+in that the extra torque due to gravity is $\tau=m l g sin(q) / 2$, where
+$m$ is the mass of the arm, $l$ is the length, and $g$ is $9.8m/s^2$.  If the
+force applied by the motor is linear in $u$, then we could simply compute this
+value and add it to our controller's output.  However, this assumes a perfectly
+even distribution of weight in the arm, ignores momentum and other forces, and 
+gets much more complex as more joints are added.  Furthermore, if this
+initial computation is slightly off, or if details of the system change,
+there is no way to adjust this compensation.
+
+Fortunately, there is an adaptive solution to this problem, and it is one that
+fits well with neuromorphic hardware.  In \cite{Slotine1987}, Slotine shows that if you
+express the influence of these other external forces as $\tau=Y(q) \omega$
+(where $Y(q)$ is a fixed set of functions of $q$, such as $sin(q)$, and \omega
+is a vector of weights, one for each function in $Y$), then you can learn
+to compensate for these external forces by using the learning rule $\Delta \omega = \alpha Y(q) u$,
+where $u$ is the basic PD control signal.
+
+Importantly, as pointed out in [ref] and [ref], rather than making explicit
+assumptions about the exact functions that should be in $Y(q)$, we can use
+a neural network approach where each neuron is a different function of $q$.  
+As long as there is enough hetereogenetity (i.e. as long as the neural activity
+forms a basis space that spans the desired space of functions), then the
+learning rule will continue to work.
+
+This then suggests an explicit neuromorphic benchmark.  The input to the
+neuromorphic hardware is $q$, the system state.  This input is fed to each
+neuron such that each neuron produces some output behaviour that is based on
+this input.  Since $q$ will be multi-dimensional (if there is more than one
+joint), we may give each neuron a random weighting of each $q$ value ($J_i=e_i \cdot q$, 
+where $J_i$ is the input to neuron $i$, and $e_i$ is a randomly chosen vector\footnote{$e$ could also be chosen so as to regularly span the space of possibilities}).
+Given this input, the neurons will produce some output $A$.
+we now form a weighted sum of these outputs $Ad$, where $d$ is a matrix (number of neurons by number of elements in $q$)
+that is initially all zeros.
+
+To use this controller, we add its output to that of the standard PD controller.
+That is, the standard controller has $u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$,
+and so our actual output to the motor is $u + Ad$.  We then apply a learning
+rule on $d$ such that $\Delta d = \alpha A \times u$.
+
+Notice that we can think of this system as a three-layer neural
+network, where the input and output layers are linear.  The first layer is $q$, the input state, one value for each joint.  The ``hidden'' layer
+is $A$, the activity of a large number of neurons.  The output layer again has
+one value per joint, and is the extra added signal to apply to the motors, $Ad$.
+Given that this is such a canonical example of the use of neural networks, we
+hope that the majority of neuromorphic hardware is flexible enough to implement
+exactly this model.
+
+\subsection{Online versus offline learning}
+
+The one major step here that does not exist in a lot of neuromorphic hardware is
+the ability to update the weights $d$.  For hardware that does have a built-in
+learning rule, this rule is at least of a very common form, where the weight
+update from a neuron is proportional to the activity of that neuron and an
+external error signal.  This makes it an instance of the ubiquitous delta rule.
+
+However, if the neuromorphic hardware being benchmarked does not have the
+ability to update weights online using a learning rule of this form, then there
+are two solutions.  First, the multiplication by $d$ could be done on the
+output from the neuromorphic hardware.  There has to be some system to take
+the neural output from the hardware and send it to the motor (or the simulation
+of the motor).  Instead of outputing the result of $Ad$, the hardware could
+output $A$ (the activity of all the neurons), and the interface to the motor
+could be responsible for doing the multiplication by $d$ and updating $d$
+according to the learning rule.
+
+Alternatively, we can use offline learning.  That is, rather than updating
+the weights $d$ all the time, we simply record $A$ and $u$, and then after
+a period of time stop the controller, compute the sum total of the changes
+to $d$, load the new value of $d$ onto the neuromorphic hardware, and then
+start the controller again.
 

### 2015-08-10 paper/paper.tex
online
min sim for adapt control

@@
     hardware excels.  Furthermore, the benchmarks make use of a hybrid of
-    real physical embodiment and a type of "minimal" simulation that has been
+    real physical embodiment and a type of ``minimal'' simulation that has been
     shown to lead to robust real-world performance, allowing the same benchmark
@@
 often much easier to control and better-behaved than the real thing.  The
-field of robotics is filled with algorithms that work well ``in theory'', but
+field of robotics is filled with algorithms that work well ``in theory,'' but
 fail when run on actual hardware.  We do not want a benchmark that falls into
@@
 we believe there is an alternate approach known as \emph{Minimal Simulation}
-\cite{Jakobi97evolutionaryrobotics}.
+(\citealt{Jakobi97evolutionaryrobotics}).
 
@@
 leading to the standard PD controller.  Both $K_p$ and $K_d$ are constants that can be
-hand-tuned to particular situations.
+tuned to particular situations.
 
@@
 gravity.  Since the PD controller always produces an output $u=0$ when $q=q_d$,
-it cannot compensate for this. [TODO: add diagram]
+it cannot compensate for this, and so the arm will stay stationary at some
+angle below the desired angle. [TODO: add diagram]
 
@@
 force is needed.  In this particular case, the answer is straight-forward,
-in that the extra torque due to gravity is $\tau=m l g sin(q) / 2$, where
+in that the extra torque due to gravity is $\tau=mg {l \over 2} sin(q)$, where
 $m$ is the mass of the arm, $l$ is the length, and $g$ is $9.8m/s^2$.  If the
@@
 Fortunately, there is an adaptive solution to this problem, and it is one that
-fits well with neuromorphic hardware.  In \cite{Slotine1987}, Slotine shows that if you
+fits well with neuromorphic hardware.  \cite{Slotine1987} shows that if you
 express the influence of these other external forces as $\tau=Y(q) \omega$
-(where $Y(q)$ is a fixed set of functions of $q$, such as $sin(q)$, and \omega
+(where $Y(q)$ is a fixed set of functions of $q$, such as $sin(q)$, and $\omega$
 is a vector of weights, one for each function in $Y$), then you can learn
@@
 
-\subsection{Online versus offline learning}
+\subsection{Online and offline learning}
 
@@
 update from a neuron is proportional to the activity of that neuron and an
-external error signal.  This makes it an instance of the ubiquitous delta rule.
+external error signal.  This makes it an instance of the ubiquitous delta rule,
+and hopefully supported by the hardware.
 
@@
 
-% For Original Research Articles, Clinical Trial Articles, and Technology Reports the introduction should be succinct, with no subheadings.
-%
-% For Clinical Case Studies the Introduction should include symptoms at presentation, physical exams and lab results.
-%
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text.
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text.
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text.Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text.
+\subsection{Minimal Simulation for Adaptive Control}
 
-%\begin{methods}
-\section{Material \& Methods}
+Now that we have defined the task, we can use the principles of Minimal
+Simulation to construct a flexible and variable simulated environment for
+testing adaptive control.  The idea is to make a bare-bones simulation of
+the system being controlled, with built-in variability.  If the neuromorphic
+controller works well across this variability, then it is likely to work well
+outside of simulation as well.
 
-Text Text Text Text Text Text  Text Text Text Text Text Text \cite{conference} Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.
-\cite{patent} might want to know about  text text text text Text Text Text Text  Text Text Text Text Text Text  Text Text.
-\citep{article} might want to know about  text text text text
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text  Text Text.
-\cite{book} might want to know about  text text text text \cite{chapter}
+The basic system variable is a vector of length $N$ holding the joint angles $q$.  
+Each joint has a velocity $v$.  The force applied by each motor is 
+related to the signal sent to the motor $u$, but will generally have some
+maximum value $T$, so we use $tanh(u)T$. For friction, we 
+simply scale the velocity by some factor $F$ every time step.  
+This results in the simplistic simulation as follows:
 
-\begin{table}[!t]
-\textbf{\refstepcounter{table}\label{Tab:01} Table \arabic{table}.}{ Maximum size of the Manuscript }
+\begin{align}
+    \Delta v & =  -v F + tanh(u)T \\
+    \Delta q & = v
+\end{align}
 
-\processtable{ }
-{\begin{tabular}{lllll}\toprule
- & Abstract max. legth (incl. spaces) & Figures or tables & Manuscript max. length \\\midrule
-Clinical Case Study & & & &\\
-Clinical Trial & & & &\\
-Hypothesis and Theory & & & &\\
-Methods & 2000 characters  & 15 & 12000 words \\
-Original Research & & & &\\
-Review & & & &\\
-Technology Report & & & &\\\midrule
-Focused Review & 2000 characters & 5 & 5000 words \\\midrule
-CPC &  1250 characters& 6 & 2500 words  \\\midrule
-Perspective & 1250 characters & 2 & 3000 words  \\
-Mini Review & & & &\\\midrule
-Data Report & None & 2 & 3000 words\\\midrule
-Classification & 1250 characters & 10 & 2000 words \\\midrule
-Editorial & None & None & 1000 words  \\\midrule
-Frontiers Commentary  & & &\\
-General Commentary & None & 1 & 1000 words\\
-Book review & & & \\\midrule
-Opinion   & & &\\
-Specialty Grand Challenge & None & 1 & 2000 words\\
-Field Grand Challenge & & & &\\\botrule
-\end{tabular}}{}
-\end{table}
+On top of this, we need to add the external force for which the adaptive system
+is trying to compensate.  Rather than deciding on what specific external
+force, we \emph{randomly generate} an equation controlling this force.  We start with a small set of
+smooth functions on $X$ which are often found in dynamics equations ($x$, $x^2$,
+$sin(x)$).  We then generate an external force of $Q (\zeta X(\beta q + \gamma) + \eta)$
+where $\zeta$, $\beta$, $\gamma$, and $\eta$ are all randomly chosen fron $N(0,1)$
+and $Q$ is a scaling factor to control how strong this external force is.  The
+result is added to Equation 1.
 
-Please note that large tables covering several pages cannot be included in the final PDF for formatting reasons.
-These tables will be published as supplementary material on the online article abstract page at the time of acceptance.
-The author will notified during the typesetting of the final article if this is the case.
-A link in the final PDF will direct to the online material.
+Finally, we add random noise, delay, and filtering to both the input and the
+output of the system.  For noise, we add $N(0,\sigma_u)$ to the control signal
+$u$ and $N(0,\sigma_q)$ to the $q$ value reported back to the controller.  We
+also use a low-pass filter to smooth both values (with time constants $\tau_u$
+and $\tau_q$) after this noise is added.  Finally, both $q$ and $u$ are delayed
+by an amount of time $t_q$ and $t_u$.
 
-\subsection{Original Research Articles, Clinical Trial Articles, and Technology Reports}
+The result is not meant to be a simulation of a particular physical embodiment.
+Rather, the variability in this simulation is meant to be extremely fast
+to simulation, and to make it hard to ``cheat''
+to control it.  In other words, if a controller manages to be able to control
+the various randomly created minimal simulations of embodiment that are
+generated with this approach, then we have reason to believe that it will also
+be successful at controlling real embodiments.  With this simulation and
+modern computers, we can easily simulate in real-time systems with dozens of joints and
+highly complex interactions between them, and measure how well an adaptive
+controller deals with these situations.
 
-For Original Research Articles, Clinical Trial Articles, and Technology Reports the section headings should be those appropriate for your field and the research itself.
-It is recommended to organize your manuscript in the following sections or their equivalents for your field:
-
-\begin{itemize}
-%for bulleted list, use itemize
-\item Introduction: Succinct, with no subheadings.
-\item Materials and Methods: This section may be divided by subheadings.
-This section should contain sufficient detail so that when read in conjunction with cited references, all procedures can be repeated.
-\item Results: This section may be divided by subheadings. Footnotes should not be used and have to be transferred into the main text.
-\item Discussion: This section may be divided by subheadings.
-Discussions should cover the key findings of the study: discuss any prior art related to the subject so to place the novelty of the discovery in the appropriate context; discuss the potential short-comings and limitations on their interpretations; discuss their integration into the current understanding of the problem and how this advances the current views; speculate on the future direction of the research and freely postulate theories that could be tested in the future.
-\end{itemize}
-
-Please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
-
-\subsection{Clinical Case Studies}
-
-For Clinical Case Studies the following sections are mandatory:
-
-\begin{itemize}
-%for bulleted list, use itemize
-\item Introduction: Include symptoms at presentation, physical exams and lab results.
-\item Background: This section may be divided by subheadings. Include history and review of similar cases.
-\item Results: This section may be divided by subheadings. Include diagnosis and treatment.
-\item Concluding Remarks
-\end{itemize}
-
-%\end{methods}
-
-\section{Results}
-
-\subsection{Figures}
-Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript.
-Figures will then be automatically embedded at the bottom of the submitted manuscript.
-Kindly ensure that each table and figure is mentioned in the text and in numerical order.
-Permission must be obtained for use of copyrighted material from other sources (including the web).
-Please note that it is compulsory to follow figure instructions.
-Figures which are not according to the guidelines will cause substantial delay during the production process.
-
-\begin{table}[!t]
-\textbf{\refstepcounter{table}\label{Tab:02} Table \arabic{table}.}{ Resolution Requirements for the figures}
-
-\processtable{}
-{\begin{tabular}{lllll}\toprule
-Image Type & Description & Format & Color Mode & Resolution\\\midrule
-Line Art & An image composed of lines and text,  & TIFF, JPEG & RGB, Bitmap & 900 - 1200 dpi\\
-           & which does not contain tonal or shaded areas.& & &\\
-           Halftone & A continuous tone photograph, which contains no text. & TIFF, EPS, JPEG & RGB, Grayscale & 300 dpi\\
-Combination & Image contains halftone + text or line art elements. & TIFF, JPEG & RGB,Grayscale & 600 - 900 dpi\\\botrule
-\end{tabular}}{}
-\end{table}
-
-\textbf{Table \ref{Tab:02}} shows the resolution requirements for the figures. The figures must be legible:
-\begin{enumerate}
-\item The smallest visible text is no less than 8 points in height, when viewed at actual size.
-\item Solid lines are not broken up.
-\item Image areas are not pixelated or stair stepped.
-\item Text is legible and of high quality.
-\item Any lines in the graphic are no smaller than 2 points width.
-\item The actual size of the figure must be of at least 8.5 cm.
-\end{enumerate}
-
-\subsection{Nomenclature}
-\begin{itemize}
-\item The use of abbreviations should be kept to a minimum.
-Non-standard abbreviations should be avoided unless they appear at least four times, and defined upon first use in the main text.
-Consider also giving a list of non-standard abbreviations at the end, immediately before the Acknowledgments.
-\item Gene symbols should be italicized; protein products are not italicized.
-\item Chemical compounds and biomolecules should be referred to using systematic nomenclature, preferably using the recommendations by IUPAC.
-\item We encourage the use of Standard International Units in all manuscripts.
-\item To take part in the Resource Identification Initiative, please cite antibodies, genetically modified organisms, software tools, data, databases and services using the corresponding catalog number and RRID in your current manuscript.
-For more information about the project and for steps on how to search for an RRID, please click \href{http://www.frontiersin.org/files/pdf/letter_to_author.pdf}{here}.
-\end{itemize}
-
-\begin{equation}
-\sum x+ y =Z\label{eq:01}
-\end{equation}
-
-\section{Discussion}
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text Text.
-Additional Requirements:
-\subsection{Corrections}
-
-If you need to communicate important changes to a published article please submit a General Commentary. Submit the article with the title “Corrigendum: Original Title of Article”.
-
-\subsection{Commentaries on Articles}
-
-At the beginning of your Commentary, please provide the citation of the article commented on.
-Rebuttals may be submitted in response to Commentaries; our limit in place is one commentary and one response.
-Rebuttals should also be submitted as General Commentary articles.
-
-\subsection{Human Search and Animal Research}
-
-All experiments on live vertebrates or higher invertebrates must be performed in accordance with relevant institutional and national guidelines and regulations.
-In the manuscript, authors must identify the committee approving the experiments and must confirm that all experiments conform to the relevant regulatory standards.
-For manuscripts reporting experiments on human subjects, authors must identify the committee approving the experiments and must also include a statement confirming that informed consent was obtained from all subjects.
-In Original Research Articles and Clinical Trial Articles these statements should appear in the Materials and Methods section.
-
-\subsection{Clinical Trial Registration}
-
-Clinical trials should be registered in a public trials registry in order to become the object of a publication at Frontiers.
-Trials must be registered at or before the start of patient enrollment.
-A clinical trial is defined as "any research study that prospectively assigns human participants or groups of humans to one or more health-related interventions to evaluate the effects on health outcomes."(\href{www.who.int/ictrp/en}{www.who.int/ictrp/en}).
-A list of acceptable registries can be found at \href{www.who.int/ictrp/en}{www.who.int/ictrp/en} and \href{www.icmje.org}{www.icmje.org}.
-
-\subsection{Inclusion of Proteomics Data}
-
-Authors should provide relevant information relating to how the peptide/protein matches were undertaken, including methods used to process and analyze data, false discovery rates (FDR) for large-scale studies and threshold or cut-off rates for peptide and protein matches.
-Further information could include software used, mass spectrometer type, sequence database and version, number of sequences in database, processing methods, mass tolerances used for matching, variable/fixed modifications, allowable missed cleavages, etc.
-
-Authors should provide as supplementary material information used to identify proteins and/or peptides.
-This should include information such as accession numbers, observed mass (m/z), charge, delta mass, matched mass, peptide/protein scores, peptide modification, miscleavages, peptide sequence, match rank, matched species (for cross species matching), number of peptide matches, ambiguous protein/peptide matches should be indicated, etc.
-For quantitative proteomics analyses authors should provide information to justify the statistical significance including biological replicates, statistical methods, estimates of uncertainty and the methods used for calculating error.
-
-For peptide matches with biologically relevant post-translational modifications (PTM) and for any protein match that has occurred using a single mass spectrum, authors should include this information as raw data, annotated spectra or submit data to an online repository (recommended option).
-Authors are encouraged to submit raw or matched data and 2-DE images to public proteomics repositories. Submission codes and/or links to data should be provided within the manuscript.
-
-\subsection{Data Sharing}
-
-Frontiers supports the policy of data sharing, and authors are advised to make freely available any materials and information described in their article, and any data relevant to the article (while not compromising confidentiality in the context of human-subject research) that may be reasonably requested by others for the purpose of academic and non-commercial research.
-In regards to deposition of data and data sharing through databases, Frontiers urges authors to comply with the current best practices within their discipline.
+\subsection{Cheap Robotics for Adaptive Control}
 
@@
 
-\section*{Author Contributions}
-%When determining authorship the following criteria should be observed:
-%•	Substantial contributions to the conception or design of the work; or the acquisition, analysis, or interpretation of data for the work; AND
-%•	Drafting the work or revising it critically for important intellectual content; AND
-%•	Final approval of the version to be published ; AND
-%•	Agreement to be accountable for all aspects of the work in ensuring that questions related to the accuracy or integrity of any part of the work are appropriately investigated and resolved.
-%Contributors who meet fewer than all 4 of the above criteria for authorship should not be listed as authors, but they should be acknowledged. (http://www.icmje.org/roles_a.html)
-
-The statement about the authors and contributors can be up to several sentences long, describing the tasks of individual authors referred to by their initials and should be included at the end of the manuscript before the References section.
-
 \section*{Acknowledgments}
@@
 \textit{Funding\textcolon} Text Text Text Text Text Text  Text Text.
-
-\section*{Supplemental Data}
-Supplementary Material should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure.
-LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder
-
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.
 

### 2015-08-10 paper/paper.tex
robot adapt control

@@
 
+Of course, we need physical confirmation that the minimal simulation defined
+in the previous section does a reasonable job.  For this, we define a
+particular easy-to-build example of a system that can be usefully controlled
+by this adaptive method.  For this, we use the Lego Mindstorms EV3 robot kit.
+
+The physical body is depicted in [TODO: add figure].  It consists of a single
+motor, mounted in such a way that there is a significant amount of weight on
+the arm itself (the weight of the motor itself).  Multiple motors can be
+added, and other configurations can be considered and should also be suitable
+for testing control, but here we just consider the basic case.
+
+To interface to the physical hardware, we installed the \texttt{ev3dev} operating
+system (\texttt{http://ev3dev.org}), a Debian-based Linux system specifically developed
+for the EV3.  We then installed and ran the \texttt{ev3\_link} program from \texttt{ev3dev-c}
+(\texttt{https://github.com/in4lio/ev3dev-c}).  This allows the EV3 to listen for
+UDP commands that tell it to set motor values and read sensor values.
+Communication with a PC was over a USB link (although the system also
+supports WiFi communication).  With constant communication, the system is
+able to adjust the power sent to the motors $u$ and give position feedback
+$q$ from those motors at a rate of around 300Hz.
+
+Figure [???] shows the system being controlled by a simple PD controller.
+The desired joint angle $q_d$ is varied randomly over time.  Notice that the
+system consistently undershoots and overshoots the desired angle, due to
+its inability to compensate for gravity.
+
+In Figure [???], we show the result of using the adaptive controller algorithm.
+The system quickly improves on the standard PD controller, learning to adjust
+its control signal $u$ to compensate for the unknown external forces.
+
+\section{Benchmark Analysis}
+
+- what do we measure?   performance: rmse and delay.  Also, power consumption?  cost?
+
+- for given hardware, how many neurons can be done in realtime?  that's how many should be used for comparison
+
+- show how these things change as we change parameters in the simulation
+
+- delays
+
+- number of motors $N$
+
+- magnitude of nonlinearity $Q$
+
+CPU, GPU, and SpiNNaker?
+
+\section{Other Benchmarks}
+
+(sketch out how this approach could be applied to other task)
+
+Adaptive Jacobian
+
+Driving down a corridor, avoiding obstacles
+
+Classical conditioning
+
+Operant conditioning
+
 \section*{Disclosure/Conflict-of-Interest Statement}
-%Frontiers follows the recommendations by the International Committee of Medical Journal Editors (http://www.icmje.org/ethical_4conflicts.html) which require that all financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed.
-If no such relationship exists, authors will be asked to declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
-When disclosing the potential conflict of interest, the authors need to address the following points:
-%•	Did you or your institution at any time receive payment or services from a third party for any aspect of the submitted work?
-%•	Please declare financial relationships with entities that could be perceived to influence, or that give the appearance of potentially influencing, what you wrote in the submitted work.
-%•	Please declare patents and copyrights, whether pending, issued, licensed and/or receiving royalties relevant to the work.
-%•	Please state other relationships or activities that readers could perceive to have influenced, or that give the appearance of potentially influencing, what you wrote in the submitted work.
 

### 2015-08-10 paper/paper.tex
Added refs

@@
 robots to perform various tasks.  These tasks can be as simple as navigation
-and obstacle avoidance, but have also included complex eight-legged walking,
+and obstacle avoidance, but have also included walking,
 object identification, and visual tracking [refs???].
@@
 All we need is for the simulation to be just good enough to indicate that
-things have gone wrong, and thus give a low score to that controller.
+things have gone wrong, and thus give a low score to that controller.  This
+means that, for example, in a minimal simulation of an eight-legged walking robot,
+it is not necessary to have a physics simulation that correctly models what
+happens when two legs collide with each other.  Rather, if legs collide
+with each other, that is an indication that the walking behaviour is very poor,
+so as long as that result is indicated we can greatly simplify the simulation
+by not including all the details necessary to deal with these sorts of physical
+interactions.
 
@@
 is not to be used as the primary benchmark.  Rather, this is meant as a
-double-check that the system does actually behave as expected.  The worry
-is always that if researchers focus on the simulated system when benchmarking
-their hardware, they may inadvertantly end up with hardware that only performs
-well in simulation.
+double-check that the system does actually behave as expected.
 
@@
 The simplest controller for such a situation is a P (proportional) controller,
-where $u=K_p(q_d - q)$.  This is often supplemented with a derivative term,
+where $u=K_p(q_d - q)$.  This is often supplemented with a D (derivative) term,
 which helps to slow the system down as it approaches the desired position,
@@
 
-The standard solution to this problem is to add an integral term ($K_i \int{q_d-q dt}$)
+The standard solution to this problem is to add an I (integral) term ($K_i \int{(q_d-q) dt}$)
 to the controller, making it a PID controller.  The idea here is that as the difference between where it is and
-where we want it to be accumulates over time, the $K-i$ term gradually increases
+where we want it to be accumulates over time, the $K_i$ term gradually increases
 how much extra force is being applied until it is large enough to counteract
@@
 (where $Y(q)$ is a fixed set of functions of $q$, such as $sin(q)$, and $\omega$
-is a vector of weights, one for each function in $Y$), then you can learn
+is a vector of scalar weights, one for each function in $Y$), then you can learn
 to compensate for these external forces by using the learning rule $\Delta \omega = \alpha Y(q) u$,
@@
 
-Importantly, as pointed out in [ref] and [ref], rather than making explicit
+Importantly, as pointed out by \cite{SannerSlotine1992} and \cite{Lewis1996}, rather than making explicit
 assumptions about the exact functions that should be in $Y(q)$, we can use
@@
 forms a basis space that spans the desired space of functions), then the
-learning rule will continue to work.
+learning rule will continue to work.  This approach has been used in the
+Recurrent Error-driven Adaptive Control Hierarchy model of human motor control
+(\citealt{dewolf2014}), and for quadcopter control [TODO: any citations for this?].
 

### 2015-08-10 paper/paper.tex
typos/formatting

@@
 reflective of the actual usefulness of the neuromorphic hardware?  It is
-very widely known that \emph{simulations of robots} (or other physical systems) are
-often much easier to control and better-behaved than the real thing.  The
+very widely known that simulations of robots (or other physical systems) are
+often \emph{much} easier to control and better-behaved than the real thing.  The
 field of robotics is filled with algorithms that work well ``in theory,'' but
@@
 
-To address this problem, Nicholas Jacobi proposed the creation of
+To address this problem, \cite{Jakobi97evolutionaryrobotics} proposed the creation of
 ``minimal'' simulations.  These are simulations where there is variability
@@
 
-It is also worth noting that a minimal simulation \emph{only has to be a
-good simulation for successful behaviour}.  That is, ``if we are evolving 
+It is also worth noting that a minimal simulation only has to be a
+good simulation \emph{for successful behaviour}.  That is, ``if we are evolving 
 corridor following behaviour, the dynamics of the simulation might
@@
 advantages.  In particular, we \emph{don't want benchmarks that rely on
-particular high-speed, high-accuracy devices}.  The purpose of the benchmark
+high-speed, high-accuracy devices}.  The purpose of the benchmark
 is not to indicate how well this neuromorphic hardware works to control this
@@
 output from the neuromorphic hardware.  There has to be some system to take
-the neural output from the hardware and send it to the motor (or the simulation
+the neural output from the hardware and send it to the motor (or to the simulation
 of the motor).  Instead of outputing the result of $Ad$, the hardware could
@@
 Rather, the variability in this simulation is meant to be extremely fast
-to simulation, and to make it hard to ``cheat''
+to simulate, and to make it hard to ``cheat''
 to control it.  In other words, if a controller manages to be able to control
@@
 particular easy-to-build example of a system that can be usefully controlled
-by this adaptive method.  For this, we use the Lego Mindstorms EV3 robot kit.
+by this adaptive method.  In particular, we use the Lego Mindstorms EV3 robot kit.
 

### 2015-08-11 code/benchmark.py
Code for running benchmark on EV3

@@ -1,0 +1,151 @@
+import argparse
+import importlib
+import inspect
+import logging
+import os
+import shelve
+import time
+
+import matplotlib.pyplot
+import numpy as np
+
+import nengo
+
+def find_offset(a, b):
+    assert len(a) == len(b)
+    corr = np.correlate(a, b, 'full')
+    index = np.argmax(corr[len(a):])
+    return index
+
+
+class Benchmark(object):
+    def __init__(self):
+        self.parser = argparse.ArgumentParser(
+                            description='Nengo benchmark: %s' %
+                                 self.__class__.__name__)
+        self.param_names = []
+        self.hidden_params = []
+        self.params()
+        self.fixed_params()
+
+    def default(self, description, **kwarg):
+        if len(kwarg) != 1:
+            raise ValueException('Must specify exactly one parameter')
+        k, v = kwarg.items()[0]
+        if k in self.param_names:
+            raise ValueException('Cannot redefine parameter "%s"' % k)
+        if v is False:
+            self.parser.add_argument('--%s' % k, action='store_true',
+                                     help=description)
+        else:
+            self.parser.add_argument('--%s' % k, type=type(v), default=v,
+                                     help=description)
+        self.param_names.append(k)
+
+    def fixed_params(self):
+        self.default('backend to use', backend='nengo')
+        self.default('time step', dt=0.001)
+        self.default('random number seed', seed=1)
+        self.default('data directory', data_dir='data')
+        self.default('display figures', show_figs=False)
+        self.default('do not generate figures', no_figs=False)
+        self.default('enable debug messages', debug=False)
+        self.hidden_params.extend(['data_dir', 'show_figs',
+                                   'no_figs', 'debug'])
+
+    def process_args(self, allow_cmdline=True, **kwargs):
+        if len(kwargs) == 0 and allow_cmdline:
+            args = self.parser.parse_args()
+        else:
+            args = argparse.Namespace()
+            for k in self.param_names:
+                v = kwargs.get(k, self.parser.get_default(k))
+                setattr(args, k, v)
+
+        name = self.__class__.__name__
+        text = []
+        for k in self.param_names:
+            if k not in self.hidden_params:
+                text.append('%s=%s' % (k, getattr(args, k)))
+
+        filename = name + '#' + ','.join(text)
+
+        return args, filename
+
[…]

### 2015-08-11 code/benchmark_ev3.py

@@ -1,0 +1,125 @@
+import time
+
+import nengo
+import numpy as np
+import ev3link
+import embodied_benchmarks as bench
+
+import benchmark
+
+class SingleArmEV3(benchmark.Benchmark):
+    def params(self):
+        self.default('Kp', Kp=2.0)
+        self.default('Kd', Kd=1.0)
+        self.default('Ki', Ki=0.0)
+        self.default('tau_d', tau_d=0.001)
+        self.default('T', T=0.5)
+        self.default('IP address of EV3', address='10.42.0.3')
+        self.default('motor index', motor_index=0)
+        self.default('desired path', path='step')
+        self.default('period', period=4.0)
+        self.default('use adaptation', adapt=False)
+        self.default('n_neurons', n_neurons=500)
+        self.default('learning rate', learning_rate=1.0)
+        self.default('max_freq', max_freq=1.0)
+
+    def model(self, p):
+        link = ev3link.EV3Link(p.address)
+        self.link = link
+        path = '/sys/class/tacho-motor/motor%d/' % p.motor_index
+        self.path = path
+        link.write(path + 'command', 'stop')
+        link.write(path + 'position', '0\n')
+        link.write(path + 'command', 'run-direct')
+        print 'current position: ', link.read(path + 'position')
+    
+        model = nengo.Network()
+        with model:
+            
+            def ev3_system(t, x):
+                value = int(100 * x[0])
+                if value > 100:
+                    value = 100
+                if value < -100:
+                    value = -100
+                value = '%d' % value
+                link.write(path + 'duty_cycle_sp', value)
+                p = link.read(path + 'position')
+                try:
+                    return float(p) / 180 * np.pi
+                except:
+                    return 0
+            
+            ev3 = nengo.Node(ev3_system, size_in=1, size_out=1)
+            
+            if p.adapt:
+                pid = bench.pid.PDAdapt(1, p.Kp, p.Kd, tau_d=p.tau_d, learning_rate=p.learning_rate,
+                                        n_neurons=p.n_neurons)
+            else:
+                pid = bench.pid.PID(p.Kp, p.Kd, p.Ki, tau_d=p.tau_d)
+            control = nengo.Node(lambda t, x: pid.step(x[:1], x[1:]), size_in=2)
+            nengo.Connection(ev3, control[:1], synapse=0)
[…]

### 2015-08-11 code/control_ev3.py

@@ -1,0 +1,65 @@
+import time
+
+import nengo
+import numpy as np
+import ev3link
+import embodied_benchmarks as bench
+
+link = ev3link.EV3Link('10.42.0.3')
+
+path0 = '/sys/class/tacho-motor/motor0/'
+link.write(path0 + 'command', 'run-direct')
+link.write(path0 + 'position', '0')
+print 'current position: ', link.read(path0 + 'position')
+    
+start_time = time.time()
+    
+model = nengo.Network()
+with model:
+    
+    def ev3_system(t, x):
+        value = int(100 * x[0])
+        if value > 100:
+            value = 100
+        if value < -100:
+            value = -100
+        value = '%d' % value
+        link.write(path0 + 'duty_cycle_sp', value)
+        p = link.read(path0 + 'position')
+        try:
+            return float(p) / 180 * np.pi
+        except:
+            return 0
+    
+    ev3 = nengo.Node(ev3_system, size_in=1, size_out=1)
+    
+    pid = bench.pid.PID(2,1,0, tau_d=0.001)
+    control = nengo.Node(lambda t, x: pid.step(x[:1], x[1:]), size_in=2)
+    nengo.Connection(ev3, control[:1], synapse=0)
+    nengo.Connection(control, ev3, synapse=None)
+    
+    actual_time = nengo.Node(lambda t: time.time() - start_time)
+
+    def desired_func(t, actual_t):
+        return np.sin(actual_t*2*np.pi)
+    desired = nengo.Node(desired_func, size_in=1)
+    nengo.Connection(actual_time, desired, synapse=None)
+    nengo.Connection(desired, control[1:], synapse=None)
+
+    p_desired = nengo.Probe(desired, synapse=None)
+    p_q = nengo.Probe(ev3, synapse=None)
+    p_t = nengo.Probe(actual_time, synapse=None)
+
+
+
+sim = nengo.Simulator(model)
+start_time = time.time()
+sim.run(0.1)
+    
+link.write(path0 + 'duty_cycle_sp', '0')
+
+import pylab
+pylab.plot(sim.data[p_t], sim.data[p_desired])
+pylab.plot(sim.data[p_t], sim.data[p_q])
+pylab.show()
+    

### 2015-08-11 code/plot.py

@@ -1,0 +1,27 @@
+import shelve
+import sys
+
+import numpy as np
+import pylab
+
+db = shelve.open(sys.argv[1])
+
+q = db['p_q'][:,0]
+d = db['p_desired'][:,0]
+t = db['p_t']
+
+def find_offset(a, b):
+    assert len(a) == len(b)
+    corr = np.correlate(a, b, 'full')
+    index = np.argmax(corr[len(a):])
+    return index
+
+offset = find_offset(q, d)
+
+dt = np.mean(t[offset:]-t[:-offset])
+print dt
+
+#pylab.plot(t, d)
+pylab.plot(t[offset:], d[:-offset])
+pylab.plot(t[offset:], q[offset:])
+pylab.show()

### 2015-08-11 code/benchmark_ev3.py
Fixed adaptation

@@ -22,6 +22,8 @@ class SingleArmEV3(benchmark.Benchmark):
         self.default('n_neurons', n_neurons=500)
         self.default('learning rate', learning_rate=1.0)
         self.default('max_freq', max_freq=1.0)
+        self.default('synapse', synapse=0.01)
+        self.default('radius', radius=1.0)
 
     def model(self, p):
         link = ev3link.EV3Link(p.address)
@@ -52,14 +54,23 @@ class SingleArmEV3(benchmark.Benchmark):
             
             ev3 = nengo.Node(ev3_system, size_in=1, size_out=1)
             
-            if p.adapt:
-                pid = bench.pid.PDAdapt(1, p.Kp, p.Kd, tau_d=p.tau_d, learning_rate=p.learning_rate,
-                                        n_neurons=p.n_neurons)
-            else:
-                pid = bench.pid.PID(p.Kp, p.Kd, p.Ki, tau_d=p.tau_d)
+            #if p.adapt:
+            #    pid = bench.pid.PDAdapt(1, p.Kp, p.Kd, tau_d=p.tau_d, learning_rate=p.learning_rate,
+            #                            n_neurons=p.n_neurons)
+            pid = bench.pid.PID(p.Kp, p.Kd, p.Ki, tau_d=p.tau_d)
             control = nengo.Node(lambda t, x: pid.step(x[:1], x[1:]), size_in=2)
             nengo.Connection(ev3, control[:1], synapse=0)
             nengo.Connection(control, ev3, synapse=None)
+
+            if p.adapt:
+                adapt = nengo.Ensemble(p.n_neurons, dimensions=1, radius=p.radius)
+                nengo.Connection(ev3, adapt, synapse=None)
+                conn = nengo.Connection(adapt, ev3, synapse=p.synapse,
+                        function=lambda x: 0,
+                        learning_rule_type=nengo.PES())
+                conn.learning_rule_type.learning_rate *= p.learning_rate
+                nengo.Connection(control, conn.learning_rule, synapse=None,
+                        transform=-1)
             
             actual_time = nengo.Node(lambda t: time.time() - self.start_time,
                     size_in=0, size_out=1)

### 2015-08-11 code/plots/fig_ev3.py
Added plot for ev3 control

@@ -1,0 +1,48 @@
+import pylab
+import shelve
+import numpy as np
+
+pylab.figure(figsize=(10,6))
+
+index = 0
+for path in ['step', 'random']:
+    for adapt in ['False', 'True']:
+
+        db = shelve.open('SingleArmEV3#path=%s,adapt=%s.db' % (path, adapt))
+
+        index += 1
+        pylab.subplot(2, 2, index)
+
+        t = db['p_t'][:,0]
+        q = db['p_q'][:,0]
+        d = db['p_desired'][:,0]
+
+        if path == 'random':
+            q *= 2
+            d *= 2
+
+        pylab.plot(t, q, color='#888888', linewidth=2, label='$q$')
+        pylab.plot(t, d, color='k', linewidth=1, label='$q_d$')
+        pylab.ylim(-2, 2)
+        pylab.yticks([-np.pi/2, 0, np.pi/2], ['$-\pi$', '0','$\pi$'])
+
+        if adapt == 'False':
+            pylab.ylabel('joint angle (radians)')
+        if path == 'random':
+            pylab.xlabel('time (s)')
+        if path == 'step' and adapt=='False':
+            pylab.title('without adaptation')
+        if path == 'step' and adapt=='True':
+            pylab.title('with adaptation')
+
+        pylab.legend(loc='upper left')
+
+
+
+
+        db.close()
+
+pylab.savefig('fig_ev3.png', dpi=600)
+pylab.show()
+
+

### 2015-08-12 code/benchmark.py
Made minsim benchmark

@@ -64,11 +64,15 @@ class Benchmark(object):
 
         name = self.__class__.__name__
         text = []
+        self.args_text = []
         for k in self.param_names:
             if k not in self.hidden_params:
                 text.append('%s=%s' % (k, getattr(args, k)))
+                self.args_text.append('_%s = %r' % (k, getattr(args, k)))
 
         filename = name + '#' + ','.join(text)
+
+        filename = name + '#' + time.strftime('%H%M%S')
 
         return args, filename
 
@@ -122,6 +126,7 @@ class Benchmark(object):
         text = []
         for k, v in sorted(result.items()):
             text.append('%s = %s' % (k, repr(v)))
+        text = self.args_text + text
         text = '\n'.join(text)
 
 
@@ -148,4 +153,6 @@ class Benchmark(object):
 
         if p.show_figs:
             plt.show()
+
+        return result
 

### 2015-08-12 code/benchmark_min.py

@@ -1,0 +1,117 @@
+import time
+
+import nengo
+import numpy as np
+import embodied_benchmarks as bench
+
+import benchmark
+
+class AdaptiveControl(benchmark.Benchmark):
+    def params(self):
+        self.default('Kp', Kp=2.0)
+        self.default('Kd', Kd=1.0)
+        self.default('Ki', Ki=0.0)
+        self.default('tau_d', tau_d=0.001)
+        self.default('T', T=0.5)
+        self.default('desired path', path='random')
+        self.default('period', period=4.0)
+        self.default('use adaptation', adapt=False)
+        self.default('n_neurons', n_neurons=500)
+        self.default('learning rate', learning_rate=1.0)
+        self.default('max_freq', max_freq=1.0)
+        self.default('synapse', synapse=0.01)
+        self.default('radius', radius=1.0)
+        self.default('number of dimensions', D=1)
+        self.default('scale_add', scale_add=1)
+        self.default('noise', noise=0.1)
+        self.default('filter', filter=0.01)
+        self.default('delay', delay=0.01)
+
+    def model(self, p):
+
+        model = nengo.Network()
+        with model:
+
+            system = bench.system.System(p.D, p.D, dt=p.dt, seed=p.seed, motor_noise=p.noise, sense_noise=p.noise, scale_add=p.scale_add,
+                    motor_scale=10,
+                    motor_delay=p.delay, sensor_delay=p.delay,
+                    motor_filter=p.filter, sensor_filter=p.filter)
+
+            def minsim_system(t, x):
+                return system.step(x)
+
+            minsim = nengo.Node(minsim_system, size_in=p.D, size_out=p.D)
+
+            pid = bench.pid.PID(p.Kp, p.Kd, p.Ki, tau_d=p.tau_d)
+            control = nengo.Node(lambda t, x: pid.step(x[:p.D], x[p.D:]), size_in=p.D*2)
+            nengo.Connection(minsim, control[:p.D], synapse=0)
+            nengo.Connection(control, minsim, synapse=None)
+
+            if p.adapt:
+                adapt = nengo.Ensemble(p.n_neurons, dimensions=p.D, radius=p.radius)
+                nengo.Connection(minsim, adapt, synapse=None)
+                conn = nengo.Connection(adapt, minsim, synapse=p.synapse,
+                        function=lambda x: 0,
+                        learning_rule_type=nengo.PES())
[…]

### 2015-08-12 code/plot_vary.py

@@ -1,0 +1,72 @@
+import os
+import sys
+
+import numpy as np
+import pylab
+
+path = sys.argv[1]
+
+data_x = []
+data_y = []
+for fn in os.listdir(path):
+    if fn.endswith('.txt'):
+        with open(os.path.join(path, fn)) as f:
+            text = f.read()
+        d = dict()
+        exec(text, d)
+
+        x = d['_delay']
+        y = d['rmse']
+
+        data_x.append(x)
+        data_y.append(y)
+
+#data_x = data_x[:30]
+#data_y = data_y[:30]
+
+do_gp = True
+
+pylab.figure(figsize=(8,4))
+
+if do_gp:
+    import pylab
+    import GPy
+
+    X = np.array(data_x)
+    X.shape = X.shape[0], 1
+    Y = np.array(data_y)
+    Y.shape = Y.shape[0], 1
+
+
+    kernel = GPy.kern.RBF(input_dim=1, variance=1, lengthscale=0.005)
+    #kernel = kernel+GPy.kern.RBF(input_dim=1, variance=1, lengthscale=0.005)
+    #kernel = GPy.kern.Linear(1)
+    #kernel = GPy.kern.Brownian(1)
+    m = GPy.models.GPRegression(X, Y,kernel)
+    #m = GPy.models.SparseGPRegression(X, Y,kernel)
+    #m = GPy.models.GPCoregionalizedRegression([X], [Y],kernel)
+    #m = GPy.models.GPHeteroscedasticRegression(X, Y,kernel)
+    #m = GPy.models.GPVariationalGaussianApproximation(X, Y, kernel)
+    #m = GPy.models.SparseGPRegression(X,Y,Z=np.random.rand(10,1)*0.04, kernel=kernel)
+    m.optimize('bfgs', messages=True)
+    #m.optimize('tnc', messages=True)
+
+    #print help(m.optimize)
+
+
+    #m.optimize('bfgs', max_iters=100)
+    #m.optimize_restarts(num_restarts=10)
+    m.optimize()
+
+    print m
+
+    m.plot()
+
+else:
+    pylab.scatter(data_x, data_y, marker='x', color='k')
+
+
+pylab.xlim(0, 0.04)
+pylab.xlabel('delay (s)')
+pylab.ylabel('rmse')
+pylab.savefig('plot_vary.png', dpi=100)

### 2015-08-12 code/vary_delay.py

@@ -1,0 +1,9 @@
+from benchmark_min import AdaptiveControl
+
+import numpy as np
+rng = np.random.RandomState()
+
+for i in range(10):
+    AdaptiveControl().run(no_figs=True, T=20, seed=i, max_freq=1,
+                scale_add=1, delay=rng.uniform(0, 0.04),
+                adapt=True, data_dir='vary_delay')

### 2015-08-12 code/plot_vary.py
simpler plots

@@ -4,6 +4,8 @@ import sys
 import numpy as np
 import pylab
 
+import bootstrapci
+
 path = sys.argv[1]
 
 data_x = []
@@ -24,12 +26,14 @@ for fn in os.listdir(path):
 #data_x = data_x[:30]
 #data_y = data_y[:30]
 
-do_gp = True
+data_x = np.array(data_x)
+data_y = np.array(data_y)
+
+do_gp = False
 
 pylab.figure(figsize=(8,4))
 
 if do_gp:
-    import pylab
     import GPy
 
     X = np.array(data_x)
@@ -63,7 +67,32 @@ if do_gp:
     m.plot()
 
 else:
-    pylab.scatter(data_x, data_y, marker='x', color='k')
+
+
+    '''
+    low = []
+    high = []
+    var = 0.0005
+    x = np.linspace(0, 0.04, 100)
+    for xx in x:
+        w = np.exp(-((data_x-xx)**2)/(2*var**2))
+        w /= sum(w)
+        samples = [np.random.choice(data_y, p=w) for i in range(200)]
+
+
+        #ci = bootstrapci.bootstrapci(samples, np.mean)
+        mean = np.mean(samples)
+        sd = np.std(samples)
+        ci = mean - 2*sd, mean+2*sd
+        low.append(ci[0])
+        high.append(ci[1])
+        print xx, ci
+
+    pylab.fill_between(x, low, high, color='#888888')
+    '''
+
+    pylab.scatter(data_x, data_y, s=30, marker='x', color='k')
+
 
 
 pylab.xlim(0, 0.04)

### 2015-08-12 code/benchmark.py
Fixes for multiple dimensions

@@ -126,13 +126,15 @@ class Benchmark(object):
         text = []
         for k, v in sorted(result.items()):
             text.append('%s = %s' % (k, repr(v)))
-        text = self.args_text + text
-        text = '\n'.join(text)
 
 
         if plt is not None:
-            plt.suptitle(fn.replace('#', '\n') +'\n' + text,
+            plt.suptitle(fn.replace('#', '\n') +'\n' + '\n'.join(text),
                          fontsize=8)
+            plt.figtext(0.12,0.12,'\n'.join(self.args_text))
+
+        text = self.args_text + text
+        text = '\n'.join(text)
 
         if not os.path.exists(p.data_dir):
             os.mkdir(p.data_dir)

### 2015-08-12 code/benchmark_min.py

@@ -12,8 +12,7 @@ class AdaptiveControl(benchmark.Benchmark):
         self.default('Kd', Kd=1.0)
         self.default('Ki', Ki=0.0)
         self.default('tau_d', tau_d=0.001)
-        self.default('T', T=0.5)
-        self.default('desired path', path='random')
+        self.default('T', T=8.0)
         self.default('period', period=4.0)
         self.default('use adaptation', adapt=False)
         self.default('n_neurons', n_neurons=500)
@@ -51,7 +50,7 @@ class AdaptiveControl(benchmark.Benchmark):
                 adapt = nengo.Ensemble(p.n_neurons, dimensions=p.D, radius=p.radius)
                 nengo.Connection(minsim, adapt, synapse=None)
                 conn = nengo.Connection(adapt, minsim, synapse=p.synapse,
-                        function=lambda x: 0,
+                        function=lambda x: [0]*p.D,
                         learning_rule_type=nengo.PES())
                 conn.learning_rule_type.learning_rate *= p.learning_rate
                 nengo.Connection(control, conn.learning_rule, synapse=None,
@@ -85,14 +84,24 @@ class AdaptiveControl(benchmark.Benchmark):
         t = sim.trange()
 
         N = len(q) / 2
-        offset = benchmark.find_offset(q[N:], d[N:])
-        if offset == 0:
-            offset = 1
+
+        # find an offset that lines up the data best (this is the delay)
+        offsets = []
+        for i in range(p.D):
+            q = sim.data[self.p_q][:,i]
+            d = sim.data[self.p_desired][:,i]
+            offset = benchmark.find_offset(q[N:], d[N:])
+            if offset == 0:
+                offset = 1
+            offsets.append(offset)
+        offset = int(np.mean(offsets))
+        delay = p.dt * offset
 
         if plt is not None:
-            plt.plot(t[offset:], d[:-offset])
-            plt.plot(t[offset:], d[offset:])
-            plt.plot(t[offset:], q[offset:])
+            plt.plot(t[offset:], d[:-offset], label='$q_d$')
+            #plt.plot(t[offset:], d[offset:])
+            plt.plot(t[offset:], q[offset:], label='$q$')
+            plt.legend(loc='upper left')
 
             #plt.plot(np.correlate(d, q, 'full')[len(q):])
 
[…]

### 2015-08-12 code/bootstrapci.py

@@ -1,0 +1,20 @@
+import random
+
+def sample(data):
+  for x in data:
+    yield random.choice(data)
+
+def bootstrapci(data,func,n=3000,p=0.95):
+    index=int(n*(1-p)/2)
+    r=[func(list(sample(data))) for i in range(n)]
+    r.sort()
+    return r[index], r[-index]
+
+def mean(x):
+    return sum(x)/len(x)
+
+
+if __name__ == '__main__':
+    data = [0.80696726764610904, 0.81252376300438622, 0.85538731552472702, 0.90055775771080115, 0.89766544499329926, 0.84625724686494552, 0.75323732419558742, 0.85881414581620064, 0.80899927994217202, 0.83466702339313126]
+
+    print bootstrapci(data, mean)

### 2015-08-12 code/benchmark.py
Adjsuted filename

@@ -72,7 +72,8 @@ class Benchmark(object):
 
         filename = name + '#' + ','.join(text)
 
-        filename = name + '#' + time.strftime('%H%M%S')
+        uid = np.random.randint(0x7FFFFFFF)
+        filename = name + '#' + time.strftime('%Y%m%d-%H%M%S')+('-%08x' % uid)
 
         return args, filename
 

### 2015-08-12 code/benchmark_min.py

@@ -12,7 +12,7 @@ class AdaptiveControl(benchmark.Benchmark):
         self.default('Kd', Kd=1.0)
         self.default('Ki', Ki=0.0)
         self.default('tau_d', tau_d=0.001)
-        self.default('T', T=8.0)
+        self.default('T', T=10.0)
         self.default('period', period=4.0)
         self.default('use adaptation', adapt=False)
         self.default('n_neurons', n_neurons=500)

### 2015-08-12 code/benchmark_min.py
Remove dependence on embodied_benchmarks

@@ -2,8 +2,11 @@ import time
 
 import nengo
 import numpy as np
-import embodied_benchmarks as bench
 
+
+from system import System
+from random_signal import Signal
+from pid import PID
 import benchmark
 
 class AdaptiveControl(benchmark.Benchmark):
@@ -31,7 +34,9 @@ class AdaptiveControl(benchmark.Benchmark):
         model = nengo.Network()
         with model:
 
-            system = bench.system.System(p.D, p.D, dt=p.dt, seed=p.seed, motor_noise=p.noise, sense_noise=p.noise, scale_add=p.scale_add,
+            system = System(p.D, p.D, dt=p.dt, seed=p.seed,
+                    motor_noise=p.noise, sense_noise=p.noise,
+                    scale_add=p.scale_add,
                     motor_scale=10,
                     motor_delay=p.delay, sensor_delay=p.delay,
                     motor_filter=p.filter, sensor_filter=p.filter)
@@ -41,13 +46,15 @@ class AdaptiveControl(benchmark.Benchmark):
 
             minsim = nengo.Node(minsim_system, size_in=p.D, size_out=p.D)
 
-            pid = bench.pid.PID(p.Kp, p.Kd, p.Ki, tau_d=p.tau_d)
-            control = nengo.Node(lambda t, x: pid.step(x[:p.D], x[p.D:]), size_in=p.D*2)
+            pid = PID(p.Kp, p.Kd, p.Ki, tau_d=p.tau_d)
+            control = nengo.Node(lambda t, x: pid.step(x[:p.D], x[p.D:]),
+                                 size_in=p.D*2)
             nengo.Connection(minsim, control[:p.D], synapse=0)
             nengo.Connection(control, minsim, synapse=None)
 
             if p.adapt:
-                adapt = nengo.Ensemble(p.n_neurons, dimensions=p.D, radius=p.radius)
+                adapt = nengo.Ensemble(p.n_neurons, dimensions=p.D,
+                                       radius=p.radius)
                 nengo.Connection(minsim, adapt, synapse=None)
                 conn = nengo.Connection(adapt, minsim, synapse=p.synapse,
                         function=lambda x: [0]*p.D,
@@ -56,20 +63,12 @@ class AdaptiveControl(benchmark.Benchmark):
                 nengo.Connection(control, conn.learning_rule, synapse=None,
                         transform=-1)
 
[…]

### 2015-08-12 code/system.py

@@ -1,0 +1,94 @@
+import numpy as np
+
+class System(object):
+    def __init__(self, d_state, d_motor, dt=0.001, seed=None,
+            scale_mult=1, scale_add=1, diagonal=True,
+            sense_noise=0.1, motor_noise=0.1,
+            motor_delay=0, motor_filter=None,
+            scale_inertia=0, motor_scale=1.0,
+            sensor_delay=0, sensor_filter=None,
+            nonlinear=True):
+
+        self.rng = np.random.RandomState(seed=seed)
+
+        self.d_motor = d_motor
+        self.d_state = d_state
+        self.dt = dt
+        self.motor_scale = motor_scale
+
+        sensor_steps = int(sensor_delay / dt) + 1
+        self.sensor_delay = np.zeros((sensor_steps, d_state), dtype=float)
+        motor_steps = int(motor_delay / dt) + 1
+        self.motor_delay = np.zeros((motor_steps, d_motor), dtype=float)
+        self.sensor_index = 0
+        self.motor_index = 0
+        self.scale_inertia = scale_inertia
+
+        self.sensor = np.zeros(d_state, dtype=float)
+        self.motor = np.zeros(d_motor, dtype=float)
+        if sensor_filter is None or sensor_filter < dt:
+            self.sensor_filter_scale = 0.0
+        else:
+            self.sensor_filter_scale = np.exp(-dt / sensor_filter)
+        if motor_filter is None or motor_filter < dt:
+            self.motor_filter_scale = 0.0
+        else:
+            self.motor_filter_scale = np.exp(-dt / motor_filter)
+
+        if diagonal:
+            assert d_state == d_motor
+            self.J = np.eye(d_motor) * scale_mult
+            #self.J = np.abs(np.diag(self.rng.randn(d_motor))) * scale_mult
+        else:
+            self.J = self.rng.randn(d_motor, d_state) * scale_mult
+        self.sense_noise = sense_noise
+        self.motor_noise = motor_noise
+
+        self.nonlinear = nonlinear
+        self.additive_a = self.rng.randn(d_state) 
+        if nonlinear:
+            D = len(self.nonlinearity(self.sensor))
+            self.additive_b = self.rng.randn(d_state, D)
+            self.additive_c = self.rng.randn(d_state) 
+            self.additive_d = self.rng.randn(d_state)
+        self.scale_add = scale_add
+
+        self.reset()
+    def reset(self):
+        self.state = self.rng.randn(self.d_state)
[…]

### 2015-08-12 code/vary_delay.py
changed fn

@@ -3,7 +3,9 @@ from benchmark_min import AdaptiveControl
 import numpy as np
 rng = np.random.RandomState()
 
-for i in range(10):
+for i in range(40):
     AdaptiveControl().run(no_figs=True, T=20, seed=i, max_freq=1,
                 scale_add=1, delay=rng.uniform(0, 0.04),
+                filter=rng.uniform(0, 0.01),
+                noise=rng.uniform(0, 0.1),
                 adapt=True, data_dir='vary_delay')

### 2015-08-12 code/benchmark_min.py
Fixed bug in rmse (was using q+noise rather than q)

@@ -46,6 +46,8 @@ class AdaptiveControl(benchmark.Benchmark):
 
             minsim = nengo.Node(minsim_system, size_in=p.D, size_out=p.D)
 
+            state_node = nengo.Node(lambda t: system.state)
+
             pid = PID(p.Kp, p.Kd, p.Ki, tau_d=p.tau_d)
             control = nengo.Node(lambda t, x: pid.step(x[:p.D], x[p.D:]),
                                  size_in=p.D*2)
@@ -68,7 +70,7 @@ class AdaptiveControl(benchmark.Benchmark):
             nengo.Connection(desired, control[p.D:], synapse=None)
 
             self.p_desired = nengo.Probe(desired, synapse=None)
-            self.p_q = nengo.Probe(minsim, synapse=None)
+            self.p_q = nengo.Probe(state_node, synapse=None)
             self.p_u = nengo.Probe(control, synapse=None)
         return model
 

### 2015-08-12 paper/paper.tex
starting on Travis' suggested fixes

@@
 hardware (and potentially across different algorithms implemented in said
-hardware).  This allows us to quantitatively compare systems, letting 
+hardware).  Having such a metric across instances and algorithms will allow us
+to quantitatively compare systems, letting
 researchers both measure the progress in the field and also directly
@@
 In this paper, we focus on the development of \emph{embodied} benchmarks.
-These are tasks where the output of the neuromorphic hardware \emph{influences
+These are dynamic tasks where the output of the neuromorphic hardware \emph{influences
 its own future input}.  This is in contrast to standard categorization or pattern
@@
 Incredibly finely detailed simulations can be created, filling in all of the
-details needed.  However, those resulting simulations \emph{cannot be run
-in real-time}.  This is a fundamental problem, in that most neuromorphic
+details needed.  However, accurate modelling of physical systems can very quickly
+become \emph{impractical to run in realtime}.
+This is a fundamental problem, in that most neuromorphic
 hardware is about real-time interactions, and there is no way to slow down
@@
 we spent the considerable amount of research effort needed to define a
-simulated environment for an embodied benchmark, that simulation could not
-be run fast enough to interact with most of the hardware that we would want
-to test.
+simulated environment for an embodied benchmark, running that simulation 
+fast enough to interact with most of the desired hardware would require more
+resources than are available.
 
@@
 The above considerations seem to indicate that even though using real-world
-phyiscal hardware for benchmarking is problematic, it is still better than
+physical hardware for benchmarking is problematic, it is still better than
 using simplistic simulations which may not generalize to real tasks.  However,
-we believe there is an alternate approach known as \emph{Minimal Simulation}
+we believe that the solution lies in an alternate approach taken from the
+field of evolutionary robotics, known as \emph{Minimal Simulation}
 (\citealt{Jakobi97evolutionaryrobotics}).

### 2015-08-12 code/plot_vary.py
nothing too useful
standard gp regression

@@ -42,34 +42,44 @@ if do_gp:
     Y.shape = Y.shape[0], 1
 
 
-    kernel = GPy.kern.RBF(input_dim=1, variance=1, lengthscale=0.005)
+    #kernel = GPy.kern.RBF(input_dim=1, variance=.3, lengthscale=0.5)
     #kernel = kernel+GPy.kern.RBF(input_dim=1, variance=1, lengthscale=0.005)
     #kernel = GPy.kern.Linear(1)
     #kernel = GPy.kern.Brownian(1)
-    m = GPy.models.GPRegression(X, Y,kernel)
+    #m = GPy.models.GPRegression(X, Y,kernel)
     #m = GPy.models.SparseGPRegression(X, Y,kernel)
     #m = GPy.models.GPCoregionalizedRegression([X], [Y],kernel)
     #m = GPy.models.GPHeteroscedasticRegression(X, Y,kernel)
     #m = GPy.models.GPVariationalGaussianApproximation(X, Y, kernel)
     #m = GPy.models.SparseGPRegression(X,Y,Z=np.random.rand(10,1)*0.04, kernel=kernel)
-    m.optimize('bfgs', messages=True)
+    #m.optimize('bfgs', messages=True)
     #m.optimize('tnc', messages=True)
 
     #print help(m.optimize)
-
-
+    
     #m.optimize('bfgs', max_iters=100)
     #m.optimize_restarts(num_restarts=10)
-    m.optimize()
+    #m.optimize()
 
-    print m
-
-    m.plot()
+    m = GPy.models.GPRegression(X,Y)
+    #m.rbf.lengthscale=.001
+    #m.rbf.variance=.009
+    
+    print m, '\n'
+    for i in range(5):
+        m.optimize('bfgs', max_iters=100) #first runs EP and then optimizes the kernel parameters
+        print 'iteration:', i,
+        print m
+        print ""
+    
+    #m.optimize(messages=True)
+    #m.optimize_restarts(num_restarts=10)
+    
+    m.plot() #posterior over the data
+    #m.plot_f() #The posterior
 
 else:
 
-
-    '''
     low = []
     high = []
     var = 0.0005
@@ -89,12 +99,11 @@ else:
         print xx, ci
 
     pylab.fill_between(x, low, high, color='#888888')
-    '''
 
     pylab.scatter(data_x, data_y, s=30, marker='x', color='k')
 
 
-
[…]

### 2015-08-13 code/plot_vary_neurons.py
results when varying number of neurons

@@ -1,0 +1,116 @@
+import os
+import sys
+
+import numpy as np
+import pylab
+
+import bootstrapci
+
+do_logx = True
+
+D = int(sys.argv[1])
+path = 'vary_neuronsD%d' % D
+
+data_x = []
+data_y = []
+for fn in os.listdir(path):
+    if fn.endswith('.txt'):
+        with open(os.path.join(path, fn)) as f:
+            text = f.read()
+        d = dict()
+        exec(text, d)
+
+        x = d['_n_neurons']
+        y = d['rmse']
+
+        if do_logx:
+            x = np.log2(x)
+        if y > 1.0:
+            continue
+
+        data_x.append(x)
+        data_y.append(y)
+        print x, y
+
+#data_x = data_x[:30]
+#data_y = data_y[:30]
+
+data_x = np.array(data_x)
+data_y = np.array(data_y)
+
+do_gp = False
+
+pylab.figure(figsize=(8,5))
+
+if do_gp:
+    import GPy
+
+    X = np.array(data_x)
+    X.shape = X.shape[0], 1
+    Y = np.array(data_y)
+    Y.shape = Y.shape[0], 1
+
+
+    kernel = GPy.kern.RBF(input_dim=1, variance=1, lengthscale=0.005)
+    #kernel = kernel+GPy.kern.RBF(input_dim=1, variance=1, lengthscale=0.005)
+    #kernel = GPy.kern.Linear(1)
+    #kernel = GPy.kern.Brownian(1)
+    m = GPy.models.GPRegression(X, Y,kernel)
+    #m = GPy.models.SparseGPRegression(X, Y,kernel)
+    #m = GPy.models.GPCoregionalizedRegression([X], [Y],kernel)
+    #m = GPy.models.GPHeteroscedasticRegression(X, Y,kernel)
+    #m = GPy.models.GPVariationalGaussianApproximation(X, Y, kernel)
+    #m = GPy.models.SparseGPRegression(X,Y,Z=np.random.rand(10,1)*0.04, kernel=kernel)
+    m.optimize('bfgs', messages=True)
+    #m.optimize('tnc', messages=True)
+
+    #print help(m.optimize)
+
+
+    #m.optimize('bfgs', max_iters=100)
+    #m.optimize_restarts(num_restarts=10)
+    m.optimize()
+
+    print m
+
+    m.plot()
+
+else:
+
+
+    def weighted_avg_and_std(values, weights):
[…]

### 2015-08-13 code/vary_neurons.py

@@ -1,0 +1,18 @@
+from benchmark_min import AdaptiveControl
+import traceback
+
+import numpy as np
+rng = np.random.RandomState()
+
+for i in range(100):
+    try:
+        AdaptiveControl().run(no_figs=True, seed=i, max_freq=1,
+                    scale_add=1, delay=rng.uniform(0, 0.01),
+                    D=15, T=60,
+                    filter=rng.uniform(0, 0.01),
+                    noise=rng.uniform(0, 0.1),
+                    n_neurons=int(2.0**rng.uniform(0,10)),
+                    adapt=True,
+                    data_dir='vary_neuronsD15')
+    except:
+        traceback.print_exc()

### 2015-08-13 code/benchmark_min.py
first attempt
gets an error in .run…

@@ -44,16 +44,16 @@ class AdaptiveControl(benchmark.Benchmark):
             def minsim_system(t, x):
                 return system.step(x)
 
-            minsim = nengo.Node(minsim_system, size_in=p.D, size_out=p.D)
+            minsim = nengo.Node(minsim_system, size_in=p.D, size_out=p.D*2)
 
             pid = PID(p.Kp, p.Kd, p.Ki, tau_d=p.tau_d)
             control = nengo.Node(lambda t, x: pid.step(x[:p.D], x[p.D:]),
                                  size_in=p.D*2)
-            nengo.Connection(minsim, control[:p.D], synapse=0)
+            nengo.Connection(minsim[:p.D], control[:p.D], synapse=0)
             nengo.Connection(control, minsim, synapse=None)
 
             if p.adapt:
-                adapt = nengo.Ensemble(p.n_neurons, dimensions=p.D,
+                adapt = nengo.Ensemble(p.n_neurons, dimensions=p.D*2,
                                        radius=p.radius)
                 nengo.Connection(minsim, adapt, synapse=None)
                 conn = nengo.Connection(adapt, minsim, synapse=p.synapse,

### 2015-08-13 code/system.py

@@ -3,7 +3,7 @@ import numpy as np
 class System(object):
     def __init__(self, d_state, d_motor, dt=0.001, seed=None,
             scale_mult=1, scale_add=1, diagonal=True,
-            sense_noise=0.1, motor_noise=0.1,
+            sense_noise=0.1, vel_sense_noise=0, motor_noise=0.1,
             motor_delay=0, motor_filter=None,
             scale_inertia=0, motor_scale=1.0,
             sensor_delay=0, sensor_filter=None,
@@ -18,6 +18,7 @@ class System(object):
 
         sensor_steps = int(sensor_delay / dt) + 1
         self.sensor_delay = np.zeros((sensor_steps, d_state), dtype=float)
+        self.vel_sensor_delay = np.zeros((sensor_steps, d_state), dtype=float)
         motor_steps = int(motor_delay / dt) + 1
         self.motor_delay = np.zeros((motor_steps, d_motor), dtype=float)
         self.sensor_index = 0
@@ -25,6 +26,7 @@ class System(object):
         self.scale_inertia = scale_inertia
 
         self.sensor = np.zeros(d_state, dtype=float)
+        self.vel_sensor = np.zeros(d_state, dtype=float)
         self.motor = np.zeros(d_motor, dtype=float)
         if sensor_filter is None or sensor_filter < dt:
             self.sensor_filter_scale = 0.0
@@ -42,6 +44,7 @@ class System(object):
         else:
             self.J = self.rng.randn(d_motor, d_state) * scale_mult
         self.sense_noise = sense_noise
+        self.vel_sense_noise = vel_sense_noise
         self.motor_noise = motor_noise
 
         self.nonlinear = nonlinear
@@ -88,7 +91,12 @@ class System(object):
         sensor = self.state + self.rng.randn(self.d_state) * self.sense_noise
         self.sensor = (self.sensor * self.sensor_filter_scale +
                        sensor * (1.0 - self.sensor_filter_scale))
+        
+        vel_sensor = self.dstate * self.dt + self.rng.randn(self.d_state) * self.vel_sense_noise            
+        self.vel_sensor = (self.vel_sensor * self.sensor_filter_scale +
+                       vel_sensor * (1.0 - self.sensor_filter_scale))
 
         self.sensor_delay[self.sensor_index] = self.sensor
+        self.vel_sensor_delay[self.sensor_index] = self.vel_sensor
[…]

### 2015-08-13 code/benchmark_min.py
Tweaks for consistent running

@@ -9,6 +9,11 @@ from random_signal import Signal
 from pid import PID
 import benchmark
 
+class ZeroDecoder(nengo.solvers.Solver):
+    weights = False
+    def __call__(self, A, Y, rng=None, E=None):
+        return np.zeros((A.shape[1], Y.shape[1]), dtype=float), []
+
 class AdaptiveControl(benchmark.Benchmark):
     def params(self):
         self.default('Kp', Kp=2.0)
@@ -55,11 +60,14 @@ class AdaptiveControl(benchmark.Benchmark):
             nengo.Connection(control, minsim, synapse=None)
 
             if p.adapt:
+
+
                 adapt = nengo.Ensemble(p.n_neurons, dimensions=p.D,
                                        radius=p.radius)
                 nengo.Connection(minsim, adapt, synapse=None)
                 conn = nengo.Connection(adapt, minsim, synapse=p.synapse,
                         function=lambda x: [0]*p.D,
+                        solver=ZeroDecoder(),
                         learning_rule_type=nengo.PES())
                 conn.learning_rule_type.learning_rate *= p.learning_rate
                 nengo.Connection(control, conn.learning_rule, synapse=None,

### 2015-08-13 code/plot_vary.py

@@ -6,7 +6,10 @@ import pylab
 
 import bootstrapci
 
+do_logx = True
+
 path = sys.argv[1]
+xkey = sys.argv[2]
 
 data_x = []
 data_y = []
@@ -17,11 +20,17 @@ for fn in os.listdir(path):
         d = dict()
         exec(text, d)
 
-        x = d['_delay']
+        x = d['_' + xkey]
         y = d['rmse']
+
+        if do_logx:
+            x = np.log2(x)
+        if y > 1.0:
+            continue
 
         data_x.append(x)
         data_y.append(y)
+        print x, y
 
 #data_x = data_x[:30]
 #data_y = data_y[:30]
@@ -31,7 +40,7 @@ data_y = np.array(data_y)
 
 do_gp = False
 
-pylab.figure(figsize=(8,4))
+pylab.figure(figsize=(8,5))
 
 if do_gp:
     import GPy
@@ -69,33 +78,29 @@ if do_gp:
 else:
 
 
-    '''
+    def weighted_avg_and_std(values, weights):
+        average = np.average(values, weights=weights)
+        variance = np.average((values-average)**2, weights=weights)
+        return (average, np.sqrt(variance))
+
     low = []
     high = []
-    var = 0.0005
-    x = np.linspace(0, 0.04, 100)
+    var = float(sys.argv[3])
+    x = np.linspace(min(data_x), max(data_x), 100)
     for xx in x:
         w = np.exp(-((data_x-xx)**2)/(2*var**2))
-        w /= sum(w)
-        samples = [np.random.choice(data_y, p=w) for i in range(200)]
-
-
-        #ci = bootstrapci.bootstrapci(samples, np.mean)
-        mean = np.mean(samples)
-        sd = np.std(samples)
-        ci = mean - 2*sd, mean+2*sd
+        mean, sd = weighted_avg_and_std(data_y, w)
+        ci = mean - 1*sd, mean+1*sd
         low.append(ci[0])
         high.append(ci[1])
-        print xx, ci
 
-    pylab.fill_between(x, low, high, color='#888888')
-    '''
+    pylab.fill_between(x, low, high, color='#aaaaaa')
 
     pylab.scatter(data_x, data_y, s=30, marker='x', color='k')
 
 
 
-pylab.xlim(0, 0.04)
-pylab.xlabel('delay (s)')
+#pylab.xlim(0, )
+#pylab.xlabel('delay (s)')
[…]

### 2015-08-13 code/plot_vary_neurons.py

@@ -18,7 +18,10 @@ for fn in os.listdir(path):
         with open(os.path.join(path, fn)) as f:
             text = f.read()
         d = dict()
-        exec(text, d)
+        try:
+            exec(text, d)
+        except:
+            continue
 
         x = d['_n_neurons']
         y = d['rmse']

### 2015-08-13 paper/paper.tex
Added travis' comments/suggestions

@@
 details needed.  However, accurate modelling of physical systems can very quickly
-become \emph{impractical to run in realtime}.
+become \emph{impractical to run in real time}.
 This is a fundamental problem, in that most neuromorphic
@@
 
+\subsection{Minimal Simulation as a Benchmark}
+
 Given the success of this approach for evolutionary robotics, we propose
-that it can be directly used for embodied neuromorphic benchmarks.  To do
-this, we create software simulations for each benchmark task.  These simulations
-must be fast enough to run in realtime (so that they can be controlled by
+using a minimal simulation as a neuromorphic benchmark.  First, we note that
+one important use of a benchmark is that by knowing how well particular
+hardware performs on that benchmark, you can reasonably infer how well
+that hardware will perform in other situations.  For example, if an image recognition
+algorithm performs well on the MNIST hand-written digit recognition
+benchmark, we can use that knowledge to guess that it may also perform well on
+a different recognition task.  Of course, this inference will fail if that
+algorithn has been specifically over-fit to exactly that one situtation.  For
+that reason, it would be useful to have a benchmark that covers a large range
+of variations on the task.  If the hardware performs well across that
+variability, then it is more likely to also work in whatever new situation we
+want to use it in.
+
+To achieve this, we need software simulations of the environment for the task.
+These simulations
+must be fast enough to run in real time (so that they can be controlled by
 real neuromorphic hardware), and they must be extremely variable.  Each time
@@
 
-All of that said, we also need real physical embodiment as part of any
-benchmark of this kind.  While we claim that reliable minimal simulations
-are relatively easy to create, they are not a widely-used technique, and the
-real proof of behaviour is always in the real world.  For this reason, we
-also argue that every benchmark based on a minimal simulation should also
-have an easy-to-construct physical analog.  However, this physical version
-is not to be used as the primary benchmark.  Rather, this is meant as a
-double-check that the system does actually behave as expected.
+The minimal simulation described above forms the core of our benchmark.
+However, the point of that benchmark is that is should do a reasonable job
+of generalizing to real-world physical tasks.  It is then useful to
+supplement that simulation benchmark with at least one easy-to-construct 
+physical analog.  This physical version would be one particular instance of the
+type of situation the benchmark is meant to cover.  For that reason, it
+is much more restrictive in terms of what general conclusions can be drawn
+from how well different hardware performs in that situation.  Rather, it is
+gives an explicit and understandable double-check that models that perform
+well on the simulation benchmark also perform well in a physical environment.
+Even considering the advantages of benchmarking using minimal simulations,
+it is still useful to have a real physical point of comparison as well.
 
@@
 the joints move to a particular desired position $q_d$.  Our only output is
-the signal $u$ (one for each motor) and our only input is the current position
-of each motor $q$.
+the signal $u$ (one for each motor) and our only inputs are the current position
+of each motor $q$ and the desired positions $q_d$.
 
@@
 external force due to gravity changes depending on the position of the arm $q$.
-The controller ends up having to use the accumulated integral to ``relearn'' the
+The controller ends up having to ``relearn'' the
 correct amount of extra force needed every time $q_d$ changes.
@@
 As long as there is enough hetereogenetity (i.e. as long as the neural activity
-forms a basis space that spans the desired space of functions), then the
-learning rule will continue to work.  This approach has been used in the
+forms a basis space that is capable of approximating the external forces), then the
+learning rule will continue to work.  This approach has been extended to
+biologically plausible neurons and been used in both the
 Recurrent Error-driven Adaptive Control Hierarchy model of human motor control
-(\citealt{dewolf2014}), and for quadcopter control [TODO: any citations for this?].
+(\citealt{dewolf2014}) and quadcopter control (\citealt{KomerThesis}).
 
@@
 Given this input, the neurons will produce some output $A$.
-we now form a weighted sum of these outputs $Ad$, where $d$ is a matrix (number of neurons by number of elements in $q$)
+We now form a weighted sum of these outputs $Ad$, where $d$ is a matrix (number of neurons by number of elements in $q$)
 that is initially all zeros.
@@
 
-On top of this, we need to add the external force for which the adaptive system
-is trying to compensate.  Rather than deciding on what specific external
-force, we \emph{randomly generate} an equation controlling this force.  We start with a small set of
-smooth functions on $X$ which are often found in dynamics equations ($x$, $x^2$,
-$sin(x)$).  We then generate an external force of $Q (\zeta X(\beta q + \gamma) + \eta)$
-where $\zeta$, $\beta$, $\gamma$, and $\eta$ are all randomly chosen fron $N(0,1)$
-and $Q$ is a scaling factor to control how strong this external force is.  The
-result is added to Equation 1.
+On top of this, we need to add an external perturbing force.  In a real system,
+this could be the effects of gravity given the current configuration of the
+motors, or of other unexpected influences.  
+Rather than chosing one particular fixed external force for our benchmark,
+we \emph{randomly generate} this force each time the benchmark is run.
+
+We start with a small set of
+smooth functions $f$ which are often found in dynamics equations (e.g. $x$, $x^2$,
+$sin(x)$).  We then generate an external force of $K_f (\zeta \cdot f(\beta \cdot q + \gamma) + \eta)$
+where $\zeta$, $\beta$, $\gamma$, and $\eta$ are all random vectors and
+$K_f$ is a scaling factor to control how strong this external force is.  The
+result is added to Equation 1.  Note that this means that if $q$ is 4-dimensional
+(i.e. if there are four joints being controlled) and if there are three smooth
+functions in $f$ (as there are here), then $\beta$, $\gamma$, and $\eta$ are 
+all vectors of length 4 and $\zeta$ is a 4x12 matrix.  All of these values are
+randomly chosen from the normal distribution $N(0,1)$.
 
@@
 
-Of course, we need physical confirmation that the minimal simulation defined
-in the previous section does a reasonable job.  For this, we define a
-particular easy-to-build example of a system that can be usefully controlled
-by this adaptive method.  In particular, we use the Lego Mindstorms EV3 robot kit.
+Of course, we also want an indication that the minimal simulation defined
+in the previous section is reasonably representative of the sorts of real-world
+situations in which we might want to use these controllers.  Importantly, 
+this physical instantiation does not have to exactly match one particular
+parameter setting of the minimal simulation.  Rather, we want a physical system
+that shares basic functional similarities to the minimal simulation defined
+previously.
 
-The physical body is depicted in [TODO: add figure].  It consists of a single
+For example, we want the inputs to the system to act like $u$, in that a
+positive number will increase some velocity $v$ which will in turn increase some
+sensor value $q$.  We want there to be some sort of external force applied that
+affects $q$, and we want that external force itself to be a function of $q$.
+We want there to be communication delays and noise in the sensory and motor
+system, and we want all of these effects to be somewhere within the extreme
+ranges covered by the minimal simulation.  While this cannot prove that hardware
+that is successful in simulation will always be successful in any similar
+real-world task, it at least gives an existence proof that there is at least
+one real-world task where it also performs reasonably.
+
+For this demonstration, we define an easy-to-build example of a system that can be usefully controlled
+by this adaptive method.  In particular, we use the Lego Mindstorms EV3 robot kit,
+organized as shown in [TODO: add figure].  It consists of a single
 motor, mounted in such a way that there is a significant amount of weight on
@@
 
-- magnitude of nonlinearity $Q$
+- magnitude of nonlinearity $K_f$
 

### 2015-08-13 paper/paper.tex
Added Chris' edits

@@
 \def\firstAuthorLast{Stewart {et~al.}} %use et al only if is more than 1 author
-\def\Authors{Terrence C. Stewart\,$^{1,*}$, Ashley Kleinhans\,$^{2}$ and Chris Eliasmith\,$^1$}
+\def\Authors{Terrence C. Stewart\,$^{1,*}$, Travis DeWolf\,$^{1}$, Ashley Kleinhans\,$^{2}$ and Chris Eliasmith\,$^1$}
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
@@
     embodied task; that is, a task where the output from the neuromorphic
-    hardware affects its future input.  However, these embodied situations
+    hardware affects its future input through some environment.  However, embodied situations
     are one of the primary potential uses of neuromorphic hardware.  To address
-    this, we present a methodology for embodied benchmarking and a particular
-    example of a benchmark for adaptive control of an arbitrary system.  These
-    benchmarks are flexible, in that they allow researchers to explicitly
+    this, we present a methodology for embodied benchmarking that makes use
+    of a hybrid of real physical embodiment and a type of ``minimal'' simulation.
+    Minimal simulation has been shown to lead to robust real-world performance,
+    while still maintaining the practical advantages of simulation, such as
+    making it easy for the same benchmark to be used by many researchers.
+    These benchmarks are flexible, in that they allow researchers to explicitly
     modify the benchmark to identify particular task domains where particular
-    hardware excels.  Furthermore, the benchmarks make use of a hybrid of
-    real physical embodiment and a type of ``minimal'' simulation that has been
-    shown to lead to robust real-world performance, allowing the same benchmark
-    to be used by many researchers.
+    hardware excels.  To demonstrate the method, we present a novel benchmark
+    where the task is to perform motor control on an arbitrary unknown system.
 
@@
 means that there is the potential to have complex algorithms running in
-embedded processing situations, without being a significant drain on battery
-life.  The question, of course, is to identify what sort of always-on or
-interactive functionality is feasible with these devices.
+embedded processing situations, without being a significant drain on available
+energy.  A crucial challenge is to identify what sort of always-on or
+interactive functionality can best exploit these devices.
 
-To evaluate applications of this hardware, we need benchmark tasks.  These
+To evaluate applications of neuromorphic hardware, we need benchmark tasks.  These
 tasks must allow us to compare across different instances of neuromorphic
-hardware (and potentially across different algorithms implemented in said
-hardware).  Having such a metric across instances and algorithms will allow us
-to quantitatively compare systems, letting
-researchers both measure the progress in the field and also directly
+hardware (and potentially across different algorithms implemented in that
+hardware).  Good benchmarks will allow us to quantitatively compare systems, letting
+researchers both measure the progress in the field, and also directly
 compare competing approaches.
@@
 These are dynamic tasks where the output of the neuromorphic hardware \emph{influences
-its own future input}.  This is in contrast to standard categorization or pattern
-identification tasks, where the input is some fixed sequence, and the main
-question is whether the hardware produces the correct output for each input
-(or input pattern).
+its own future input} through some environment.  This is in contrast to standard categorization or pattern
+identification tasks, where the input is some fixed sequence and the hardware
+must produce the correct output for each input (or input pattern).
 
 We believe embodied benchmarks should be of particular interest to
-neuromorphic research, given that many of the applications of neuromorphic
-hardware are in exactly this domain of embedded and interactive control
-of robotic or other physical systems.  However, embodiment raises a number
+neuromorphic research, given that the most compelling applications of neuromorphic
+hardware are likely to be in this domain of embedded and interactive control
+of robotic or other physical systems.  However, embodiment itself raises a number
 of issues that complicate the development of such benchmarks.  Rather than
@@
 must either specify a full physical system for that embodiment, or it must 
-provide software for a simulation of that system.  As we discuss below, either
-approach is problematic, and addressing this difficulty is the primary goal of
+provide software for a simulation of that system.  As we discuss below, both
+approaches are problematic.  
+Describing a method for overcoming these shortcomings is the primary goal of
 this paper.
@@
 two-way interaction with some sort of environment.  That is, the outputs from
-the neuromorphic hardware are sent to the environment where they cause some
-sort of effect, the results of which change the subsequent input.  For example,
+the neuromorphic hardware are sent to the environment where they cause
+an effect, the results of which change the subsequent input.  For example,
 the outputs might control the movement of a robot, which in turn affects the
@@
 To define such a benchmark, we need to be explicit about the embodiment.  If
-it is a situation where a robot is to be controlled, we need to be explicit
-about all the details of that robot.  What motors are there?  How are they
+a robot is to be controlled, we need to be explicit
+about all of the details of that robot.  What motors does it have?  How are they
 configured?  How strong are they?  What sensors are there?  Where are they
@@
 need access to that exact robot}.  If a benchmark is to be widely used, other
-researchers developing their own hardware should be able to do their own 
-testing.
+researchers developing their own neuromorphic hardware should be able to do their own 
+testing on the same benchmark system.
 
-Furthermore, using a physical robot also imposes significant practical
+Furthermore, using a physical robot imposes significant practical
 difficulties when performing extensive benchmark testing.  When testing, we
@@
 manually setting up the task, letting the test run, gathering the resulting
-data, and then resetting the robot back to the initial state.  This means that
+data, and then resetting the robot back to the initial state.  Consequently,
 issues like battery life become problematic, and not just because there is a
@@
 the system fails.  This means that some of the testing will involve parameter
-settings that lead to poor behaviour, which could involve physical damage to
-the robot itself.
+settings that lead to poor behaviour, which might have the undesireable 
+result of causing physical damage.
 
@@
 problematic.  First and foremost, without an actual real-world task, why
-should anyone have any confidence that the performance on the benchmark is
+should we have any confidence that the performance on the benchmark is
 reflective of the actual usefulness of the neuromorphic hardware?  It is
-very widely known that simulations of robots (or other physical systems) are
-often \emph{much} easier to control and better-behaved than the real thing.  The
+widely known that simulations of robots (or other physical systems) are
+often \emph{much} easier to control and better-behaved than the real thing [TODO: ref??].  The
 field of robotics is filled with algorithms that work well ``in theory,'' but
 fail when run on actual hardware.  We do not want a benchmark that falls into
-this trap of failing to generalize to real situations.
+this trap, giving high scores to hardware that does not turn out to
+functional well when generalized to real situations.
 
 Furthermore, neuromorphic hardware has another constraint that severely limits
-the possibilities of simulation.  The normal approach when a simulation is
-too simple to reflect reality is to add details to the simulation itself.
+the utility of simulation.  Typically, when a simulation is
+too simple to reflect reality, we add details to the simulation itself.
 Incredibly finely detailed simulations can be created, filling in all of the
@@
 become \emph{impractical to run in real time}.
-This is a fundamental problem, in that most neuromorphic
-hardware is about real-time interactions, and there is no way to slow down
+This is a fundamental problem, in that neuromorphic
+hardware is often tied to real-time interactions, and there is no way to slow down
 the hardware to match the simulated environment.  This means that even if
@@
 simulated environment for an embodied benchmark, running that simulation 
-fast enough to interact with most of the desired hardware would require more
+fast enough to interact with the desired hardware would require more
 resources than are available.
@@
 using simplistic simulations which may not generalize to real tasks.  However,
-we believe that the solution lies in an alternate approach taken from the
-field of evolutionary robotics, known as \emph{Minimal Simulation}
+we believe neuromorphic benchmarking can effectively exploit an approach 
+known as \emph{Minimal Simulation}
 (\citealt{Jakobi97evolutionaryrobotics}).
 
-First, we note that the problem faced here is remarkably similar to a problem
-faced by the evolutionary robotics community.  In evolutionary robotics, the
+This approach was first suggested in the context of evolutionary robotics.  
+Notably, the problem faced by embodied neuromorphic benchmarking is remarkably 
+similar to that faced earlier by these researchers. 
+In evolutionary robotics, the
 goal is to use genetic algorithms to \emph{evolve} systems that can control
 robots to perform various tasks.  These tasks can be as simple as navigation
-and obstacle avoidance, but have also included walking,
+and obstacle avoidance, but have also included more difficult tasks such as walking,
 object identification, and visual tracking [refs???].
 
-However, performing this evolution on real physical robots is problematic
+However, performing evolution on real physical robots is problematic
 for the same reasons that benchmarks on physical robots are problematic.  The
@@
 behaviour that can physically damage the robots; and they take a very long
-time to run.  For this reason, attempts were made to evolve using simulated
+time to run.  For this reason, attempts were made to evolve algorithms using simulated
 robots.  However, the general finding was that algorithms that worked on the
@@
 \emph{within the simulation itself}.  In other words, we make \emph{poor}
-simulations, but ensure that the way in which it is poor is itself
-variable.  We then make sure that the controllers work across that whole
+simulations, but ensure that the way in which they are poor is itself
+variable.  We are then in a position to ensure that the controllers work across that whole
 range of variability.  ``Instead of trying to eliminate the differences between
@@
 
-With this approach, it was possible to build minimal simulations that would
+With this approach, it became possible to build minimal simulations that would
 run faster than real-time and yet also be complex enough that if a system
 could successfully control the simulation, it was also likely to successfully
-control a real robot.  To achieve this, the simulations are made to be
+control a real robot.  To achieve this kind of transfer, the simulations were made to be
 unreliable in almost every respect.  For example, for a simulation of a simple
@@
 with this wide range of variability, and if it could handle that variability
-then we would have reason to believe it could also handle the real system.
+then there would be reason to believe it could also handle the real physical system.
 
-It is also worth noting that a minimal simulation only has to be a
+It is worth noting that a minimal simulation only has to be a
 good simulation \emph{for successful behaviour}.  That is, ``if we are evolving 
@@
 across the reality gap will neither
-hit walls nor go round in circles.''  If the controller is poor, we do not 
+hit walls nor go round in circles.''  [ref: jacobi thesis]  If the controller is poor, we do not 
 need the simulation to be at all accurate in exactly \emph{how} that poor
@@
 happens when two legs collide with each other.  Rather, if legs collide
-with each other, that is an indication that the walking behaviour is very poor,
-so as long as that result is indicated we can greatly simplify the simulation
+with each other, that is an indication that the walking behaviour is very poor.
+As long as that result is indicated we can greatly simplify the simulation
 by not including all the details necessary to deal with these sorts of physical

### 2015-08-13 paper/paper.tex
Added calibration discussion

@@
     hardware excels.  To demonstrate the method, we present a novel benchmark
-    where the task is to perform motor control on an arbitrary unknown system.
+    where the task is to perform motor control on an arbitrary system with
+    unknown external forces.
 
@@
 able to adjust the power sent to the motors $u$ and give position feedback
-$q$ from those motors at a rate of around 300Hz.
+$q$ from those motors at a rate of around 200Hz.
 
-Figure [???] shows the system being controlled by a simple PD controller.
-The desired joint angle $q_d$ is varied randomly over time.  Notice that the
-system consistently undershoots and overshoots the desired angle, due to
-its inability to compensate for gravity.
+Figure \ref{fig:ev3} shows the effects of adaptive control on this physical
+system.  Without adaptation (i.e. with a simple PD controller), there system
+state $q$ (the joint angle) overshoots the desired $q_d$.  This overshoot
+is largest when $q$ is large.  This is because the external force applied to
+the joint due to gravity is proportional to $sin(q)$.  The $q$ value also
+overshoots and comes back part-way, due to physical momentum.
 
-In Figure [???], we show the result of using the adaptive controller algorithm.
-The system quickly improves on the standard PD controller, learning to adjust
-its control signal $u$ to compensate for the unknown external forces.
+However, with adaptation (the right-hand side of Figure \ref{fig:ev3}),
+the system learns to counteract this extra force due to gravity.  After
+the first 5 seconds, the system is able to bring $q$ much closer to the
+desired $q_d$.
+
+Now that we have this physical example of the task out minimal simulation
+benchmark is meant to cover, we can use it to calibrate the parameters of
+the simulation.  For example, communication with the EV3 happens around
+200Hz, meaning that there must be a delay on the order of 0.005 seconds.  Given
+this, we set the delays in the simulation to be uniformly chosen between
+0 and 0.01.  Importantly, we do not need to exactly measure the delay in the
+EV3 robot --- we just make sure that the minimal simulation is worse.
+
+For sensor noise, we note that the EV3 rotation encoders for the motors (the
+devices that measure $q$) have a resolution of 0.0175 (1 degree).  This is a
+very different sort of noise than the gaussian noise used in the simulation,
+so we set the simulation noise to be much larger (uniformly distributed
+between 0 and 0.1).  Similarly, the motor resolution is 0.01, as it accepts 
+integer values up to 100, so we set the motor noise to be uniform between
+0 and 0.1.
+
+Finally, we can use the physical system to calibrate the relationship between
+$T$ (the maximum torque applied by the motor) and $K_f$ (the scaling factor
+of the external force).  After all, we do not want external forces that are
+so strong that the system does not have enough strength to counteract them.
+On the physical robot, in the worst-case scenario ($q=\pi/2$ or $-\pi/2$),
+the motors must be driven at around 0.3 times their maximum strength to
+balance the force of gravity.  (This can, of course, be adjusted by changing
+the weight and its position on the end of the arm).  If we arbitrarily fix $K_f$ to 1 and randomly
+generate external forces given the process described above, then 95\% of
+the time we get values between -3.75 and +3.75.  Since we want the motors to
+be strong enough to compensate for forces in that range, we set $T$ to 10.
 
@@
 \begin{center}
-\includegraphics[width=10cm]{logo1}% This is a *.jpg file
+\includegraphics[width=18cm]{figures/fig_ev3}
 \end{center}
- \textbf{\refstepcounter{figure}\label{fig:01} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
+ \textbf{\refstepcounter{figure}\label{fig:ev3} Figure \arabic{figure}.}{ Adaptive control of the EV3 lego robot used for calibrating the minimal simulation.
+ The effects of adaptation over two different desired trajectories are shown.  Without adaptation, the joints $q$ do not reach the desired $q_d$ when $q_d$ is large (which
+ is when the external force is largest).  With adaptation, $q$ is closer to $q_d$ after about 5 seconds, showing that the system has quickly learned to compensate.}
 \end{figure}

### 2015-08-13 paper/paper.tex
clarified minsim worseness
calrifications

@@
 
-\subsection{Cheap Robotics for Adaptive Control}
+\subsection{Calibrating the Minimal Simulation via Cheap Robotics}
 
@@
 0 and 0.01.  Importantly, we do not need to exactly measure the delay in the
-EV3 robot --- we just make sure that the minimal simulation is worse.
+EV3 robot --- we just make sure that the minimal simulation covers even worse
+cases.
 
@@
 very different sort of noise than the gaussian noise used in the simulation,
-so we set the simulation noise to be much larger (uniformly distributed
+so we set the simulation noise $\sigma_q$ to be much larger (uniformly distributed
 between 0 and 0.1).  Similarly, the motor resolution is 0.01, as it accepts 
-integer values up to 100, so we set the motor noise to be uniform between
+integer values up to 100, so we set the motor noise $\sigma_u$ to be uniform between
 0 and 0.1.

### 2015-08-13 code/benchmark_min.py
with velocity to adaptation pop

@@ -16,6 +16,7 @@ class AdaptiveControl(benchmark.Benchmark):
         self.default('Ki', Ki=0.0)
         self.default('tau_d', tau_d=0.001)
         self.default('T', T=10.0)
+        self.default('path', path='random')
         self.default('period', period=4.0)
         self.default('use adaptation', adapt=False)
         self.default('n_neurons', n_neurons=500)
@@ -62,9 +63,23 @@ class AdaptiveControl(benchmark.Benchmark):
                 conn.learning_rule_type.learning_rate *= p.learning_rate
                 nengo.Connection(control, conn.learning_rule, synapse=None,
                         transform=-1)
+                        
+            if p.path == 'random':
+                signal = Signal(p.D, p.period, dt=p.dt, max_freq=p.max_freq, seed=p.seed)
+                def desired_func(t):
+                    return signal.value(t) * np.pi / 2
 
-            signal = Signal(p.D, p.period, dt=p.dt, max_freq=p.max_freq, seed=p.seed)
-            desired = nengo.Node(signal.value)
+            elif p.path == 'step':
+                def desired_func(t):
+                    t = (t % p.period) / p.period
+                    if 0.25 < t < 0.5:
+                        return -np.pi / 2, -np.pi / 2
+                    elif 0.75 < t < 1.0:
+                        return np.pi / 2, np.pi / 2
+                    else:
+                        return 0, 0
+            
+            desired = nengo.Node(desired_func)
             nengo.Connection(desired, control[p.D:], synapse=None)
 
             self.p_desired = nengo.Probe(desired, synapse=None)
@@ -104,8 +119,8 @@ class AdaptiveControl(benchmark.Benchmark):
 
             #plt.plot(np.correlate(d, q, 'full')[len(q):])
 
+        diff = sim.data[self.p_desired][:-offset] - sim.data[self.p_q][offset:][:,:p.D]
 
-        diff = sim.data[self.p_desired][:-offset] - sim.data[self.p_q][offset:]
         diff = diff[N:]
         rmse = np.sqrt(np.mean(diff.flatten()**2))
 

### 2015-08-13 code/system.py

@@ -99,4 +99,4 @@ class System(object):
         self.sensor_delay[self.sensor_index] = self.sensor
         self.vel_sensor_delay[self.sensor_index] = self.vel_sensor
         self.sensor_index = (self.sensor_index + 1) % len(self.sensor_delay)
-        return self.sensor_delay[self.sensor_index], self.vel_sensor_delay[self.sensor_index]
+        return np.concatenate([self.sensor_delay[self.sensor_index], self.vel_sensor_delay[self.sensor_index]])

### 2015-08-14 paper/paper.tex
Added Chris' comments v2

@@
     embodied task; that is, a task where the output from the neuromorphic
-    hardware affects its future input through some environment.  However, embodied situations
+    hardware affects some environment, which then in turn affects the
+    hardware's future input.  However, embodied situations
     are one of the primary potential uses of neuromorphic hardware.  To address
-    this, we present a methodology for embodied benchmarking that makes use
+    this, we present a methodology for generating embodied benchmarks that makes use
     of a hybrid of real physical embodiment and a type of ``minimal'' simulation.
@@
     making it easy for the same benchmark to be used by many researchers.
-    These benchmarks are flexible, in that they allow researchers to explicitly
-    modify the benchmark to identify particular task domains where particular
-    hardware excels.  To demonstrate the method, we present a novel benchmark
-    where the task is to perform motor control on an arbitrary system with
+    This method is flexible enough to allow researchers to explicitly
+    modify the benchmarks to identify specific task domains where particular
+    hardware excels.  To demonstrate the method, we present a set of novel benchmarks
+    that focus on motor control for an arbitrary system with
     unknown external forces.
@@
 
-To define such a benchmark, we need to be explicit about the embodiment.  If
-a robot is to be controlled, we need to be explicit
-about all of the details of that robot.  What motors does it have?  How are they
+To define an embodied benchmark, we need to be explicit about the embodiment.  If
+a robot is to be controlled, we need to specify all of the details of that robot.  What motors does it have?  How are they
 configured?  How strong are they?  What sensors are there?  Where are they
@@
 this trap, giving high scores to hardware that does not turn out to
-functional well when generalized to real situations.
+functional well when deployed in real situations.
 
@@
 
+[TODO: maybe add a paragraph summarizing the horns of the dilemma so as to set up the
+role of minimal simulation more?]
+
 \subsection{Minimal Simulation}
@@
 using simplistic simulations which may not generalize to real tasks.  However,
-we believe neuromorphic benchmarking can effectively exploit an approach 
+we do not think this is the case.  Instead, we believe neuromorphic benchmarking can effectively exploit an approach 
 known as \emph{Minimal Simulation}
@@
 As long as that result is indicated we can greatly simplify the simulation
-by not including all the details necessary to deal with these sorts of physical
+by not including all the details necessary to model these complex physical
 interactions.
 
-\subsection{Minimal Simulation as a Benchmark}
+\subsection{Minimal Simulation for Benchmarking}
 
 Given the success of this approach for evolutionary robotics, we propose
-using a minimal simulation as a neuromorphic benchmark.  First, we note that
-one important use of a benchmark is that by knowing how well particular
-hardware performs on that benchmark, you can reasonably infer how well
+using minimal simulation for neuromorphic benchmarks.  We would argue that
+one important use of a benchmark is \emph{generalization}.  That is, by knowing how well particular
+hardware performs on a benchmark, you can reasonably infer how well
 that hardware will perform in other situations.  For example, if an image recognition
 algorithm performs well on the MNIST hand-written digit recognition
-benchmark, we can use that knowledge to guess that it may also perform well on
+benchmark, this suggests that it may also perform well on
 a different recognition task.  Of course, this inference will fail if that
-algorithn has been specifically over-fit to exactly that one situtation.  For
-that reason, it would be useful to have a benchmark that covers a large range
+algorithn has been specifically over-fit to that situtation.  For
+that reason, it is useful to have benchmarks that covers a wide range
 of variations on the task.  If the hardware performs well across that
-variability, then it is more likely to also work in whatever new situation we
-want to use it in.
+variability, then it is more likely to also work in sufficiently similar new situations.
 
-To achieve this, we need software simulations of the environment for the task.
+To achieve this kind of transfer, we need software simulations of the environment for the task.
 These simulations
 must be fast enough to run in real time (so that they can be controlled by
-real neuromorphic hardware), and they must be extremely variable.  Each time
-the simulation is run, different parameters will be chosed for this variability
+real neuromorphic hardware), and they must be extremely variable, to encourage
+robustness of the methods being benchmarked.  Each time
+the simulation is run, different parameters will be chosen to give significant variability
 (so one run might have a large degree of sensor noise while the next run has
@@
 
-The result should be a benchmark that can be run by any researcher.  The fact
+The result is a benchmark that can be used by any researcher.  The fact
 that it is a simulation means that source code can be shared, and that no
@@
 the benchmarked hardware.  For example, some hardware might only work with
-small amounts of sensor noise, or other hardware might only work when there
+small amounts of sensor noise, while other hardware might be most effective when there
 is significant delay in the motor response.  This flexibility in parameters
-in the benchmark allows researchers to explicitly characterize that particular
-siuations where their hardware excells.
+in the benchmark allows researchers to explicitly characterize those particular
+situations where their hardware excels.
 
@@
 
-The minimal simulation described above forms the core of our benchmark.
-However, the point of that benchmark is that is should do a reasonable job
-of generalizing to real-world physical tasks.  It is then useful to
-supplement that simulation benchmark with at least one easy-to-construct 
+The minimal simulation described above forms the core of our method for generating benchmarks.
+The purpose of these benchmarks is that is that they should do a reasonable job
+of generalizing to real-world physical tasks.  Consequently, it is important to
+supplement simulation benchmarks with at least one easy-to-construct 
 physical analog.  This physical version would be one particular instance of the
-type of situation the benchmark is meant to cover.  For that reason, it
+type of situation the benchmarks are meant to cover.  For that reason, the physical task
 is much more restrictive in terms of what general conclusions can be drawn
-from how well different hardware performs in that situation.  Rather, it is
-gives an explicit and understandable double-check that models that perform
-well on the simulation benchmark also perform well in a physical environment.
-Even considering the advantages of benchmarking using minimal simulations,
-it is still useful to have a real physical point of comparison as well.
+from how well different hardware performs in that situation.  Rather, the purpose is to give
+an explicit double-check that hardware that performs
+well on the simulation benchmarks also perform well in a physical environment.
 
-For this physical aspect of the benchmark, we recommend cheap, widely-available
+To keep the physical aspect simple, we recommend cheap, widely-available
 components.  This allows a greater chance for other researchers to have
@@
 
-It is important to note that there is actually a theoretical advantage to
+It is important to note that there is a theoretical advantage to
 using cheap robotics hardware for benchmarking, in addition to the practical
-advantages.  In particular, we \emph{don't want benchmarks that rely on
-high-speed, high-accuracy devices}.  The purpose of the benchmark
+advantages.  In particular, we \emph{do not want benchmarks that rely on
+high-speed, high-accuracy devices}.  The purpose of benchmarks
 is not to indicate how well this neuromorphic hardware works to control this
-one particular robot in this task.  Rather, the purpose of a benchmark is to
-characterise how well this neuromorphic hardware works on this task 
-\emph{in general}.  The variability in the minimal simulation means that it
+particular robot in this task.  Rather, the purpose of a benchmark is to
+characterise how well some specific neuromorphic hardware works on a task 
+\emph{in general}.  The variability introduced in the minimal simulation means that the hardware
 should be able to function across a wide variety of physical embodiments, and
 so if we are to choose one particular physical embodiment to test in the
-real world, then we should choose one that is not extremely high-precision.
-For this reason, we believe using cheap Lego robot is actually more useful
-for benchmarking than an expensive high-precision robot.\footnote{Of course,
+real world, then we should choose one that is not high-precision.
+For this reason, we believe using cheap Lego robots is actually more informative
+for benchmarking than expensive high-precision robots.\footnote{Of course,
 for more complex benchmark tasks we may need sensory and motor capabiities
@@
 
-\section{A Benchmark: Adaptive Motor Control}
+\section{Example: Adaptive Motor Control}
 
@@
 we now consider a basic control task.  Suppose we have a system with a number
-of joints $q$ and we want to send an output $u$ to those motors such that
-the joints move to a particular desired position $q_d$.  Our only output is
-the signal $u$ (one for each motor) and our only inputs are the current position
+of joints with positions $q = [q_1, q_2, ..., q_n]$ and we want to send a control signal $u = [u_1, u_2, ..., u_n]$ to the motors at each joint
+such that
+the joints move to a particular desired position $q_d = [q_{d,1}, q_{d,2}, ..., q_{d,n}]$.  The only output from the controller is
+the signal $u$ and the inputs are the current position
 of each motor $q$ and the desired positions $q_d$.
@@
 which helps to slow the system down as it approaches the desired position,
-thus avoiding overshooting and oscillation ($u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$),
-leading to the standard PD controller.  Both $K_p$ and $K_d$ are constants that can be
+thus avoiding overshoot and oscillation. This combination of terms leads to the
+standard PD controller $u=K_p(q_d - q) + K_d(\dot{q_d} - \dot{q})$.
+Both $K_p$ and $K_d$ are constants that can be
 tuned to particular situations.
@@
 gravity.  Since the PD controller always produces an output $u=0$ when $q=q_d$,
-it cannot compensate for this, and so the arm will stay stationary at some
+it cannot compensate for gravity, and so the arm will stay stationary at some
 angle below the desired angle. [TODO: add diagram]
 
-The standard solution to this problem is to add an I (integral) term ($K_i \int{(q_d-q) dt}$)
-to the controller, making it a PID controller.  The idea here is that as the difference between where it is and
-where we want it to be accumulates over time, the $K_i$ term gradually increases
-how much extra force is being applied until it is large enough to counteract
+The standard solution to this steady-state error is to add an I (integral) term ($K_i \int{(q_d-q) dt}$)
+to the controller, making it a PID controller.  As the difference between where it is ($q$) and
+where we want it to be ($q_d$) accumulates over time, the $K_i$ term will gradually increase the
+the extra controlled force $u$ that is being applied until it is large enough to counteract
 the external force of gravity (or whatever other external forces are present).
@@
 
-In some robotics applications, the solution to this problem is to mathematically
-analyze the geometry and mass of the system and compute exactly how much extra
+In some robotics applications, this problem is solved by mathematically
+analyzing the geometry and mass of the system to compute exactly how much extra
 force is needed.  In this particular case, the answer is straight-forward,
@@
 value and add it to our controller's output.  However, this assumes a perfectly
-even distribution of weight in the arm, ignores momentum and other forces, and 
+even distribution of weight in the arm, ignores momentum, friction, and other forces, and 
 gets much more complex as more joints are added.  Furthermore, if this
@@
 Fortunately, there is an adaptive solution to this problem, and it is one that
-fits well with neuromorphic hardware.  \cite{Slotine1987} shows that if you
-express the influence of these other external forces as $\tau=Y(q) \omega$
+fits well with neuromorphic hardware.  \cite{Slotine1987} show that if you
+express the influence of the external forces as $\tau=Y(q) \omega$
 (where $Y(q)$ is a fixed set of functions of $q$, such as $sin(q)$, and $\omega$
@@
 biologically plausible neurons and been used in both the
-Recurrent Error-driven Adaptive Control Hierarchy model of human motor control
+Recurrent Error-driven Adaptive Control Hierarchy (REACH) model of human motor control
 (\citealt{dewolf2014}) and quadcopter control (\citealt{KomerThesis}).
 
-This then suggests an explicit neuromorphic benchmark.  The input to the
+These considerations suggest that there is a neuromorphic-friendly family of 
+algorithms to address the general problem of controlling a wide variety of 
+physical systems.  Identifying those algorithms will allow us to benchmark 
+their performance across example tasks and physical configurations.  
+To implement these algorithms, the input to the
 neuromorphic hardware is $q$, the system state.  This input is fed to each
-neuron such that each neuron produces some output behaviour that is based on
+neuron such that each neuron produces some output activity that is based on
 this input.  Since $q$ will be multi-dimensional (if there is more than one
-joint), we may give each neuron a random weighting of each $q$ value ($J_i=e_i \cdot q$, 
-where $J_i$ is the input to neuron $i$, and $e_i$ is a randomly chosen vector\footnote{$e$ could also be chosen so as to regularly span the space of possibilities}).
+joint), we may give each neuron a random weighting of each $q$ value ($J_i=e_i \cdot q + b_i$, 
+where $J_i$ is the input to neuron $i$, and $e_i$ is a randomly chosen 
+vector\footnote{$e$ could also be chosen so as to regularly span the space of possibilities,
+or could be learned using some back-propogation of error method.  Here, for simplicity, we only
+consider the approach of randomly selecting $e_i$ and $b_i$.},
+and $b_i$ is a randomly chosen bias term).  
 Given this input, the neurons will produce some output $A$.
@@
 and so our actual output to the motor is $u + Ad$.  We then apply a learning
-rule on $d$ such that $\Delta d = \alpha A \times u$.
+rule on $d$ such that $\Delta d = \alpha A \times u$.  Here, $\alpha$ is a learning
+rate and the cross product is used so that we are applying the learning rule
+on all the joints simultaneously.
 
@@
 network, where the input and output layers are linear.  The first layer is $q$, the input state, one value for each joint.  The ``hidden'' layer
-is $A$, the activity of a large number of neurons.  The output layer again has
+is the neurons producing activities $A$, the activity of a large number of neurons.  The output layer again has
 one value per joint, and is the extra added signal to apply to the motors, $Ad$.
-Given that this is such a canonical example of the use of neural networks, we
-hope that the majority of neuromorphic hardware is flexible enough to implement
-exactly this model.
+Given that this is a canonical example of the use of neural networks, we
+expect that the majority of neuromorphic hardware is flexible enough to implement
+this model.  Importantly, it functions well with spiking neuron models as 
+well as non-spiking ones.  For spiking neurons, we consider $A$ to be the
+instantaneous measure of the output of a neuron (i.e. whether or not it is currently outputing a spike),
+filtered through a low-pass filter.  Further discussion of this sort of learning
+rule and comparison to biological spiking neurons is given by [ref?].
 
@@
 
-The one major step here that does not exist in a lot of neuromorphic hardware is
-the ability to update the weights $d$.  For hardware that does have a built-in
-learning rule, this rule is at least of a very common form, where the weight
+The rule for modifying the weights $d$ described here is
+of a very common form, as the weight
 update from a neuron is proportional to the activity of that neuron and an
-external error signal.  This makes it an instance of the ubiquitous delta rule,
-and hopefully supported by the hardware.
+external error signal.  This makes it an instance of the ubiquitous delta rule.
+Thus, neuromorphic hardware that has built-in learning will often be able to natively support this rule.
+However, some neuromorphic hardware does not intrinsically have
+the ability to update connection weights in this manner.
 
-However, if the neuromorphic hardware being benchmarked does not have the
-ability to update weights online using a learning rule of this form, then there
-are two solutions.  First, the multiplication by $d$ could be done on the
-output from the neuromorphic hardware.  There has to be some system to take
-the neural output from the hardware and send it to the motor (or to the simulation
-of the motor).  Instead of outputing the result of $Ad$, the hardware could
-output $A$ (the activity of all the neurons), and the interface to the motor
-could be responsible for doing the multiplication by $d$ and updating $d$
+In that case, there are at least two possible ways to implement this algorithm.
+First, the multiplication by $d$ can be done on the
+output from the neuromorphic hardware.  Any embodied neuromorphic system will have some method that takes
+the neural output from the hardware and sends it to the motors (or to the simulation
+of the motors).  Instead of sending the result of $Ad$, the hardware could
+send $A$ (the activity of all the neurons), and the interface to the motor
+can be responsible for doing the multiplication by $d$ and updating $d$
 according to the learning rule.
 
-Alternatively, we can use offline learning.  That is, rather than updating
-the weights $d$ all the time, we simply record $A$ and $u$, and then after
-a period of time stop the controller, compute the sum total of the changes
-to $d$, load the new value of $d$ onto the neuromorphic hardware, and then
+Alternatively, it may be possible to use offline learning.  That is, rather than updating
+the weights $d$ during the simulation, we record $A$ and $u$, and after
+a period of time stop the controller, compute the sum of the changes
+to $d$, load the new value of $d$ onto the neuromorphic hardware, and
 start the controller again.
+
+Given this variety of options for implementing adaptive algorithms of this type, 
+we believe it should be possible to benchmark most neuromorphic systems on 
+adaptive control tasks in this manner.
 
@@
 
-Now that we have defined the task, we can use the principles of Minimal
-Simulation to construct a flexible and variable simulated environment for
-testing adaptive control.  The idea is to make a bare-bones simulation of
-the system being controlled, with built-in variability.  If the neuromorphic
+Now that we have defined the task domain, we can use the principles of minimal
+simulation to construct a flexible and variable simulated environment for
+testing adaptive control.  In this case, we would like to develop a bare-bones simulation of
+the system being controlled, with significant variability.  If the neuromorphic
 controller works well across this variability, then it is likely to work well
@@
 
-The basic system variable is a vector of length $N$ holding the joint angles $q$.  
+The basic system variable is a vector of joint angles $q$.  
 Each joint has a velocity $v$.  The force applied by each motor is 
-related to the signal sent to the motor $u$, but will generally have some
-maximum value $T$, so we use $tanh(u)T$. For friction, we 
-simply scale the velocity by some factor $F$ every time step.  
-This results in the simplistic simulation as follows:
+related to the signal $u$ sent to the motors, but will generally have some
+maximum value $T$, so we use $tanh(u)T$ to determine the force as a function of the control signal. To account for friction, we 
+scale the velocity by some factor $F$ every time step.  
+This results in the simplistic simulation described by:
 
@@
 
-On top of this, we need to add an external perturbing force.  In a real system,
+In addition, we add a external perturbing forces.  In a real system,
 this could be the effects of gravity given the current configuration of the
@@
 Rather than chosing one particular fixed external force for our benchmark,
-we \emph{randomly generate} this force each time the benchmark is run.
+we \emph{randomly generate} this force each time the benchmark is run.  This
+ensures that the benchmark covers a wide range of possible external forces
+and motor configurations, rather than just one particular situation.
 
-We start with a small set of
-smooth functions $f$ which are often found in dynamics equations (e.g. $x$, $x^2$,
+Specifically, to generate this force, we start with a small set of
+smooth functions $f$ which are often found in dynamics equations ($x$, $x^2$,
 $sin(x)$).  We then generate an external force of $K_f (\zeta \cdot f(\beta \cdot q + \gamma) + \eta)$
 where $\zeta$, $\beta$, $\gamma$, and $\eta$ are all random vectors and
-$K_f$ is a scaling factor to control how strong this external force is.  The
-result is added to Equation 1.  Note that this means that if $q$ is 4-dimensional
+$K_f$ is a scaling factor to control how strong the external force is.  The
+result is added to Equation 1.  For example, if $q$ is 4-dimensional
 (i.e. if there are four joints being controlled) and if there are three smooth
-functions in $f$ (as there are here), then $\beta$, $\gamma$, and $\eta$ are 
-all vectors of length 4 and $\zeta$ is a 4x12 matrix.  All of these values are
+functions in $f$, then $\beta$, $\gamma$, and $\eta$ are 
+all vectors of length 4 and $\zeta$ is a 4x12 matrix.  To introduce
+significant variability, all of these values are
 randomly chosen from the normal distribution $N(0,1)$.
 
-Finally, we add random noise, delay, and filtering to both the input and the
+To complete the simulation, we introduce additional sources of variability: random noise, delay, and filtering to both the input and the
 output of the system.  For noise, we add $N(0,\sigma_u)$ to the control signal
@@
 also use a low-pass filter to smooth both values (with time constants $\tau_u$
-and $\tau_q$) after this noise is added.  Finally, both $q$ and $u$ are delayed
-by an amount of time $t_q$ and $t_u$.
+and $\tau_q$) after this noise is added, giving a damping effect.  Finally, both $q$ and $u$ are delayed
+by an amount of time $t_q$ and $t_u$ to reflect communication delays that are
+common in physical systems.
 
-The result is not meant to be a simulation of a particular physical embodiment.
-Rather, the variability in this simulation is meant to be extremely fast
-to simulate, and to make it hard to ``cheat''
-to control it.  In other words, if a controller manages to be able to control
+The resulting simulation is not meant to be an accurate portrayal of a particular physical embodiment.
+Rather, this simulation is meant to be extremely fast
+to simulate, and it is meant to be similarly difficult to control as a real system.
+In other words, if a controller manages to be able to control
 the various randomly created minimal simulations of embodiment that are
 generated with this approach, then we have reason to believe that it will also
-be successful at controlling real embodiments.  With this simulation and
-modern computers, we can easily simulate in real-time systems with dozens of joints and
-highly complex interactions between them, and measure how well an adaptive
+be successful at controlling real embodiments.  With this minimal simulation and
+modern computers, we can run real-time simulations of systems with dozens of joints that have
+highly complex interactions between them.  Consequently, we can effectively
+benchmark how well an adaptive
 controller deals with these situations.
@@
 
-Of course, we also want an indication that the minimal simulation defined
-in the previous section is reasonably representative of the sorts of real-world
-situations in which we might want to use these controllers.  Importantly, 
-this physical instantiation does not have to exactly match one particular
+It is important to ensure that the minimal simulation defined
+in the previous section is representative of the sorts of real-world
+situations in which we want to use these same controllers.  Importantly, 
+this physical instantiation does not have to exactly match any particular
 parameter setting of the minimal simulation.  Rather, we want a physical system
@@
 positive number will increase some velocity $v$ which will in turn increase some
-sensor value $q$.  We want there to be some sort of external force applied that
+sensor value $q$.  We want there to be some sort of external applied force that
 affects $q$, and we want that external force itself to be a function of $q$.
-We want there to be communication delays and noise in the sensory and motor
-system, and we want all of these effects to be somewhere within the extreme
-ranges covered by the minimal simulation.  While this cannot prove that hardware
-that is successful in simulation will always be successful in any similar
-real-world task, it at least gives an existence proof that there is at least
-one real-world task where it also performs reasonably.
+We want there to be communication delays and noise in the sensor and motor
+systems, and we want all of these effects to be somewhere within the
+ranges covered by the minimal simulation.  While implementing this kind of
+hardware analog cannot guarantee that neuromorphic hardware
+that is successful in simulation will be successful in every similar
+real-world task, it does provide an existence proof that there is at least
+one real-world task where the hardware performs similarly to how it performs in simulation.
 
-For this demonstration, we define an easy-to-build example of a system that can be usefully controlled
+For our specific demonstration, we describe an easy-to-build system that can be usefully controlled
 by this adaptive method.  In particular, we use the Lego Mindstorms EV3 robot kit,
@@
 motor, mounted in such a way that there is a significant amount of weight on
-the arm itself (the weight of the motor itself).  Multiple motors can be
+the arm itself (i.e., the weight of the motor itself).  Multiple motors can be
 added, and other configurations can be considered and should also be suitable
-for testing control, but here we just consider the basic case.
+for testing control, but here we consider only this basic case.
 

### 2015-08-14 paper/paper.tex
Added analysis section

@@
 
-- what do we measure?   performance: rmse and delay.  Also, power consumption?  cost?
+With the minimal simulation approach, we perform a benchmark by taking a
+particular piece of hardware, using it to control a minimal simulation, and
+measure some aspect of that simulation.  We perform this test multiple times,
+and each time we choose a different set of random parameters for the
+simulation.  A particular benchmark, then, will also provide particular
+distributions for these parameters.
 
-- for given hardware, how many neurons can be done in realtime?  that's how many should be used for comparison
+For example, Figure \ref{fig:analysis_basic} shows the root-mean-squared
+error (rmse) between $q$ and $q_d$ for three different neural systems.  For
+this benchmark, $t_q$, $t_u$, $\tau_q$, and $\tau_u$ are chosen from $U(0, 0.01)$ (the uniform distribution), $\sigma_q$ and $\sigma_u$ are from $U(0, 0.1)$,
+and $\beta$, $\gamma$, $\eta$, and $\zeta$ are all $N(0, 1)$.  As discussed
+above, $K_f$ is 1 and $T$ is 10.  $q_d$ is set to be gaussian white noise with
+a maximum frequency of 1Hz and RMS power of 1.  Each simulation is run
+for 20 seconds, and the error is computed on the last 10 seconds.
 
-- show how these things change as we change parameters in the simulation
+The first hardware tested on this benchmark is an Intel i5-3337U CPU running
+at 1.80GHz.  This is \emph{not} neuromorphic hardware, but serves as a
+point of comparison.  The learning algorithm was implemented using Nengo,
+a software toolkit for developing large-scale neural models that can be run
+on various hardware platforms \cite{nengo}.  For the neuron model, we used
+500 spiking Leaky-Integrate-and-Fire neurons.
 
-- delays
+The second hardware shown in Figure \ref{fig:analysis_basic} is an
+Nvidia Tesla C2075 GPU.  The same Nengo implementation was used, but
+retargetted to run on the GPU using OpenCL, with the same neuron model and
+number of neurons as the CPU.
 
-- number of motors $N$
+The third hardware system benchmarked is SpiNNaker \cite{furber2014}.  This
+neuromorphic hardware consists of 18 ARM processors on a single chip, 
+optimized for running neural models.  Thanks to a SpiNNaker implementation
+for Nengo \cite{mundy2015}, the same Nengo implementation as was used on the
+CPU and GPU can be run on this hardware as well.  Importantly, while the
+neuron model is the same, the actual implementation of this neuron model on
+SpiNNaker is very different from the implementation on the CPU and GPU, in
+that it relies on fixed-point computations and an asynchronous on-chip
+communication system.
 
-- magnitude of nonlinearity $K_f$
+\subsection{Computational Power Benchmark}
 
-CPU, GPU, and SpiNNaker?
+In Figure \ref{fig:analysis_basic}, all three systems perform equally well.
+However, on that benchmark all three systems are implementing exactly 500
+neurons.  The only thing that can be concluded from this is that the
+differences in neuron model across that hardware does not significantly
+impact performance.  A more interesting measure would be to compare based
+on how many neurons each piece of hardware is capable of running in real time.
+This is shown in Figure \ref{fig:analysis_compute}.  In this case, the CPU can
+run XXX neurons, the GPU XXX neurons, and a single SpiNNaker chip can run XXX
+neurons. Furthermore, we also show the effects of increasing the number
+of joints in the simulation.
+
+\subsection{Computational Efficiency Benchmark}
+
+While it is possible to run large neural models on standard CPUs and GPUs,
+one of the primary advantages of neuromorphic hardware is its power
+efficiency.  For this reason, the third benchmark fixes the number of neurons
+at however many that hardware can compute in 0.5 Watts.  While this is a
+reasonable number for SpiNNaker, neither the CPU nor the GPU can even run
+with that little power.  For this reason, we simply take the number of
+neurons that hardware can simulate in real time and scale it by the power
+consumption of the hardware.  The result is that the CPU can run XXX neurons,
+the GPU XXX neurons, and SpiNNaker can run XXX neurons.  The results of
+this benchmark are shown in Figure \ref{fig:analysis_power}.  This clearly
+shows the power advantage of neuromorphic hardware over standard CPUs and
+GPUs on this task.
+
+\subsection{Communication Delay Benchmark}
+
+We can also use this sort of benchmark to examine the effects of various
+parameters in the model.  For example, in Figure \ref{fig:analysis_delay_cpu}
+we show what happens as the delays $\t_q$ and $\t_u$ are increased.
+Importantly, the range on these parameters is larger than before ($U(0, 0.04)$
+rather than $U(0, 0.01)$).  This shows that if the delay is short (less than 0.02), the controller performs well, and if it is very large (greater than 0.03),
+the controller performs poorly.  However, for delays between 0.02 and 0.03,
+the controller will sometimes perform well and sometimes perform poorly.  The
+difference is due to the other random parameters in the system.  Interestingly,
+running this same benchmark on SpiNNaker in Figure \ref{fig:analysis_delay_spinn} shows that SpiNNaker actually performs better on this
+task.  Further analysis is needed to determine exactly why this is.
+
+\subsection{Scaling Benchmark}
+
+As a final comparison, we look at how this algorithm scales as the number
+of neurons increases and as the number of controlled motors $N$ increases.  This
+is a crucial benchmark, as the complexity of the task itself quickly
+increases with $N$ (since there the function computing the force applied
+at each joint is a function of \emph{all} the joint angles $q$).  This means
+that the external force the neural system must learn to predict increases
+in complexity exponentially as $N$ increases.  The quality of the control
+is thus dependent on how good an approximation the neurons can make of this
+complex nonlinear function.
+
+As shown in Figure \ref{fig:analysis_neurons_cpu}, adapting for unknown
+interacting forces on 15 joints is possible with 500 neurons.  This gives
+an indication of how many neurons are needed for different tasks, and can
+be extended to larger systems.
 
@@
 Operant conditioning
+
+\section {Conclusions}
+
+- And then a concluding thing pointing out that we've now defined N benchmarks, each of which looks at different aspects of the hardware.
+If people have new hardware that is good in particular situations, they can adjust the core benchmark be specifying different parameter distributions.
+This lets people be explicit about what their hardware is good at, while still using the same basic benchmark framework.
 

### 2015-08-14 code/benchmark_opencl.py
Added OpenCL benchmark

@@ -1,0 +1,132 @@
+import time
+
+import nengo
+import numpy as np
+
+
+from system import System
+from random_signal import Signal
+from pid import PID
+import benchmark
+
+class ZeroDecoder(nengo.solvers.Solver):
+    weights = False
+    def __call__(self, A, Y, rng=None, E=None):
+        return np.zeros((A.shape[1], Y.shape[1]), dtype=float), []
+
+class AdaptiveControl(benchmark.Benchmark):
+    def params(self):
+        self.default('Kp', Kp=2.0)
+        self.default('Kd', Kd=1.0)
+        self.default('Ki', Ki=0.0)
+        self.default('tau_d', tau_d=0.001)
+        self.default('T', T=10.0)
+        self.default('period', period=4.0)
+        self.default('use adaptation', adapt=False)
+        self.default('n_neurons', n_neurons=500)
+        self.default('learning rate', learning_rate=1.0)
+        self.default('max_freq', max_freq=1.0)
+        self.default('synapse', synapse=0.01)
+        self.default('radius', radius=1.0)
+        self.default('number of dimensions', D=1)
+        self.default('scale_add', scale_add=1)
+        self.default('noise', noise=0.1)
+        self.default('filter', filter=0.01)
+        self.default('delay', delay=0.01)
+
+    def model(self, p):
+
+        model = nengo.Network()
+        with model:
+
+            system = System(p.D, p.D, dt=p.dt, seed=p.seed,
+                    motor_noise=p.noise, sense_noise=p.noise,
+                    scale_add=p.scale_add,
+                    motor_scale=10,
+                    motor_delay=p.delay, sensor_delay=p.delay,
+                    motor_filter=p.filter, sensor_filter=p.filter)
+
+            self.system_state = []
+            def minsim_system(t, x):
+                self.system_state.append(system.state)
+                return system.step(x)
+
+            minsim = nengo.Node(minsim_system, size_in=p.D, size_out=p.D)
+
+            state_node = nengo.Node(lambda t: system.state)
+
+            pid = PID(p.Kp, p.Kd, p.Ki, tau_d=p.tau_d)
+            control = nengo.Node(lambda t, x: pid.step(x[:p.D], x[p.D:]),
+                                 size_in=p.D*2)
+            nengo.Connection(minsim, control[:p.D], synapse=0)
+            nengo.Connection(control, minsim, synapse=None)
+
+            if p.adapt:
+
+
[…]

### 2015-08-14 code/benchmark_min_spinn.py
Added spinnaker benchmarks

@@ -1,0 +1,137 @@
+import time
+
+import nengo
+import numpy as np
+
+
+from system import System
+from random_signal import Signal
+from pid import PID
+import benchmark
+
+class ZeroDecoder(nengo.solvers.Solver):
+    weights = False
+    def __call__(self, A, Y, rng=None, E=None):
+        return np.zeros((A.shape[1], Y.shape[1]), dtype=float), []
+
+class AdaptiveControl(benchmark.Benchmark):
+    def params(self):
+        self.default('Kp', Kp=2.0)
+        self.default('Kd', Kd=1.0)
+        self.default('Ki', Ki=0.0)
+        self.default('tau_d', tau_d=0.001)
+        self.default('T', T=10.0)
+        self.default('period', period=4.0)
+        self.default('use adaptation', adapt=False)
+        self.default('n_neurons', n_neurons=500)
+        self.default('learning rate', learning_rate=1.0)
+        self.default('max_freq', max_freq=1.0)
+        self.default('synapse', synapse=0.01)
+        self.default('radius', radius=1.0)
+        self.default('number of dimensions', D=1)
+        self.default('scale_add', scale_add=1)
+        self.default('noise', noise=0.1)
+        self.default('filter', filter=0.01)
+        self.default('delay', delay=0.01)
+
+    def model(self, p):
+
+        model = nengo.Network()
+        with model:
+
+            system = System(p.D, p.D, dt=p.dt, seed=p.seed,
+                    motor_noise=p.noise, sense_noise=p.noise,
+                    scale_add=p.scale_add,
+                    motor_scale=10,
+                    motor_delay=p.delay, sensor_delay=p.delay,
+                    motor_filter=p.filter, sensor_filter=p.filter)
+
+            self.system_state = []
+            self.system_times = []
+            self.system_desired = []
+            def minsim_system(t, x):
+                r = system.step(x)
+                self.system_state.append(system.state)
+                self.system_times.append(t)
+                self.system_desired.append(signal.value(t))
+                return r
+
+            minsim = nengo.Node(minsim_system, size_in=p.D, size_out=p.D)
+
+            state_node = nengo.Node(lambda t: system.state)
+
+            pid = PID(p.Kp, p.Kd, p.Ki, tau_d=p.tau_d)
+            control = nengo.Node(lambda t, x: pid.step(x[:p.D], x[p.D:]),
+                                 size_in=p.D*2)
[…]

### 2015-08-14 code/plot_vary.py

@@ -6,7 +6,7 @@ import pylab
 
 import bootstrapci
 
-do_logx = True
+do_logx = False
 
 path = sys.argv[1]
 xkey = sys.argv[2]
@@ -100,7 +100,8 @@ else:
 
 
 
-#pylab.xlim(0, )
-#pylab.xlabel('delay (s)')
+pylab.xlim(0, 0.04)
+pylab.xlabel('delay (s)')
 pylab.ylabel('rmse')
+pylab.ylim(-0.1, 1.0)
 pylab.savefig(path + '.png', dpi=100)

### 2015-08-14 code/plot_vary_neurons_all.py

@@ -1,0 +1,85 @@
+import os
+import sys
+
+import numpy as np
+import pylab
+
+pylab.figure(figsize=(12,8))
+
+do_logx = True
+
+index = 1
+for D in [1, 5, 10, 15]:
+    path = 'vary_neuronsD%d' % D
+
+    data_x = []
+    data_y = []
+    for fn in os.listdir(path):
+        if fn.endswith('.txt'):
+            with open(os.path.join(path, fn)) as f:
+                text = f.read()
+            d = dict()
+            try:
+                exec(text, d)
+            except:
+                continue
+
+            x = d['_n_neurons']
+            y = d['rmse']
+
+            if do_logx:
+                x = np.log2(x)
+            if y > 1.0:
+                continue
+
+            data_x.append(x)
+            data_y.append(y)
+            print x, y
+
+    #data_x = data_x[:30]
+    #data_y = data_y[:30]
+
+    data_x = np.array(data_x)
+    data_y = np.array(data_y)
+
+    pylab.subplot(2, 2, index)
+
+    def weighted_avg_and_std(values, weights):
+        average = np.average(values, weights=weights)
+        variance = np.average((values-average)**2, weights=weights)
+        return (average, np.sqrt(variance))
+
+    low = []
+    high = []
+    var = 1.0
+    x = np.linspace(0,10, 100)
+    for xx in x:
+        w = np.exp(-((data_x-xx)**2)/(2*var**2))
+        try:
+            mean, sd = weighted_avg_and_std(data_y, w)
+        except ZeroDivisionError:
+            mean, sd = 0.0, 0.0
+        ci = mean - 1*sd, mean+1*sd
+        low.append(ci[0])
+        high.append(ci[1])
+
+    pylab.fill_between(x, low, high, color='#aaaaaa')
+
+    pylab.scatter(data_x, data_y, s=30, marker='x', color='k')
+
+
+    if D > 1:
+        pylab.title('Adaptive control of %d interacting values' % D)
+    else:
+        pylab.title('Adaptive control of 1 value')
+
+    pylab.xlim(-1, 10)
+    xticks = [0, 2, 4, 6, 8, 10]
+    pylab.xticks(xticks, ['%d' % (2**xx) for xx in xticks])
+    if index in [3, 4]:
+        pylab.xlabel('number of neurons')
+    if index in [1,3]:
+        pylab.ylabel('rmse')
+    pylab.ylim(-0.05, 0.5)
+    index += 1
+pylab.savefig('fig_vary_neurons.png', dpi=100)

### 2015-08-14 code/plot_vary_neurons_all_spinn.py

@@ -1,0 +1,85 @@
+import os
+import sys
+
+import numpy as np
+import pylab
+
+pylab.figure(figsize=(12,8))
+
+do_logx = True
+
+index = 1
+for D in [1, 5, 10, 15]:
+    path = 'vary_neuronsD%d_spinn' % D
+
+    data_x = []
+    data_y = []
+    for fn in os.listdir(path):
+        if fn.endswith('.txt'):
+            with open(os.path.join(path, fn)) as f:
+                text = f.read()
+            d = dict()
+            try:
+                exec(text, d)
+            except:
+                continue
+
+            x = d['_n_neurons']
+            y = d['rmse']
+
+            if do_logx:
+                x = np.log2(x)
+            if y > 1.0:
+                continue
+
+            data_x.append(x)
+            data_y.append(y)
+            print x, y
+
+    #data_x = data_x[:30]
+    #data_y = data_y[:30]
+
+    data_x = np.array(data_x)
+    data_y = np.array(data_y)
+
+    pylab.subplot(2, 2, index)
+
+    def weighted_avg_and_std(values, weights):
+        average = np.average(values, weights=weights)
+        variance = np.average((values-average)**2, weights=weights)
+        return (average, np.sqrt(variance))
+
+    low = []
+    high = []
+    var = 1.0
+    x = np.linspace(0,10, 100)
+    for xx in x:
+        w = np.exp(-((data_x-xx)**2)/(2*var**2))
+        try:
+            mean, sd = weighted_avg_and_std(data_y, w)
+        except ZeroDivisionError:
+            mean, sd = 0.0, 0.0
+        ci = mean - 1*sd, mean+1*sd
+        low.append(ci[0])
+        high.append(ci[1])
+
+    pylab.fill_between(x, low, high, color='#aaaaaa')
+
+    pylab.scatter(data_x, data_y, s=30, marker='x', color='k')
+
+
+    if D > 1:
+        pylab.title('Adaptive control of %d interacting values' % D)
+    else:
+        pylab.title('Adaptive control of 1 value')
+
+    pylab.xlim(-1, 10)
+    xticks = [0, 2, 4, 6, 8, 10]
+    pylab.xticks(xticks, ['%d' % (2**xx) for xx in xticks])
+    if index in [3, 4]:
+        pylab.xlabel('number of neurons')
+    if index in [1,3]:
+        pylab.ylabel('rmse')
[…]

### 2015-08-14 code/system.py

@@ -82,6 +82,7 @@ class System(object):
             additive *= self.scale_add
         else:
             additive = self.additive_a * self.scale_add
+        self.additive = additive
         self.dstate += (np.dot(self.motor, self.J) + additive)
         self.state = self.state + self.dstate * self.dt
 
@@ -92,3 +93,19 @@ class System(object):
         self.sensor_delay[self.sensor_index] = self.sensor
         self.sensor_index = (self.sensor_index + 1) % len(self.sensor_delay)
         return self.sensor_delay[self.sensor_index]
+
+
+if __name__ == '__main__':
+    D = 1
+
+    data = []
+    for i in range(10000):
+        system = System(D, D)
+        system.step(np.zeros(D))
+        data.append(system.additive)
+
+    data = np.array(data).flatten()
+    data = np.sort(data)
+    index = int(len(data)*0.025)
+    print('95%% range: %f to %f' % (data[index], data[-index]))
+    

### 2015-08-14 code/vary_neurons_spinn.py

@@ -1,0 +1,21 @@
+from benchmark_min_spinn import AdaptiveControl
+import traceback
+
+import numpy as np
+rng = np.random.RandomState()
+
+for i in range(100):
+    for D in [1, 5, 10, 15]:
+        try:
+            AdaptiveControl().run(no_figs=True, seed=i, max_freq=1,
+                        scale_add=1, delay=rng.uniform(0, 0.01),
+                        D=D, T=60,
+                        filter=rng.uniform(0, 0.01),
+                        noise=rng.uniform(0, 0.1),
+                        n_neurons=int(2.0**rng.uniform(0,10)),
+                        adapt=True,
+                        debug=True,
+                        backend='nengo_spinnaker',
+                        data_dir='vary_neuronsD%d_spinn' % D)
+        except:
+            traceback.print_exc()

### 2015-08-14 paper/paper.tex
Fixed typo
Added edits Chris and Travis

@@
 filtered through a low-pass filter.  Further discussion of this sort of learning
-rule and comparison to biological spiking neurons is given by [ref?].
+rule and comparison to biological spiking neurons can be found in (\citealt{bekolay2013}).
 
@@
 
-In addition, we add a external perturbing forces.  In a real system,
+In addition, we add an external perturbing force.  In a real system,
 this could be the effects of gravity given the current configuration of the
@@
 With the minimal simulation approach, we perform a benchmark by taking a
-particular piece of hardware, using it to control a minimal simulation, and
-measure some aspect of that simulation.  We perform this test multiple times,
+particular piece of hardware and using it to control a minimal simulation while
+measuring some aspect of that simulation.  We perform this test multiple times,
 and each time we choose a different set of random parameters for the
@@
 parameters in the model.  For example, in Figure \ref{fig:analysis_delay_cpu}
-we show what happens as the delays $\t_q$ and $\t_u$ are increased.
+we show what happens as the delays $t_q$ and $t_u$ are increased.
 Importantly, the range on these parameters is larger than before ($U(0, 0.04)$

### 2015-08-15 paper/paper.tex
Added discussion and conclusion

@@
 
-\subsection{Online and offline learning}
+\subsection{Online and Offline Learning}
 
@@
 running this same benchmark on SpiNNaker in Figure \ref{fig:analysis_delay_spinn} shows that SpiNNaker actually performs better on this
-task.  Further analysis is needed to determine exactly why this is.
+task.  This is somewhat surprising, as we are currently using the slow
+Ethernet interface to SpiNNaker, rather than the high-speed I/O system that
+is meant for motor control.  Further analysis is needed to determine exactly why this is.
 
@@
 
-\section{Other Benchmarks}
+\section{Discussion}
 
-(sketch out how this approach could be applied to other task)
+The benchmarks described above all use the same underlying minimal simulation
+as a way to characterize the overall performance of particular hardware
+across a range of situations.  By adjusting the random distributions that
+define that range of situations, we generate different benchmarks that
+explore the capabilities of the systems in different ways.  This allows for
+an explicit depiction of the sorts of conditions in which particular
+neuromorphic hardware perfroms well.  After all, it is unlikely that one
+piece of neuromorphic hardware will be the best choice in all situations;
+rather, these benchmarks allow us to demonstrate the advantages and disadvantages
+of the hardware by looking at the same underlying system, but with multiple
+different distributions of parameters.
 
-Adaptive Jacobian
+Python software for the minimal simulation and the full benchmarks are
+available at \texttt{http://github.com/ctn-waterloo/ctn_benchmarks}.
 
-Driving down a corridor, avoiding obstacles
+\subsection{Benchmark Improvements}
 
-Classical conditioning
+The benchmarks presented here can be improved and further developed in many
+ways.  Most obviously, we need to benchmark more hardware, and in particular
+we would most like to consider analog neuromorphic hardware.  Interestingly, these
+benchmarks can also be run on software simulations of proposed hardware.
 
-Operant conditioning
+However, it is also clear that performance these benchmarks is a result of
+a combination of the hardware itself, the algorithm being run, and the system that
+interfaces the hardware to the environment.  Thus, for any given hardware, 
+we can explore improvements to the algorithm (better choices for $e$, 
+different learning rules, adaptive learning rates, adapting $K_p$ and $K_d$, etc.).
+For example, in the SpiNNaker hardware implementation not only can the neuron model
+be adjusted, but the distribution of the task across the multiple cores is also
+under programmer control.  Futhermore, SpiNNaker provides a custom I/O interface
+for high-speed communication that could be used to reduce communication delay.
+
+In addition, we would like to clarify and expand the space of physical tasks
+being approximated by the minimal simulation.  To do this, we will use the
+previously discussed method to 
+calibrate the model across a variety of different physical systems.  Of
+particular interest will be the space of random external force functions that
+are randomly generated.  Right now, this is a simple set of randomly
+generated vectors with a few common nonlinearities.  However, if we can
+identify classes of tasks that we are likely to want to control, we can
+create minimal simulations across those sorts of tasks.  For example, it
+would be possible to randomly generate $N$-joint arms with random arm
+lengths and random masses, and derive (an approximation of) the actual forces 
+that would be seen in that situation.  However, when doing this careful
+consideration would be needed to make sure that the randomly generated simulated arm
+had motors powerful enough to lift itself against gravity.
+
+\subsection{Other Benchmarks}
+
+The approach to benchmarking presented here can also be applied to other tasks,
+defining minimal simulations and evaluating hardware across a random sampling
+of that space.
+
+The most straightforward expansion would be to consider motor control where
+the state being controlled \emph{is not the motor state itself}.  That is,
+suppose we want to control the position of a hand $x$, but our output $u$
+only directly controls the joints $q$ of an arm.  The position of the hand $x$
+is some function of $q$, but this function may be unknown or highly complex.
+Crucially, there is a learning rule that is similar to the one discussed above
+that can learn this mapping (\citealt{adaptiveJacobian}), and we have had
+some success in using it for particular arm control tasks (citealt{dewolf2014}).
+This would give a benchmark with a much more flexible set of capabilities;
+for example a non-neural version of this learning algorithm was able to
+learn to control a robot arm to follow a certain path, given only the
+visual feedback of where the tip of its hand was, without knowing the
+geometry of its own hand or the position of the camera with relation to that
+hand (\citealt{adaptiveJacobian}).  While it is unlikely that such a
+system would work for \emph{any} possible sensorimotor combination, developing
+such a benchmark could identify the space of such embodiments that a
+particular algorithm running on particular neuromorphic hardware would be
+good at controlling.
+
+We also believe this approach to developing benchmarks would be applicable
+to more complex and abstract tasks.  For example, much of the original work
+on minimal simulation involved robot navigation and obstacle avoidance.  
+\cite{Jakobi97evolutonaryrobotics} created a
+minimal simulation of an environment with fixed walls, a mobile robot, and
+various sensory systems.  A benchmark involving movement between locations
+while avoiding collisions could be defined, and the environment could be
+enhanced with moving obstacles and a richer set of sensors.  The eventual
+goal would be to achieve reliable movement in an environment that includes
+people.
+
+Another direction to take this sort of benchmark would be classical and
+operant conditioning.  This sort of learning is extremely common in living
+creatures, and is vital for dealing with novel environments.  As such,
+it is natural to define tasks involving learning the associates between
+sensory events in the environment (akin to classical conditioning) and the
+associations between actions and desired sensory states (akin to operant
+conditioning).  Again, the approach of minimal simulation could be used
+to generate random complex sensory mappings, and various neuromorphic
+learning algorithms could be evaluated in terms of their ability to
+discern these patterns.  Importantly, this would also include the ability
+to unlearn these patterns as the environment changes.  Interestingly, the
+neuron-based classical conditioning learning rule used by \cite{verschure2003}
+is quite similar to the learning rule used in the adaptive control benchmark
+here.
 
@@
 
-- And then a concluding thing pointing out that we've now defined N benchmarks, each of which looks at different aspects of the hardware.
-If people have new hardware that is good in particular situations, they can adjust the core benchmark be specifying different parameter distributions.
-This lets people be explicit about what their hardware is good at, while still using the same basic benchmark framework.
+The primary purpose of this paper is to describe a new method for benchmarking
+neuromorphic hardware that addresses the problem of reliably benchmarking
+in a complex task that involves interaction with an environment.  This
+method involves building a ``minimal'' simulation; one that is extremely
+simple in terms of required computation, but that has a high degree of
+adjustable variability.  By benchmarking across that whole space of
+possibilities, we can identify hardware that performs well across that whole
+space, and is thus likely to be useful in real-world situations.  In order
+to identify what real-world situations are covered by a minimal simulation,
+we can calibrate the variability in that simulation with particular physical
+systems.
+
+We demonstrated this approach by defining a minimal simulation and a task
+based around adaptive motor control.  We presented a basic algorithm that
+can use neuromorphic hardware to improve performance on this task.  While this
+algorithm can be run on a variety of hardware, other algorithms could also
+be used on this task, and there are many implementation possibilities.
+Importantly, by measuring performance while adjusting the distributions of
+parameters in the benchmark, it is possible to characterize different aspects
+of the hardware, identifying how different aspects of the task affect
+performance for different hardware.  This was demonstrated by providing
+five different benchmarks, each based on the same minimal simulation, but
+setting parameters in different ways.  We believe this sort of flexibility
+is important in a benchmark, as it lets researchers be explicit about what
+their hardware is good at, while still using the same basic and shareable
+benchmark framework.
 
@@
 \section*{Acknowledgments}
+
 Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.

### 2015-08-15 paper/paper.tex
Added numbers for power comparison

@@
 widely known that simulations of robots (or other physical systems) are
-often \emph{much} easier to control and better-behaved than the real thing [TODO: ref??].  The
+often \emph{much} easier to control and better-behaved than the real thing 
+(see \citealt{Jakobi95}; \citealt{Koos2013}).  The
 field of robotics is filled with algorithms that work well ``in theory,'' but
@@
 vector\footnote{$e$ could also be chosen so as to regularly span the space of possibilities,
-or could be learned using some back-propogation of error method.  Here, for simplicity, we only
+or could be learned using some back-propagation of error method.  Here, for simplicity, we only
 consider the approach of randomly selecting $e_i$ and $b_i$.},
@@
 on how many neurons each piece of hardware is capable of running in real time.
-This is shown in Figure \ref{fig:analysis_compute}.  In this case, the CPU can
-run XXX neurons, the GPU XXX neurons, and a single SpiNNaker chip can run XXX
-neurons. Furthermore, we also show the effects of increasing the number
-of joints in the simulation.
+This is shown in Figure \ref{fig:analysis_compute}.  In this case, with the
+current implementation, the CPU can
+run 5200 neurons, the GPU 1500 neurons, and a single SpiNNaker core (of which
+there are 18 on one chip) can run 500
+neurons.  These values were measured empirically with the current versions
+of the reference Nengo implementation, the Nengo OpenCL implementation,
+and the Nengo SpiNNaker implementation (as of August 10, 2015).  
+Figure \ref{fig:analysis_compute}
+also shows the effects of varying the number of joints in the simulation,
+giving an indication of the complexity of the external forces that can be
+learned by this many neurons.
 
@@
 one of the primary advantages of neuromorphic hardware is its power
-efficiency.  For this reason, the third benchmark fixes the number of neurons
-at however many that hardware can compute in 0.5 Watts.  While this is a
-reasonable number for SpiNNaker, neither the CPU nor the GPU can even run
-with that little power.  For this reason, we simply take the number of
-neurons that hardware can simulate in real time and scale it by the power
-consumption of the hardware.  The result is that the CPU can run XXX neurons,
-the GPU XXX neurons, and SpiNNaker can run XXX neurons.  The results of
-this benchmark are shown in Figure \ref{fig:analysis_power}.  This clearly
+efficiency.  For this reason, the third benchmark normalizes the number of neurons
+based on power consumption.  With a power budget of 1W per chip (with 16 used cores),
+we estimate 0.0625W for the 500 neurons used here and round up to 0.1W to be
+conservative.  Of course, neither the CPU nor the
+GPU are designed to run on that little power.  For this reason,
+on this benchmark we simply and scale it by the power consumption of the hardware.
+For this power consumption we measure the difference between the idle power
+consumption and the consumption when running the benchmark.  For the Intel
+i5-3337U this was $34W - 11.5W = 12.5W$ and for the Nvidia Tesla C2075 GPU
+this was $74W - 70W = 4W$.  This value is much lower than the peak power
+consumption supported by the GPUi ($215W$), indicating that the current implementation
+does not make extensive use of the GPU.  Indeed, initial analysis indicates that
+the main bottleneck is communication between the GPU and the environment (i.e.,
+the minimal simulation), and we feel it is appropriate that this benchmark
+captures that limitation of the current GPU implementation.
+
+Given the above considerations, we run the benchmark indicating that the CPU can run 42 neurons per 0.1W,
+the GPU 38 neurons per 0.1W, and SpiNNaker can run 500 neurons per 0.1W.  If
+this number of neurons is kept fixed and the number of joints controlled is
+varied, we get the benchmark shown in Figure \ref{fig:analysis_power}.  This clearly
 shows the power advantage of neuromorphic hardware over standard CPUs and
@@
 an explicit depiction of the sorts of conditions in which particular
-neuromorphic hardware perfroms well.  After all, it is unlikely that one
+neuromorphic hardware performs well.  After all, it is unlikely that one
 piece of neuromorphic hardware will be the best choice in all situations;
@@
 Python software for the minimal simulation and the full benchmarks are
-available at \texttt{http://github.com/ctn-waterloo/ctn_benchmarks}.
+available at \texttt{http://github.com/ctn-waterloo/ctn\_benchmarks}.
 
@@
 be adjusted, but the distribution of the task across the multiple cores is also
-under programmer control.  Futhermore, SpiNNaker provides a custom I/O interface
+under programmer control.  Furthermore, SpiNNaker provides a custom I/O interface
 for high-speed communication that could be used to reduce communication delay.
@@
 that can learn this mapping (\citealt{adaptiveJacobian}), and we have had
-some success in using it for particular arm control tasks (citealt{dewolf2014}).
+some success in using it for particular arm control tasks (\citealt{dewolf2014}).
 This would give a benchmark with a much more flexible set of capabilities;

### 2015-08-15 paper/paper.tex
Added references

@@
 and obstacle avoidance, but have also included more difficult tasks such as walking,
-object identification, and visual tracking [refs???].
+collecting objects, and visual tracking (\citealt{Nolfi2000}).
 
@@
 possible to generalize to real behaviour; unfortunately, such complex
-simulations would run slower than real-time [refs].
+simulations would run slower than real-time (see \citealt{Husbands1992}; 
+\citealt{Husbands1993}).
 
@@
 simulation and reality, they are acknowledged, and mechanisms are put in place
-to prevent evolving controllers from relying on them.'' [ref: jacobi thesis].
+to prevent evolving controllers from relying on them.'' \citep[p.~48]{jakobiThesis}
 
@@
 across the reality gap will neither
-hit walls nor go round in circles.''  [ref: jacobi thesis]  If the controller is poor, we do not 
+hit walls nor go round in circles.''  \citep[p.~41]{jakobiThesis}  If the controller is poor, we do not 
 need the simulation to be at all accurate in exactly \emph{how} that poor
@@
 
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.
-Text Text Text Text Text Text  Text Text Text Text Text Text Text Text  Text Text Text Text Text Text Text Text Text  Text Text Text.
+We thank Andrew Mundy for extensive work on the Nengo SpiNNaker backend, and
+James Knight for his SpiNNaker implementation of the learning rule used here.
 
-\textit{Funding\textcolon} Text Text Text Text Text Text  Text Text.
+\textit{Funding\textcolon} 
+NSERC Discovery (grant 261453),
+ONR (N000141310419)
+AFOSR (FA8655-13-1-3084),
+Canada Research Chairs,
+and Canadian Foundation for Innovation
+Ontario Innovation Trust.
 

### 2015-08-16 paper/paper.tex
Figures and finishing sections

@@
 the system fails.  This means that some of the testing will involve parameter
-settings that lead to poor behaviour, which might have the undesireable 
+settings that lead to poor behaviour, which might have the undesirable 
 result of causing physical damage.
@@
 a different recognition task.  Of course, this inference will fail if that
-algorithn has been specifically over-fit to that situtation.  For
+algorithm has been specifically over-fit to that situation.  For
 that reason, it is useful to have benchmarks that covers a wide range
@@
 for benchmarking than expensive high-precision robots.\footnote{Of course,
-for more complex benchmark tasks we may need sensory and motor capabiities
+for more complex benchmark tasks we may need sensory and motor capabilities
 that are beyond that of a simple Lego robot.}
@@
 it cannot compensate for gravity, and so the arm will stay stationary at some
-angle below the desired angle. [TODO: add diagram]
+angle below the desired angle (Figure \ref{fig:system}).
 
@@
 to the controller, making it a PID controller.  As the difference between where it is ($q$) and
-where we want it to be ($q_d$) accumulates over time, the $K_i$ term will gradually increase the
+where we want it to be ($q_d$) accumulates over time, the $K_i$ term will gradually increase
 the extra controlled force $u$ that is being applied until it is large enough to counteract
@@
 well as non-spiking ones.  For spiking neurons, we consider $A$ to be the
-instantaneous measure of the output of a neuron (i.e. whether or not it is currently outputing a spike),
+instantaneous measure of the output of a neuron (i.e. whether or not it is currently outputting a spike),
 filtered through a low-pass filter.  Further discussion of this sort of learning
@@
 motors, or of other unexpected influences.  
-Rather than chosing one particular fixed external force for our benchmark,
+Rather than choosing one particular fixed external force for our benchmark,
 we \emph{randomly generate} this force each time the benchmark is run.  This
@@
 devices that measure $q$) have a resolution of 0.0175 (1 degree).  This is a
-very different sort of noise than the gaussian noise used in the simulation,
+very different sort of noise than the Gaussian noise used in the simulation,
 so we set the simulation noise $\sigma_q$ to be much larger (uniformly distributed
@@
 and $\beta$, $\gamma$, $\eta$, and $\zeta$ are all $N(0, 1)$.  As discussed
-above, $K_f$ is 1 and $T$ is 10.  $q_d$ is set to be gaussian white noise with
+above, $K_f$ is 1 and $T$ is 10.  $q_d$ is set to be Gaussian white noise with
 a maximum frequency of 1Hz and RMS power of 1.  Each simulation is run
@@
 The first hardware tested on this benchmark is an Intel i5-3337U CPU running
-at 1.80GHz.  This is \emph{not} neuromorphic hardware, but serves as a
+at 1.80GHz.  This is not neuromorphic hardware, but serves as a
 point of comparison.  The learning algorithm was implemented using Nengo,
@@
 of the reference Nengo implementation, the Nengo OpenCL implementation,
-and the Nengo SpiNNaker implementation (as of August 10, 2015).  
-Figure \ref{fig:analysis_compute}
-also shows the effects of varying the number of joints in the simulation,
-giving an indication of the complexity of the external forces that can be
-learned by this many neurons.
+and the Nengo SpiNNaker implementation (as of August 10, 2015).  All other
+parameters are as before.
 
@@
 consumption and the consumption when running the benchmark.  For the Intel
-i5-3337U this was $34W - 11.5W = 12.5W$ and for the Nvidia Tesla C2075 GPU
+i5-3337U this was $34W - 11.5W = 22.5W$ and for the Nvidia Tesla C2075 GPU
 this was $74W - 70W = 4W$.  This value is much lower than the peak power
-consumption supported by the GPUi ($215W$), indicating that the current implementation
+consumption supported by the GPU ($215W$), indicating that the current implementation
 does not make extensive use of the GPU.  Indeed, initial analysis indicates that
@@
 
-Given the above considerations, we run the benchmark indicating that the CPU can run 42 neurons per 0.1W,
-the GPU 38 neurons per 0.1W, and SpiNNaker can run 500 neurons per 0.1W.  If
-this number of neurons is kept fixed and the number of joints controlled is
-varied, we get the benchmark shown in Figure \ref{fig:analysis_power}.  This clearly
-shows the power advantage of neuromorphic hardware over standard CPUs and
-GPUs on this task.
+Given the above considerations, we run the benchmark indicating that the CPU can run 23 neurons per 0.1W,
+the GPU 38 neurons per 0.1W, and SpiNNaker can run 500 neurons per 0.1W.  
+As shown in Figure \ref{fig:analysis_power}, the neuromorphic hardware
+outperforms the CPU and GPU on this task.
 
@@
 We can also use this sort of benchmark to examine the effects of various
-parameters in the model.  For example, in Figure \ref{fig:analysis_delay_cpu}
-we show what happens as the delays $t_q$ and $t_u$ are increased.
+parameters in the model.  For example, Figure \ref{fig:analysis_delay}
+shows what happens as the delays $t_q$ and $t_u$ are increased.
 Importantly, the range on these parameters is larger than before ($U(0, 0.04)$
@@
 difference is due to the other random parameters in the system.  Interestingly,
-running this same benchmark on SpiNNaker in Figure \ref{fig:analysis_delay_spinn} shows that SpiNNaker actually performs better on this
-task.  This is somewhat surprising, as we are currently using the slow
+running SpiNNaker actually performs better on this
+task than the CPU (the GPU data is equivalent to the CPU and is not shown).  
+This is somewhat surprising, as we are currently using the slow
 Ethernet interface to SpiNNaker, rather than the high-speed I/O system that
@@
 on minimal simulation involved robot navigation and obstacle avoidance.  
-\cite{Jakobi97evolutonaryrobotics} created a
+\cite{Jakobi97evolutionaryrobotics} created a
 minimal simulation of an environment with fixed walls, a mobile robot, and
@@
 We thank Andrew Mundy for extensive work on the Nengo SpiNNaker backend, and
-James Knight for his SpiNNaker implementation of the learning rule used here.
+James Knight for his prototype SpiNNaker implementation of the learning rule used here.
 
@@
 \begin{center}
+\includegraphics[width=10cm]{figures/SingleArmSystem}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:system} Figure \arabic{figure}.}{ 
+     An abstract physical system that could be controlled.  In order to hold
+     the joint $q$ at a desired angle $q_d$, a force must be applied to
+     counteract the pull of gravity.  The magnitude of this force is a
+ function of $q$ which can be learned.}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
 \includegraphics[width=18cm]{figures/fig_ev3}
@@
 
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=18cm]{figures/plot_basic}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:analysis_basic} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems.
+     Each system is running 500 neurons.  The hardware do not statistically
+     significantly differ, and are all clear improvements over the having
+     no adaptation at all.
+     Scatterplots show individual runs, the 
+     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=18cm]{figures/plot_compute}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:analysis_compute} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems
+     in terms of their performance when running as many neurons as they are capable of in real time.  Scatterplots show individual runs, the 
+     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  In this case,
+     all three hardware systems perform equally well.}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=18cm]{figures/plot_power}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:analysis_power} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems
+     in terms of their performance when running as many neurons as they are capable of per 0.1 Watt of power consumption.  Scatterplots show individual runs, the 
+     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  The SpiNNaker system is a statistically
+ significant improvement over the other two.}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=18cm]{figures/plot_delay}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:analysis_delay} Figure \arabic{figure}.}{ Benchmark results comparing the effects
+     of communication delay for the CPU and SpiNNaker systems.  The delays $t_q$ and $t_u$ are randomly varied.  Shaded
+ area is the mean plus or minus one standard deviation, smoothed with a Gaussian kernel of $\sigma=0.005$.}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=18cm]{figures/plot_vary_neurons}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:analysis_neurons_cpu} Figure \arabic{figure}.}{ Benchmark results examining the relationship
+     between number of neurons and the number of simulated joints $N$.  The
+     benchmark was run on the Intel i5-3337U CPU.  Shaded
+ area is the mean plus or minus one standard deviation, smoothed with a Gaussian kernel of $\sigma=1$ in the $log_2$ domain.}
+\end{figure}
 %\begin{figure}

### 2015-08-16 paper/paper.tex
Added circles to figure for clarity

@@
  The effects of adaptation over two different desired trajectories are shown.  Without adaptation, the joints $q$ do not reach the desired $q_d$ when $q_d$ is large (which
- is when the external force is largest).  With adaptation, $q$ is closer to $q_d$ after about 5 seconds, showing that the system has quickly learned to compensate.}
+ is when the external force is largest).  With adaptation, $q$ is closer to $q_d$ after about 5 seconds, showing that the system has quickly learned to compensate.
+ Points in time where the improvement is clearest are circled.}
 \end{figure}

### 2015-08-16 paper/paper.tex
Added horns of dilemma; added ref to NEF

@@
 
-[TODO: maybe add a paragraph summarizing the horns of the dilemma so as to set up the
-role of minimal simulation more?]
+We are thus left with a situation where any benchmark we might define for
+a rich, embodied task will be either impractical for different researchers to
+run (if it is physically embodied), or inapplicable to real-world situations
+(if it is performed in simulation).  Neither option is particularly appealing.
 
@@
 
-The above considerations seem to indicate that even though using real-world
+The above considerations could be taken as an argument that even though using real-world
 physical hardware for benchmarking is problematic, it is still better than
@@
 rule and comparison to biological spiking neurons can be found in (\citealt{bekolay2013}).
+This type of neural modelling forms the foundation of \cite{eliasmith2003}'s
+Neural Engineering Framework, and \cite{stewart2014} have shown that spiking
+and non-spiking neurons can be used in this manner to implement a wide
+variety of computations.
 
@@
 by this adaptive method.  In particular, we use the Lego Mindstorms EV3 robot kit,
-organized as shown in [TODO: add figure].  It consists of a single
-motor, mounted in such a way that there is a significant amount of weight on
-the arm itself (i.e., the weight of the motor itself).  Multiple motors can be
+organized as shown in Figure \ref{fig:lego}.  It consists of a single
+motor, mounted such that the full weight of a second (unused) motor applies
+a significant force on the arm itself.  Multiple motors can be
 added, and other configurations can be considered and should also be suitable
-for testing control, but here we consider only this basic case.
+for benchmarking, but here we consider only this basic case.
 
@@
 AFOSR (FA8655-13-1-3084),
+Mitacs Postdoctoral Fellowship,
 Canada Research Chairs,
-and Canadian Foundation for Innovation
+Canadian Foundation for Innovation,
 Ontario Innovation Trust.
@@
  function of $q$ which can be learned.}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=18cm]{figures/lego}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:lego} Figure \arabic{figure}.}{ A
+     simple physical robot embodiment for calibrating the minimal simulation.
+     To rotate the central motor to the desired position $q$, enough force
+ must be added to $u$ to counteract the weight of the second unused motor.}
 \end{figure}

### 2015-08-16 paper/paper.tex
Tweaked working of NEF reference

@@
 This type of neural modelling forms the foundation of \cite{eliasmith2003}'s
-Neural Engineering Framework, and \cite{stewart2014} have shown that spiking
+Neural Engineering Framework, which has shown that spiking
 and non-spiking neurons can be used in this manner to implement a wide
-variety of computations.
+variety of computations (e.g., \citealt{stewart2014}).
 

### 2015-08-16 paper/paper.tex
Added Chris's penultimate comments

@@
 
-With the minimal simulation approach, we perform a benchmark by taking a
-particular piece of hardware and using it to control a minimal simulation while
-measuring some aspect of that simulation.  We perform this test multiple times,
-and each time we choose a different set of random parameters for the
-simulation.  A particular benchmark, then, will also provide particular
-distributions for these parameters.
+To run a benchmark using the proposed
+minimal simulation approach, there are three main steps: 1. identify
+the neuromorphic hardware to be tested; 2. construct the minimal
+simulation; 3. determine a metric (e.g. root-mean-squared error; rmse)
+to record; 5. specify distributions for any parameters in the
+simulation.  We then perform multiple runs of the simulation, each
+time choosing different values from the parameter distributions.  We
+can then plot how the metric varies as a function of a particular parameter
+of interest, or how it compares across different hardware for a
+given set of parameter distributions.
 
 For example, Figure \ref{fig:analysis_basic} shows the root-mean-squared
-error (rmse) between $q$ and $q_d$ for three different neural systems.  For
+error (rmse) between $q$ and $q_d$ for three different hardware systems.  For
 this benchmark, $t_q$, $t_u$, $\tau_q$, and $\tau_u$ are chosen from $U(0, 0.01)$ (the uniform distribution), $\sigma_q$ and $\sigma_u$ are from $U(0, 0.1)$,
@@
 The first hardware tested on this benchmark is an Intel i5-3337U CPU running
-at 1.80GHz.  This is not neuromorphic hardware, but serves as a
-point of comparison.  The learning algorithm was implemented using Nengo,
+at 1.80GHz.  This is not neuromorphic hardware, but provides a
+useful baseline.  The learning algorithm was implemented using Nengo,
 a software toolkit for developing large-scale neural models that can be run
-on various hardware platforms \cite{nengo}.  For the neuron model, we used
-500 spiking Leaky-Integrate-and-Fire neurons.
+on various hardware platforms (\citealt{nengo}).  For the neuron model, we used
+500 spiking Leaky-Integrate-and-Fire (LIF) neurons.
 
-The second hardware shown in Figure \ref{fig:analysis_basic} is an
-Nvidia Tesla C2075 GPU.  The same Nengo implementation was used, but
+The second hardware used to generate Figure \ref{fig:analysis_basic} is an
+Nvidia Tesla C2075 GPU, hardware that is often used for special purpose computing
+and neural network simulations.  The same Nengo implementation was used, but
 retargetted to run on the GPU using OpenCL, with the same neuron model and
@@
 
-The third hardware system benchmarked is SpiNNaker \cite{furber2014}.  This
+The third hardware system benchmarked is SpiNNaker (\citealt{furber2014}).  This
 neuromorphic hardware consists of 18 ARM processors on a single chip, 
 optimized for running neural models.  Thanks to a SpiNNaker implementation
-for Nengo \cite{mundy2015}, the same Nengo implementation as was used on the
-CPU and GPU can be run on this hardware as well.  Importantly, while the
-neuron model is the same, the actual implementation of this neuron model on
+for Nengo (\citealt{mundy2015}), the same Nengo implementation that was used on the
+CPU and GPU is run on this hardware as well.  Importantly, while the
+basic neuron model is the same, the actual implementation of this neuron model on
 SpiNNaker is very different from the implementation on the CPU and GPU, in
@@
 However, on that benchmark all three systems are implementing exactly 500
-neurons.  The only thing that can be concluded from this is that the
+neurons.  This demonstrates that the
 differences in neuron model across that hardware does not significantly
-impact performance.  A more interesting measure would be to compare based
-on how many neurons each piece of hardware is capable of running in real time.
-This is shown in Figure \ref{fig:analysis_compute}.  In this case, with the
-current implementation, the CPU can
+impact performance.  Given that embodied models rely on real-time simulation,
+it is also important to determine how many neurons each piece of hardware is capable of running in real time.
+This is shown in Figure \ref{fig:analysis_compute}.  With the current implementation, the CPU can
 run 5200 neurons, the GPU 1500 neurons, and a single SpiNNaker core (of which
 there are 18 on one chip) can run 500
-neurons.  These values were measured empirically with the current versions
+neurons.\footnote{The full chip can run much more than 500 neurons; here we use just a single one of the cores on the SpiNNaker chip.}  These values were measured empirically with the current versions
 of the reference Nengo implementation, the Nengo OpenCL implementation,
@@
 we estimate 0.0625W for the 500 neurons used here and round up to 0.1W to be
-conservative.  Of course, neither the CPU nor the
+conservative.  Neither the CPU nor the
 GPU are designed to run on that little power.  For this reason,
-on this benchmark we simply and scale it by the power consumption of the hardware.
+on this benchmark we scale the number of neurons by the power consumption of the hardware.
 For this power consumption we measure the difference between the idle power
@@
 consumption supported by the GPU ($215W$), indicating that the current implementation
-does not make extensive use of the GPU.  Indeed, initial analysis indicates that
+does not make extensive use of the GPU for this task.  Indeed, initial analysis indicates that
 the main bottleneck is communication between the GPU and the environment (i.e.,
 the minimal simulation), and we feel it is appropriate that this benchmark
-captures that limitation of the current GPU implementation.
+captures that limitation of the current GPU implementation.  The GPU could easily
+run many more neurons than this in real time, if those neurons were not connected
+to an environment.  However, that would not be useful for an embodied task.
 
-Given the above considerations, we run the benchmark indicating that the CPU can run 23 neurons per 0.1W,
+Given the above considerations, the benchmark indicates that the CPU can run 23 neurons per 0.1W,
 the GPU 38 neurons per 0.1W, and SpiNNaker can run 500 neurons per 0.1W.  
@@
 
-We can also use this sort of benchmark to examine the effects of various
+We can also use minimal simulation benchmarks to examine the effects of various
 parameters in the model.  For example, Figure \ref{fig:analysis_delay}
-shows what happens as the delays $t_q$ and $t_u$ are increased.
-Importantly, the range on these parameters is larger than before ($U(0, 0.04)$
-rather than $U(0, 0.01)$).  This shows that if the delay is short (less than 0.02), the controller performs well, and if it is very large (greater than 0.03),
+shows the effect of increasing the delays $t_q$ and $t_u$.
+Importantly, the range on the delay parameters is larger than in previous benchmarks ($U(0, 0.04)$
+rather than $U(0, 0.01)$).  Figure \ref{fig:analysis_delay} shows that if the delay is short (less than 0.02), the controller performs well, and if it is very large (greater than 0.03),
 the controller performs poorly.  However, for delays between 0.02 and 0.03,
-the controller will sometimes perform well and sometimes perform poorly.  The
+the controller sometimes performs well and sometimes performs poorly.  The
 difference is due to the other random parameters in the system.  Interestingly,
-running SpiNNaker actually performs better on this
+SpiNNaker performs better on this
 task than the CPU (the GPU data is equivalent to the CPU and is not shown).  
@@
 Ethernet interface to SpiNNaker, rather than the high-speed I/O system that
-is meant for motor control.  Further analysis is needed to determine exactly why this is.
+is meant for motor control.  Further analysis is needed to determine exactly why this is the case.
 
@@
 is a crucial benchmark, as the complexity of the task itself quickly
-increases with $N$ (since there the function computing the force applied
-at each joint is a function of \emph{all} the joint angles $q$).  This means
-that the external force the neural system must learn to predict increases
-in complexity exponentially as $N$ increases.  The quality of the control
+increases with $N$ because the function computing the force applied
+at each joint is an interaction of \emph{all} the joint angles $q$.  Consequently,
+the number of parameter interactions in ithe external force that the neural system must learn to predict increases
+in exponentially as $N$ increases.  The quality of the control
 is thus dependent on how good an approximation the neurons can make of this
@@
 interacting forces on 15 joints is possible with 500 neurons.  This gives
-an indication of how many neurons are needed for different tasks, and can
-be extended to larger systems.
+an indication of how many neurons are needed for different tasks, and suggests
+that this controller could be used to control larger systems than those tested here.
 
@@
 
-The benchmarks presented here can be improved and further developed in many
+The benchmarks presented here can be improved and further developed in several
 ways.  Most obviously, we need to benchmark more hardware, and in particular
-we would most like to consider analog neuromorphic hardware.  Interestingly, these
-benchmarks can also be run on software simulations of proposed hardware.
+we note that none of the systems tested here are analog neuromorphic hardware.
+While getting access to such hardware can be difficult, we believe the fact
+that our benchmark is easily shared with others as source code and interacts
+with existing hardware using a Python interface will help this process.
+Interestingly, it is worth noting that these
+benchmarks can also be run on software simulations of hardware (analog or
+digital), and could even be used to help form design decisions about hardware
+that has not yet been produced.
 
@@
 
-In addition, we would like to clarify and expand the space of physical tasks
-being approximated by the minimal simulation.  To do this, we will use the
-previously discussed method to 
-calibrate the model across a variety of different physical systems.  Of
-particular interest will be the space of random external force functions that
-are randomly generated.  Right now, this is a simple set of randomly
-generated vectors with a few common nonlinearities.  However, if we can
+In addition, other classes of benchmarks could rely on expanded or completely
+different minimal simulations.  For example, other physical systems could
+be used to calibrate the minimal simulation.  This would lead to other classes
+of randomly generated external forces that may be more (or less) difficult
+for the neuromorphic system to learn.  If we
 identify classes of tasks that we are likely to want to control, we can
-create minimal simulations across those sorts of tasks.  For example, it
-would be possible to randomly generate $N$-joint arms with random arm
+create modify those randomly generated forces to ones that are more appropriate
+for different tasks.  For example, it
+may be of interest to randomly generate $N$-joint arms with random arm
 lengths and random masses, and derive (an approximation of) the actual forces 
-that would be seen in that situation.  However, when doing this careful
-consideration would be needed to make sure that the randomly generated simulated arm
-had motors powerful enough to lift itself against gravity.
+that would be seen in those situations.
 

### 2015-08-16 paper/paper.tex
Merged last of Chris' edits! And a whole 20 mins left to the deadline...

@@
 
+All three hardware systems drastically improve performance on this task, as
+compared to the non-adaptive controller.
+
 \subsection{Computational Power Benchmark}
@@
 
-The approach to benchmarking presented here can also be applied to other tasks,
-defining minimal simulations and evaluating hardware across a random sampling
-of that space.
+The approach to benchmarking presented here can be applied to a wide variety of other tasks,
+by defining different minimal simulations and evaluating hardware and algorithms across a random sampling
+of simulation parameters.
 
-The most straightforward expansion would be to consider motor control where
+One possibility is to consider motor control where
 the state being controlled \emph{is not the motor state itself}.  That is,
@@
 visual feedback of where the tip of its hand was, without knowing the
-geometry of its own hand or the position of the camera with relation to that
+geometry of its own hand, or the position of the camera with relation to that
 hand (\citealt{adaptiveJacobian}).  While it is unlikely that such a
@@
 
-We also believe this approach to developing benchmarks would be applicable
-to more complex and abstract tasks.  For example, much of the original work
+More generally, we believe that this approach to developing benchmarks is applicable
+to entirely different kinds of tasks.  For example, much of the original work
 on minimal simulation involved robot navigation and obstacle avoidance.  
@@
 minimal simulation of an environment with fixed walls, a mobile robot, and
-various sensory systems.  A benchmark involving movement between locations
+various sensory systems.  Using such a simulation, a benchmark involving movement between locations
 while avoiding collisions could be defined, and the environment could be
 enhanced with moving obstacles and a richer set of sensors.  The eventual
-goal would be to achieve reliable movement in an environment that includes
+goal might be to achieve reliable movement in an environment that includes
 people.
 
-Another direction to take this sort of benchmark would be classical and
-operant conditioning.  This sort of learning is extremely common in living
-creatures, and is vital for dealing with novel environments.  As such,
-it is natural to define tasks involving learning the associates between
+Another example is the application of this method to classical and
+operant conditioning.  Conditioning effects are extremely common in living
+creatures, and are clearly evident when animals are exposed to novel environments.  As such,
+it is natural to define benchmark tasks involving learning the associations between
 sensory events in the environment (akin to classical conditioning) and the
@@
 
-The primary purpose of this paper is to describe a new method for benchmarking
+We have described a new method for benchmarking
 neuromorphic hardware that addresses the problem of reliably benchmarking
-in a complex task that involves interaction with an environment.  This
-method involves building a ``minimal'' simulation; one that is extremely
+complex tasks that involve interaction with an environment.  This
+method involves building a minimal simulation; a simulation that is extremely
 simple in terms of required computation, but that has a high degree of
-adjustable variability.  By benchmarking across that whole space of
-possibilities, we can identify hardware that performs well across that whole
+adjustable variability.  By benchmarking across a space of
+possibilities, we can identify hardware that performs well across that
 space, and is thus likely to be useful in real-world situations.  In order
-to identify what real-world situations are covered by a minimal simulation,
-we can calibrate the variability in that simulation with particular physical
+to identify which real-world situations are covered by a minimal simulation,
+we can tune the variability in the simulation to particular physical
 systems.
@@
 We demonstrated this approach by defining a minimal simulation and a task
-based around adaptive motor control.  We presented a basic algorithm that
-can use neuromorphic hardware to improve performance on this task.  While this
-algorithm can be run on a variety of hardware, other algorithms could also
-be used on this task, and there are many implementation possibilities.
+appropriate for adaptive motor control.  We presented an algorithm that
+can use neuromorphic hardware to improve performance on this task over that
+of a standard non-adaptive controller.  
 Importantly, by measuring performance while adjusting the distributions of
@@
 setting parameters in different ways.  We believe this sort of flexibility
-is important in a benchmark, as it lets researchers be explicit about what
+is important in a benchmark method, as it lets researchers be explicit about what
 their hardware is good at, while still using the same basic and shareable
@@
  \textbf{\refstepcounter{figure}\label{fig:system} Figure \arabic{figure}.}{ 
-     An abstract physical system that could be controlled.  In order to hold
+     A simple example of a physical embodiment for control.  In order to hold
      the joint $q$ at a desired angle $q_d$, a force must be applied to

### 2015-08-26 paper/paper.tex
Fixed typos

@@
 algorithm has been specifically over-fit to that situation.  For
-that reason, it is useful to have benchmarks that covers a wide range
+that reason, it is useful to have benchmarks that cover a wide range
 of variations on the task.  If the hardware performs well across that
@@
 To run a benchmark using the proposed
-minimal simulation approach, there are three main steps: 1. identify
+minimal simulation approach, there are four main steps: 1. identify
 the neuromorphic hardware to be tested; 2. construct the minimal
 simulation; 3. determine a metric (e.g. root-mean-squared error; rmse)
-to record; 5. specify distributions for any parameters in the
+to record; 4. specify distributions for any parameters in the
 simulation.  We then perform multiple runs of the simulation, each

### 2015-08-26 paper/paper.tex
another typo

@@
 
-However, it is also clear that performance these benchmarks is a result of
+However, it is also clear that performance on these benchmarks is a result of
 a combination of the hardware itself, the algorithm being run, and the system that

### 2015-09-30 paper/paper.tex
changed embodied to closed-loop

@@
 
-\title[Embodied Neuromorphic Benchmarks]{Embodied Neuromorphic Benchmarks}
+\title[Closed-loop Neuromorphic Benchmarks]{Closed-loop Neuromorphic Benchmarks}
 
@@
     Evaluating the effectiveness and performance of neuromorphic hardware is
-    difficult.  It is even more difficult when the task of interest is an
-    embodied task; that is, a task where the output from the neuromorphic
+    difficult.  It is even more difficult when the task of interest is a
+    closed-loop task; that is, a task where the output from the neuromorphic
     hardware affects some environment, which then in turn affects the
-    hardware's future input.  However, embodied situations
+    hardware's future input.  However, closed-loop situations
     are one of the primary potential uses of neuromorphic hardware.  To address
-    this, we present a methodology for generating embodied benchmarks that makes use
+    this, we present a methodology for generating closed-loop benchmarks that makes use
     of a hybrid of real physical embodiment and a type of ``minimal'' simulation.
@@
 
-In this paper, we focus on the development of \emph{embodied} benchmarks.
+In this paper, we focus on the development of \emph{closed-loop} benchmarks.
 These are dynamic tasks where the output of the neuromorphic hardware \emph{influences
@@
 
-We believe embodied benchmarks should be of particular interest to
+We believe closed-loop benchmarks should be of particular interest to
 neuromorphic research, given that the most compelling applications of neuromorphic
 hardware are likely to be in this domain of embedded and interactive control
-of robotic or other physical systems.  However, embodiment itself raises a number
+of robotic or other physical systems.  However, the closed loop itself raises a number
 of issues that complicate the development of such benchmarks.  Rather than
 simply providing a data file of inputs and desired outputs, the benchmark
-must either specify a full physical system for that embodiment, or it must 
+must either specify a full physical system to be controlled, or it must 
 provide software for a simulation of that system.  As we discuss below, both
@@
 
-\section{Embodied Benchmarks}
+\section{Closed-Loop Benchmarks}
 
-An embodied benchmark task is one where the system we are studying has a 
+A closed-loop benchmark task is one where the system we are studying has a 
 two-way interaction with some sort of environment.  That is, the outputs from
@@
 
-To define an embodied benchmark, we need to be explicit about the embodiment.  If
+To define a closed-loop benchmark, we need to be explicit about the interaction with the environment.  If
 a robot is to be controlled, we need to specify all of the details of that robot.  What motors does it have?  How are they
@@
 we spent the considerable amount of research effort needed to define a
-simulated environment for an embodied benchmark, running that simulation 
+simulated environment for a closed-loop benchmark, running that simulation 
 fast enough to interact with the desired hardware would require more
@@
 We are thus left with a situation where any benchmark we might define for
-a rich, embodied task will be either impractical for different researchers to
+a closed-loop task will be either impractical for different researchers to
 run (if it is physically embodied), or inapplicable to real-world situations
@@
 This approach was first suggested in the context of evolutionary robotics.  
-Notably, the problem faced by embodied neuromorphic benchmarking is remarkably 
+Notably, the problem faced by closed-loop neuromorphic benchmarking is remarkably 
 similar to that faced earlier by these researchers. 
@@
 
-To demonstrate this approach to creating embodied neuromorphic benchmarks,
+To demonstrate this approach to creating closed-loop neuromorphic benchmarks,
 we now consider a basic control task.  Suppose we have a system with a number
@@
 First, the multiplication by $d$ can be done on the
-output from the neuromorphic hardware.  Any embodied neuromorphic system will have some method that takes
+output from the neuromorphic hardware.  Any closed-loop neuromorphic system will have some method that takes
 the neural output from the hardware and sends it to the motors (or to the simulation
@@
 differences in neuron model across that hardware does not significantly
-impact performance.  Given that embodied models rely on real-time simulation,
+impact performance.  Given that closed-loop models rely on real-time simulation,
 it is also important to determine how many neurons each piece of hardware is capable of running in real time.
@@
 run many more neurons than this in real time, if those neurons were not connected
-to an environment.  However, that would not be useful for an embodied task.
+to an environment.  However, that would not be useful for a closed-loop task.
 

### 2015-10-01 paper/paper.tex
Added conclusion about adaptive control

@@
     that focus on motor control for an arbitrary system with
-    unknown external forces.
+    unknown external forces.  Using these benchmarks, we show that an error-driven
+    learning rule can consistently improve motor control performance across
+    a randomly generated family of closed-loop simulations, even when there are
+    up to 15 interacting joints to be controlled.
 
@@
 
+While the primary purpose of this paper is in decribing the benchmarking
+methodology, it is also worth noting that these benchmarks indicate
+that the neuromorphic learning rule under investigation here is extremely robost.  As
+shown in \ref{fig:analysis_neurons_cpu}, even just 500 neurons can consistently
+adapt to control a \emph{randomly generated} 15-joint body simulation, and
+deal with larger delays and noise than were seen in the example 1-joint
+physical embodiment.  Since this learning system is robust across such a
+wide range of conditions, and since it is efficiently implementable in a
+wide variety of neuromorohic hardware, we feel it is worth further study.  This must include 
+both a wider variety of minimal simulation benchmarks and also a few more 
+traditional benchmarks.  These traditional
+benchmarks would be particular real physical systems (specific robot arms,
+for example), but testing on those would only reveal performance on those
+particular arms.  As we argued here, benchmarking against a wide variety of 
+randomly generated minimal simulation systems is needed to demonstrate the
+space of potential situations in which neuromorphic adaptive control performs well.
+
 The benchmarks described above all use the same underlying minimal simulation
@@
 
+Finally, we note that the benchmarking results show that this learning rule
+can consistently improve control performance across a wide variety of randomly
+generated situations, and is suitable for implementation on a wide variety
+of neuromorphic hardware.  Given this promising result, we will be further
+evaluating it on specific physical embodiments, and comparing it to more
+complex variants of PID control.
+
 \section*{Disclosure/Conflict-of-Interest Statement}

### 2015-10-01 paper/paper.tex
Added model summary

@@
 
+For each of three hardware platforms, we implemented the neural control
+system with the learning rule described above.  That is, we started with a
+standard non-neural PD controller that produced an output $u$.  The state
+information $q$ was fed into a group of neurons using randomly generated
+input weights, producing output activity $A$.  The actual output to the motor
+was $u+Ad$ where $d$ is a vector of learned weights, initialized to all
+zeros.  The learning rule was $\Delta d=\alpha A \times u$, and the
+learning rate $\alpha$ was fixed at 0.001.
+
 The first hardware tested on this benchmark is an Intel i5-3337U CPU running

### 2015-10-01 paper/paper.tex
Tweaked captions for figs 4,5,6

@@
 that the neuromorphic learning rule under investigation here is extremely robost.  As
-shown in \ref{fig:analysis_neurons_cpu}, even just 500 neurons can consistently
+shown in Figure \ref{fig:analysis_neurons_cpu}, even just 500 neurons can consistently
 adapt to control a \emph{randomly generated} 15-joint body simulation, and
@@
      Each system is running 500 neurons.  The hardware do not statistically
-     significantly differ, and are all clear improvements over the having
-     no adaptation at all.
-     Scatterplots show individual runs, the 
+     significantly differ, but are all statistically significant improvements 
+     over no adaptation (p<0.05).
+     Scatterplots show individual runs (with random jitter on the x-axis
+     to avoid overlap), the 
      shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.}
@@
  \textbf{\refstepcounter{figure}\label{fig:analysis_compute} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems
-     in terms of their performance when running as many neurons as they are capable of in real time.  Scatterplots show individual runs, the 
+     in terms of their performance when running as many neurons as they are capable of in real time.  Scatterplots show individual runs (with random
+     jitter on the x-axis to avoid overlap), the 
      shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  In this case,
-     all three hardware systems perform equally well.}
+ there is no statistical difference between the three hardware systems (p>0.05).}
 \end{figure}
@@
  \textbf{\refstepcounter{figure}\label{fig:analysis_power} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems
-     in terms of their performance when running as many neurons as they are capable of per 0.1 Watt of power consumption.  Scatterplots show individual runs, the 
+     in terms of their performance when running as many neurons as they are capable of per 0.1 Watt of power consumption.
+Scatterplots show individual runs (with random jitter on the x-axis to avoid overlap), the
      shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  The SpiNNaker system is a statistically
- significant improvement over the other two.}
+ significant improvement over the other two (p<0.05).}
 \end{figure}

### 2015-10-01 paper/paper.tex
Edits to address reviewer 1

@@
 this trap, giving high scores to hardware that does not turn out to
-functional well when deployed in real situations.
-
-Furthermore, neuromorphic hardware has another constraint that severely limits
-the utility of simulation.  Typically, when a simulation is
-too simple to reflect reality, we add details to the simulation itself.
+function well when deployed in real situations.
+
+A variety of robotics simulators, such as Webots and Gazebo, already exist
+and are intended for evaluating robotics performance.  These are extremely
+useful, but have two important limitations.  First, they are generally meant
+to evaluate \emph{one particular} robot body, and it is difficult to, for
+example, automatically generate a large number of different physical bodies 
+to evaluate over.  This means that such a system is good for evaluating a
+particular control system for a particular robot body, but is not suitable
+for the more general question of how well the control system will work over
+a large space of different robot bodies.  Second, these simulators tend
+to run slower than real-time.  Typically, when a simulation is
+too simple to reflect reality, more details are added to the simulation itself.
 Incredibly finely detailed simulations can be created, filling in all of the
@@
 This is a fundamental problem, in that neuromorphic
-hardware is often tied to real-time interactions, and there is no way to slow down
+hardware is often tied to real-time interactions, and there can be 
+no way to slow down
 the hardware to match the simulated environment.  This means that even if
@@
 simulated environment for a closed-loop benchmark, running that simulation 
-fast enough to interact with the desired hardware would require more
-resources than are available.
+fast enough to interact with the desired hardware would require significant
+computing resources.  Indeed, one of the major efforts in the Neurorobotics
+section of the Human Brain Project is to develop exactly this sort of 
+computing infrastructure (\citealt{HBProbot}), with a dedicated supercomputer to run the
+physics simulations.  Until this hardware is publically available 
+(and until software
+is available to create a variety of physical robot models), this approach
+is problematic for other researchers.
 
 We are thus left with a situation where any benchmark we might define for
-a closed-loop task will be either impractical for different researchers to
-run (if it is physically embodied), or inapplicable to real-world situations
-(if it is performed in simulation).  Neither option is particularly appealing.
+a closed-loop task will be impractical for different researchers to
+run (if it is physically embodied), inapplicable to real-world situations
+(if it is a simulation that is simple enough to run in real-time), or impossible to connect to
+some neuromorphic hardware (if it is a simulation that runs slower than 
+real-time).  We thus need a new approach to provide a sharable real-time
+simulation that is robust enough that neuromorphic hardware that learns to
+deal with the simulation might also be able to deal with reality.
 
@@
 by not including all the details necessary to model these complex physical
-interactions.
+interactions.  This approach was successfully used to develop models of
+multi-legged walking robots (\citealt{jakobiThesis}; \citealt{Meyer2003})
+and vision-based tracking of a moving object (\citealt{Nolfi2000}).
 
@@
 
-Given the success of this approach for evolutionary robotics, we propose
+Although minimal simulation has not previously been used outside the domain of evolutionary robotics, we propose
 using minimal simulation for neuromorphic benchmarks.  We would argue that
@@
 
-\subsection{Cheap Robotics}
+\subsection{Cost-Effective Robotics}
 
@@
 
-To keep the physical aspect simple, we recommend cheap, widely-available
+To keep the physical aspect simple, we recommend cheap, cost-effective, 
+widely-available
 components.  This allows a greater chance for other researchers to have
@@
 It is important to note that there is a theoretical advantage to
-using cheap robotics hardware for benchmarking, in addition to the practical
+using simple robotics hardware for benchmarking, in addition to the practical
 advantages.  In particular, we \emph{do not want benchmarks that rely on
@@
 real world, then we should choose one that is not high-precision.
-For this reason, we believe using cheap Lego robots is actually more informative
+For this reason, we believe using Lego robots is actually more informative
 for benchmarking than expensive high-precision robots.\footnote{Of course,
@@
 
+It should be noted that, while this algorithm fits well into neuromorphic
+hardware, other hardware might be better (in terms of 
+accuracy, energy efficiency, cost, or even development time).  Answering
+this sort of question is exactly why we need to use a benchmark that can
+compare multiple different hardware implementations of this algorithm.  
+Furthermore, since some hardware may be better in different situations,
+we need a benchmark that has flexible parameters, rather than one that is
+based on a single particular physical system.
+
 \subsection{Online and Offline Learning}
@@
 
-\subsection{Calibrating the Minimal Simulation via Cheap Robotics}
+\subsection{Calibrating the Minimal Simulation via Cost-Effective Robotics}
 
@@
 simulation.  We then perform multiple runs of the simulation, each
-time choosing different values from the parameter distributions.  We
+time choosing different values from the parameter distributions.  For
+each run, we reset the hardware to its initial state, so there is no
+learning from one run to the next.  This means our metric indicates
+how well the system will perform on a single environment, rather than attempting
+to use the same learned parameters across different environments.  We
 can then plot how the metric varies as a function of a particular parameter
@@
 All three hardware systems drastically improve performance on this task, as
-compared to the non-adaptive controller.
+compared to the non-adaptive controller.  Statistical significance was
+measured with a two-tailed t-test.
 
@@
 In Figure \ref{fig:analysis_basic}, all three systems perform equally well.
+This means that the timing and accuracy differences between the fixed-point
+asynchronous SpiNNaker implementation and the floating-point synchronous
+CPU/GPU impementations do not affect performance on this task.
 However, on that benchmark all three systems are implementing exactly 500
@@
 it is also important to determine how many neurons each piece of hardware is capable of running in real time.
-This is shown in Figure \ref{fig:analysis_compute}.  With the current implementation, the CPU can
-run 5200 neurons, the GPU 1500 neurons, and a single SpiNNaker core (of which
-there are 18 on one chip) can run 500
-neurons.\footnote{The full chip can run much more than 500 neurons; here we use just a single one of the cores on the SpiNNaker chip.}  These values were measured empirically with the current versions
+This is shown in Figure \ref{fig:analysis_compute}.  With the current
+implementation, including both the Leaky Integrate-and-Fire neuron model
+and the learning rule, the CPU can
+run 5200 neurons, the GPU 1500 neurons, and a single SpiNNaker core can
+run 500 neurons (or 500 $\times$ 16 = 8000 neurons for the whole chip).
+These values were measured empirically with the current versions
 of the reference Nengo implementation, the Nengo OpenCL implementation,
@@
 
-While the primary purpose of this paper is in decribing the benchmarking
+While the primary purpose of this paper is in describing the benchmarking
 methodology, it is also worth noting that these benchmarks indicate
-that the neuromorphic learning rule under investigation here is extremely robost.  As
+that the neuromorphic learning rule under investigation here is quite robost.  As
 shown in Figure \ref{fig:analysis_neurons_cpu}, even just 500 neurons can consistently
@@
 wide range of conditions, and since it is efficiently implementable in a
-wide variety of neuromorohic hardware, we feel it is worth further study.  This must include 
+wide variety of neuromorphic hardware, we feel it is worth further study.  This must include 
 both a wider variety of minimal simulation benchmarks and also a few more 
@@
 lengths and random masses, and derive (an approximation of) the actual forces 
-that would be seen in those situations.
+that would be seen in those situations.  In particular, we feel benchmarks
+based on the biologically-inspired ``soft-robotics'' systems (e.g. \citealt{Pfeifer2013})
+would be particularly appropriate for neural control, given the complexity
+involved in generating traditional controllers for them.
 
@@
      significantly differ, but are all statistically significant improvements 
-     over no adaptation (p<0.05).
+     over no adaptation ($p<0.05$).
      Scatterplots show individual runs (with random jitter on the x-axis
@@
      shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  In this case,
- there is no statistical difference between the three hardware systems (p>0.05).}
+ there is no statistical difference between the three hardware systems ($p>0.05$).}
 \end{figure}
@@
      shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  The SpiNNaker system is a statistically
- significant improvement over the other two (p<0.05).}
+ significant improvement over the other two ($p<0.05$).}
 \end{figure}

### 2015-10-14 paper/paper.tex
Updated discussion of scaling up

@@
 methodology, it is also worth noting that these benchmarks indicate
-that the neuromorphic learning rule under investigation here is quite robost.  As
+that the neuromorphic learning rule under investigation here is quite robust.  As
 shown in Figure \ref{fig:analysis_neurons_cpu}, even just 500 neurons can consistently
@@
 
-The approach to benchmarking presented here can be applied to a wide variety of other tasks,
-by defining different minimal simulations and evaluating hardware and algorithms across a random sampling
-of simulation parameters.
-
-One possibility is to consider motor control where
-the state being controlled \emph{is not the motor state itself}.  That is,
+While the particular minimal simulation shown here suggests that
+this adaptive control algorithm is worth further investigation, the overall
+goal of this paper is to present the general idea of using minimal simulation
+as a way to benchmark neuromorphic hardware.  That is, we believe this same
+approach could be scaled up to other, more complex, closed-loop tasks.
+Importantly, benchmarking these other tasks would require both the creation
+of new minimal simulations \emph{and} the specification of new algorithms
+suitable for performing those tasks.  These algorithms would then be
+implemented with the neuromorphic hardware and connected to the minimal
+simulations to construct new benchmarks.
+
+As a first step towards scaling up, consider the more complex task of
+controlling a system where the values to be controlled are not the 
+joints themselves.  For example, 
 suppose we want to control the position of a hand $x$, but our output $u$
@@
 is some function of $q$, but this function may be unknown or highly complex.
-Crucially, there is a learning rule that is similar to the one discussed above
+This is often expressed as $\dot{x}=J(q)\dot{q}$, where $J(q)$ is the
+Jacobian.  In order to successfully control $x$ (the hand), the system
+needs to learn the relationship $J$ that indicates how adjusting various
+joints $q$ will affect the position of the hand.  Crucially, there is a
+learning rule similar to the one discussed above
 that can learn this mapping (\citealt{adaptiveJacobian}), and we have had
-some success in using it for particular arm control tasks (\citealt{dewolf2014}).
-This would give a benchmark with a much more flexible set of capabilities;
-for example a non-neural version of this learning algorithm was able to
-learn to control a robot arm to follow a certain path, given only the
-visual feedback of where the tip of its hand was, without knowing the
-geometry of its own hand, or the position of the camera with relation to that
-hand (\citealt{adaptiveJacobian}).  While it is unlikely that such a
-system would work for \emph{any} possible sensorimotor combination, developing
-such a benchmark could identify the space of such embodiments that a
-particular algorithm running on particular neuromorphic hardware would be
-good at controlling.
-
-More generally, we believe that this approach to developing benchmarks is applicable
-to entirely different kinds of tasks.  For example, much of the original work
-on minimal simulation involved robot navigation and obstacle avoidance.  
-\cite{Jakobi97evolutionaryrobotics} created a
-minimal simulation of an environment with fixed walls, a mobile robot, and
-various sensory systems.  Using such a simulation, a benchmark involving movement between locations
-while avoiding collisions could be defined, and the environment could be
-enhanced with moving obstacles and a richer set of sensors.  The eventual
-goal might be to achieve reliable movement in an environment that includes
-people.
-
-Another example is the application of this method to classical and
-operant conditioning.  Conditioning effects are extremely common in living
+some success in using it for particular arm control tasks
+(\citealt{dewolf2014}).  So far we have only tested this algorithm
+in the context of one particular arm, but it was successful in learning
+this relationship, and thus learning to correctly move its hand given
+an unknown arm geometry.  To establish that this is a generally
+useful task for neuromorphic hardware, we need to benchmark this rule against
+a large family of different arms (and other systems to be controlled).  This
+can be done by generating minimal simulations very similar to the
+one presented here; the main difference is that there would also be a
+randomly generated Jacobian function $J(q)$.  It should also be noted that
+in this context, the dimensionality of $x$ and the dimensionality of $q$
+are separate variables.  It may be that some algorithms work well when
+$q$ is much larger than $x$, while others work best when they are similar.
+Exploring this relationship is fairly straightforward with minimal simulation,
+and would be an important result to know when choosing neuromorphic hardware
+for a particular new situation.
+
+Given this, we believe that the combination of minimal simulation and
+neuromorphic hardware is useful for adaptive control problems in general,
+whether the adaptation is in terms of an additive bias term to compensate
+for external forces such as gravity (as seen in the benchmarks presented
+in this paper) or if it is in terms of learning the Jacobian term relating
+the controlled variables $q$ to the desired target space $x$ (as in
+the adaptive Jacobian model discussed in the previous paragraph).  This
+should allow systems to adapt to both unknown external forces and to unknown
+bodily geometries.  However,
+it is less clear whether this approach will scale to more complex robotics
+tasks.
+
+One more complex robotic task where this approach might be applicable is
+navigation and obstacle avoidance.  Here, we would need both a more complex
+minimal simulation for the environment, and an explicit neuromorphic
+algorithm capable of performing this avoidance.  The minimal simulation itself
+would need to include some sort of sensory modality (vision,
+range sensing, or both), and movement in a two-dimensional environment
+(probably wheeled movement, for simplicity).  To run such a simulation in
+real-time, we would use many of the same tricks and simplifications used
+in Jakobi's original work (\citealt{Jakobi97evolutionaryrobotics}).  These
+included making separate simulations for corridors and intersections (rather
+than generic simulations for any possible geometries), using noisy lookup
+tables (rather than detailed physics simulations), treating collisions as
+failures (rather than modelling them), and using shifting random dot patterns
+for visual stimuli (rather than high-fidelity image rendering).  Given
+Jakobi's success at building high-speed simulations over 20 years ago, we
+believe real-time simulations of this type are feasible now.
+
+However, having such a simulation is only half of what is required.  We would
+also need a control algorithm suitable for such a situation.  This is,
+itself, a topic of much research, and there is no clear best approach.
+We have been exploring the use of reinforcement learning in
+neural models (\citealt{Stewart2012a, Rasmussen2014b}), and note that these 
+make use of the same learning algorithm as described here, with additional
+neural components needed to implement action selection.  In this case, the
+learning rule would adjust the system's estimate of which action is most
+appropriate given the current sensory state.  We are currently investigating
+this approach further.
+
+As a more speculative possibility, we also intend to apply this approach
+to tasks involving classical and operant conditioning.  Conditioning effects
+are extremely common in living
 creatures, and are clearly evident when animals are exposed to novel environments.  As such,
@@
 associations between actions and desired sensory states (akin to operant
-conditioning).  Again, the approach of minimal simulation could be used
-to generate random complex sensory mappings, and various neuromorphic
-learning algorithms could be evaluated in terms of their ability to
-discern these patterns.  Importantly, this would also include the ability
-to unlearn these patterns as the environment changes.  Interestingly, the
-neuron-based classical conditioning learning rule used by \cite{verschure2003}
-is quite similar to the learning rule used in the adaptive control benchmark
-here.
+conditioning).  In this case, the minimal simulations would consist of a
+set of small, controlled rooms with controllable buttons and stimuli,
+matching the sort of ``Skinner Box'' environments used in experimental
+psychology.  The minimal simulation will also require a basic simulated
+body, capable of movement, pushing buttons, and observing stimuli.  The
+tasks would consist of pairing stimuli together and determining if the
+learning algorithm is able to respond correctly.  For example, a model
+might have a built-in response where it will salivate when presented with
+food.  If the sound of a bell is paired with the presentation of food, it
+should learn to salivate with presented with just the sound of a bell.
+Importantly, there are extensive results showing the rate at which such
+associations are learned and un-learned in various animals.  Furthermore,
+we would test the ability to learn associations that are separated in time (delayed conditioning), and to recover associations that had been previously learned (spontaneous recovery).
+Interestingly, there already exist
+neuron-based classical conditioning learning rules that may be suitable
+for such implementation, given their similarity to the learning rule used in
+the adaptive control benchmark (\citealt{verschure2003}).
 

### 2015-11-12 code/benchmark_ev3.py
Added code for running benchmark on robot

@@ -8,12 +8,14 @@ import embodied_benchmarks as bench
 import benchmark
 
 class SingleArmEV3(benchmark.Benchmark):
+    links = {}
+
     def params(self):
         self.default('Kp', Kp=2.0)
         self.default('Kd', Kd=1.0)
         self.default('Ki', Ki=0.0)
         self.default('tau_d', tau_d=0.001)
-        self.default('T', T=0.5)
+        self.default('T', T=20)
         self.default('IP address of EV3', address='10.42.0.3')
         self.default('motor index', motor_index=0)
         self.default('desired path', path='step')
@@ -26,7 +28,9 @@ class SingleArmEV3(benchmark.Benchmark):
         self.default('radius', radius=1.0)
 
     def model(self, p):
-        link = ev3link.EV3Link(p.address)
+        if p.address not in SingleArmEV3.links:
+            SingleArmEV3.links[p.address] = ev3link.EV3Link(p.address)
+        link = SingleArmEV3.links[p.address]
         self.link = link
         path = '/sys/class/tacho-motor/motor%d/' % p.motor_index
         self.path = path
@@ -115,17 +119,20 @@ class SingleArmEV3(benchmark.Benchmark):
             plt.plot(sim.data[self.p_t], sim.data[self.p_q])
             #plt.plot(sim.data[self.p_t], sim.data[self.p_u])
 
-        rate = len(sim.data[self.p_t]) / p.T
+        rate = len(sim.data[self.p_t]) / float(p.T)
 
         q = sim.data[self.p_q][:,0]
         d = sim.data[self.p_desired][:,0]
         t = sim.data[self.p_t][:,0]
         
         offset = benchmark.find_offset(q, d)
+        #offset = 1
 
         delay = np.mean(t[offset:]-t[:-offset])
 
         diff = d[:-offset] - q[offset:]
+        N = len(q) / 2
+        diff = diff[N:]
         rmse = np.sqrt(np.mean(diff**2))
 
 

### 2015-11-12 code/exp_ev3.py

@@ -1,0 +1,15 @@
+from benchmark_ev3 import SingleArmEV3
+
+import numpy as np
+rng = np.random.RandomState()
+
+start = 45
+
+for i in range(start,start+5):
+    for adapt in [False, True]:
+        SingleArmEV3().run(no_figs=True, T=20, seed=i, max_freq=1,
+                scale_add=1,
+                n_neurons=500,
+                adapt=adapt, 
+                path='random',
+                data_dir='ev3_%s' % adapt)

### 2015-11-12 code/plot_ev3.py

@@ -1,0 +1,81 @@
+import os
+import sys
+import bootstrapci
+
+import numpy as np
+import pylab
+
+fig = pylab.figure(figsize=(6,5))
+ax = fig.add_subplot(1,1,1)
+
+do_logx = False
+
+error_bars = []
+error_means = []
+index = 1
+for adapt in [False, True]:
+    path = 'ev3_%s' % adapt
+
+    data_x = []
+    data_y = []
+    for fn in os.listdir(path):
+        if fn.endswith('.txt'):
+            with open(os.path.join(path, fn)) as f:
+                text = f.read()
+            d = dict()
+            try:
+                exec(text, d)
+            except:
+                continue
+
+            x = d['_adapt']
+            y = d['rmse']
+
+            if do_logx:
+                x = np.log2(x)
+
+            data_x.append(x)
+            data_y.append(y)
+            print x, y
+
+    #data_x = data_x[:30]
+    #data_y = data_y[:30]
+
+    data_x = np.array(data_x)
+    data_y = np.array(data_y)
+
+
+    mean = np.mean(data_y)
+    sd = np.std(data_y)
+    ci = bootstrapci.bootstrapci(data_y, np.mean)
+
+    width = 0.3
+
+    pylab.fill_between([index-width, index+width], mean-sd, mean+sd, color='#aaaaaa')
+    pylab.scatter(np.random.uniform(index-width, index+width, data_y.shape), data_y, s=30, marker='x', color='k', alpha=0.3)
+    error_means.append(mean)
+    error_bars.append([mean-ci[0], ci[1]-mean])
+    index += 1
+
+error_bars = np.array(error_bars)
+
+pylab.errorbar(np.arange(len(error_bars))+1, error_means, yerr=error_bars.T, color='k', linewidth=3, capthick=2, capsize=4)
+
+pylab.xticks([1, 2], ['without adaptation', 'with adaptation\n(500 neurons)'])
+
+def annotate_bar(ax, text, x1, x2, y, text_offset, shrink=50):
+    ax.annotate(text, xy=((x1+x2)*0.5, y+text_offset), ha='center')
+    props = dict(connectionstyle='bar', arrowstyle='-', shrinkA=shrink, 
+                 shrinkB=shrink, lw=2)
[…]

### 2015-11-13 paper/paper.tex
Clarified stats

@@
 All three hardware systems drastically improve performance on this task, as
-compared to the non-adaptive controller.  Statistical significance was
-measured with a two-tailed t-test.
+compared to the non-adaptive controller.
 
@@
 the GPU 38 neurons per 0.1W, and SpiNNaker can run 500 neurons per 0.1W.  
-As shown in Figure \ref{fig:analysis_power}, the neuromorphic hardware
-outperforms the CPU and GPU on this task.
+As shown in Figure \ref{fig:analysis_power}, while the GPU outperforms
+the CPU on this task, the neuromorphic hardware outperforms both of them.
 
@@
 (probably wheeled movement, for simplicity).  To run such a simulation in
-real-time, we would use many of the same tricks and simplifications used
+real-time, we would use many of the same optimizations and simplifications used
 in Jakobi's original work (\citealt{Jakobi97evolutionaryrobotics}).  These
@@
      significantly differ, but are all statistically significant improvements 
-     over no adaptation ($p<0.05$).
+     over no adaptation ($p<0.05$; two-tailed t-test with Bonferroni correction; 400 samples per condition).
      Scatterplots show individual runs (with random jitter on the x-axis
@@
      shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  In this case,
- there is no statistical difference between the three hardware systems ($p>0.05$).}
+ there is no statistical difference between the three hardware systems ($p>0.05$; two-tailed t-test with Bonferroni correction; 400 samples per condition).}
 \end{figure}
@@
 Scatterplots show individual runs (with random jitter on the x-axis to avoid overlap), the
-     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  The SpiNNaker system is a statistically
- significant improvement over the other two ($p<0.05$).}
+     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  
+ All differences are statistically significant ($p<0.05$; two-tailed t-test with Bonferroni correction; 400 samples per condition).}
 \end{figure}

### 2015-11-13 code/plot_basic.py
Updated plot labels, re-ran data

@@ -14,7 +14,7 @@ error_bars = []
 error_means = []
 index = 1
 for n_neurons in [0, 500, 500, 500]:
-    path = 'vary_neurons3_%d' % n_neurons
+    path = 'vary_neurons4_%d' % n_neurons
 
     data_x = []
     data_y = []
@@ -58,14 +58,14 @@ for n_neurons in [0, 500, 500, 500]:
     width = 0.3
 
     pylab.fill_between([index-width, index+width], mean-sd, mean+sd, color='#aaaaaa')
-    pylab.scatter(np.random.uniform(index-width, index+width, data_y.shape), data_y, s=30, marker='x', color='k')
+    pylab.scatter(np.random.uniform(index-width, index+width, data_y.shape), data_y, s=30, marker='x', color='k', alpha=0.3)
     error_means.append(mean)
     error_bars.append([mean-ci[0], ci[1]-mean])
     index += 1
 
 error_bars = np.array(error_bars)
 
-pylab.errorbar(np.arange(len(error_bars))+1, error_means, yerr=error_bars.T, color='k', linewidth=3)
+pylab.errorbar(np.arange(len(error_bars))+1, error_means, yerr=error_bars.T, color='k', linewidth=3, capthick=2, capsize=4)
 
 pylab.xticks([1, 2, 3, 4], ['No adaptation', 'Intel i5-337U CPU\n(500 neurons)', 'Nvidia Tesla C2075 GPU\n(500 neurons)', 'SpiNNaker core\n(500 neurons)'])
 
@@ -81,7 +81,7 @@ annotate_bar(ax, 'p>0.05', 2, 4, 0, 0.08, shrink=80)
 annotate_bar(ax, '', 2, 2.9, 0.03, 0.04, shrink=35)
 annotate_bar(ax, '', 3.1, 4, 0.03, 0.04, shrink=35)
 
-pylab.ylabel('rmse')
+pylab.ylabel('rmse (radians)')
 pylab.ylim(0, 0.25)
 pylab.savefig('plot_basic.png', dpi=300)
 #pylab.show()

### 2015-11-13 code/plot_vary_neurons_all.py

@@ -79,7 +79,7 @@ for D in [1, 5, 10, 15]:
     if index in [3, 4]:
         pylab.xlabel('number of neurons')
     if index in [1,3]:
-        pylab.ylabel('rmse')
+        pylab.ylabel('rmse (radians)')
     pylab.ylim(-0.05, 0.5)
     index += 1
 pylab.savefig('plot_vary_neurons.png', dpi=300)

### 2015-11-13 paper/paper.tex
Added analysis of lego robot improvement

@@
 the first 5 seconds, the system is able to bring $q$ much closer to the
-desired $q_d$.
+desired $q_d$.  Figure \ref{fig:ev3runs} shows the average improvement over
+50 experimental runs with different randomly-generated desired target paths
+$q_d(t)$.  Adaptation provides a clear improvement.
 
@@
 \begin{center}
+\includegraphics[width=9cm]{figures/plot_ev3}
+\end{center}
+ \textbf{\refstepcounter{figure}\label{fig:ev3runs} Figure \arabic{figure}.}{ The effect
+     of adaptive control on a single-joint lego robot (Figure \ref{fig:lego}).
+     Each run uses a randomly generated desired trajectory $q_d(t)$ over
+     20 seconds, and root-mean-squared-error is computed over the last
+     10 seconds only.  The adaptive algorithm provides a significant improvement
+     ($p<0.05$; two-tailed t-test; 50 samples).
+     Scatterplots show individual runs (with random jitter on the x-axis
+     to avoid overlap), the 
+     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.}
+\end{figure}
+
+\begin{figure}[h!]
+\begin{center}
 \includegraphics[width=18cm]{figures/plot_basic}

### 2015-11-13 paper/paper.tex
Added labelled breakdown of lego robot
Added analysis of EV3 delay

@@
 benchmark is meant to cover, we can use it to calibrate the parameters of
-the simulation.  For example, communication with the EV3 happens around
-200Hz, meaning that there must be a delay on the order of 0.005 seconds.  Given
+the simulation.  For example, to characterize the communication delay
+between the computing hardware and the EV3 robot, we simply measure the number
+of times per second we can send a motor command $u$ and read the position of
+the motor $q$ per second.  This works because the \texttt{ev3\_link} software
+is entirely synchronous and only responds with motor positions when it
+processes a command to do so.  This rate of communication averaged 154-156Hz (95\%
+bootstrap confidence interval over 100 trials) with a standard deviation of
+3.3-4.7Hz (95\% bootstrap C.I.).  This indicates a round-trip
+delay on the order of 0.006 seconds.  Given
 this, we set the delays in the simulation to be uniformly chosen between
-0 and 0.01.  Importantly, we do not need to exactly measure the delay in the
-EV3 robot --- we just make sure that the minimal simulation covers even worse
-cases.
+0 and 0.01, so that the minimal simulation covers delay conditions even
+worse than those seen in the EV3 robot.
 
@@
 \begin{center}
-\includegraphics[width=18cm]{figures/lego}
+\includegraphics[width=18cm]{figures/lego_labelled}
 \end{center}
@@
      simple physical robot embodiment for calibrating the minimal simulation.
+     All components come with the Lego Mindstorms EV3 kit, and are shown on
+     the left (2 large motors; 2 1x11 beams; 1 1x13 beam; 12 pins;
+     1 EV3 brick; 1 connector cable).
      To rotate the central motor to the desired position $q$, enough force

### 2015-11-14 paper/paper.tex
Added calibration info on motor power

@@
 For sensor noise, we note that the EV3 rotation encoders for the motors (the
-devices that measure $q$) have a resolution of 0.0175 (1 degree).  This is a
+devices that measure $q$) have a resolution of 0.0175 radians (1 degree).  This is a
 very different sort of noise than the Gaussian noise used in the simulation,
@@
 so strong that the system does not have enough strength to counteract them.
-On the physical robot, in the worst-case scenario ($q=\pi/2$ or $-\pi/2$),
-the motors must be driven at around 0.3 times their maximum strength to
-balance the force of gravity.  (This can, of course, be adjusted by changing
-the weight and its position on the end of the arm).  If we arbitrarily fix $K_f$ to 1 and randomly
+To measure this on the physical robot, we applied a standard PID controller
+with a target $q_d$ of $\pi/2$ (the position at which maximum torque must
+be applied to counteract gravity).  After giving the system 5 seconds to
+stabilize, we recorded the required motor command sent to the robot (from -1
+to +1).  On average, this was 0.11 to 0.16 (95\% bootstrap confidence interval
+over 50 trials), with a standard deviation of 0.07 to 0.12 (95\% bootstrap C.I.),
+and a maximum value of 0.36.  Considering this a worst-case scenario, if we
+arbitrarily fix $K_f$ to 1 and randomly
 generate external forces given the process described above, then 95\% of
 the time we get values between -3.75 and +3.75.  Since we want the motors to
-be strong enough to compensate for forces in that range, we set $T$ to 10.
+be strong enough to compensate for forces in that range, we set $T$ to 10 ($\approx 3.75 / 0.36$).
 
